### 2024-08-15 Test1.Rmd
First substantive commit, I believe.  Initial mortality analysis.

---
title: "Test1"
author: "Jeremiah Groom"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

To find the mean and variance of proportions where the number of values for each proportion are not necessarily equal, you can follow these steps:

### Mean of Proportions

1. **Collect Data**: Suppose you have multiple groups, each with a different number of observations. Let $n_i$ be the number of observations in the $i$-th group, and let $p_i$ be the proportion of successes in the $i$-th group.

2. **Total Observations and Successes**: Calculate the total number of observations $N$ and the total number of successes $S$ across all groups.
   $$
   N = \sum_{i=1}^{k} n_i
   $$
   $$
   S = \sum_{i=1}^{k} (p_i \times n_i)
   $$

3. **Overall Proportion (Mean of Proportions)**: The mean of the proportions is the overall proportion of successes across all groups.
   $$
   \hat{p} = \frac{S}{N} = \frac{\sum_{i=1}^{k} (p_i \times n_i)}{\sum_{i=1}^{k} n_i}
   $$

### Variance of Proportions

1. **Calculate Individual Variances**: For each group, the variance of the proportion $p_i$ is:
   $$
   \text{Var}(p_i) = \frac{p_i (1 - p_i)}{n_i}
   $$

2. **Weighted Variance**: Since the groups have different sizes, you need to weight the variances by the group sizes. The overall variance of the proportions can be calculated as follows:

   $$
   \text{Var}(\hat{p}) = \frac{1}{N} \left( \sum_{i=1}^{k} n_i \times \text{Var}(p_i) \right)
   $$
   where $\text{Var}(p_i) = \frac{p_i (1 - p_i)}{n_i}$.

3. **Plugging in Individual Variances**: Substitute the individual variances into the formula.
   $$
   \text{Var}(\hat{p}) = \frac{1}{N} \left( \sum_{i=1}^{k} n_i \times \frac{p_i (1 - p_i)}{n_i} \right)
   $$
   $$
   \text{Var}(\hat{p}) = \frac{1}{N} \left( \sum_{i=1}^{k} p_i (1 - p_i) \right)
   $$

### Summary

- **Mean of Proportions**: 
  $$
  \hat{p} = \frac{\sum_{i=1}^{k} (p_i \times n_i)}{\sum_{i=1}^{k} n_i}
  $$

- **Variance of Proportions**:
  $$
  \text{Var}(\hat{p}) = \frac{1}{\sum_{i=1}^{k} n_i} \left( \sum_{i=1}^{k} p_i (1 - p_i) \right)
  $$

By following these steps, you can accurately calculate the mean and variance of proportions, even when the number of values for each proportion are not necessarily equal.