### 2025-09-20 Manuscript.Rmd
I updated the Manuscript.RMD file and created the first iteration of the permutation test (Permutation_Test.R)

---
title: "Tree growth, mortality, and climatic stress in west-coast states, USA"
author:
  - name: "Jeremiah D. Groom"
    affiliations:
      - name: "Groom Analytics LLC"
        address: "1975 SE Crystal Lake Dr., Unit 173"
        city: "Corvallis"
        state: "OR"
        postal-code: "97333"
    orcid: "0000-0002-0110-1036"
    email: "jeremy@groomanalytics.com"
  - name: "Bryce Frank"
    affiliations:
      - name: "USDA Forest Service"
        department: "Pacific Northwest Field Station"
        address: "3625 93rd Ave SW"
        city: "Olympia"
        state: "WA"
        postal-code: "98512"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  officedown::rdocx_document:
    reference_docx: "template_use2.docx"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(kableExtra)   # Enhanced table formatting
library(scales)       # Scale functions for visualization
library(flextable)    # Professional tables for Word output
library(officedown)
library(officer)
library(rvg)
#devtools::install_github("adletaw/captioner")
library(captioner) # For figure captions.

```

```{r other_settings}
fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)
```

```{r constants}

source("Global.R")
source(paste0(CODE.LOC, "Functions.R"))

# Define any custom functions
format_p_value <- function(p) {
  case_when(
    p < 0.001 ~ "< 0.001",
    p < 0.01 ~ sprintf("= %.3f", p),
    TRUE ~ sprintf("= %.2f", p)
  )
}

set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  theme_fun = "theme_vanilla"
)

centeredP <- fp_par(text.align = "center")

```

```{r load-data}
# Some general analysis info + plots
analysis.stats <- read_rds(paste0(RESULTS.OTHER, "analysis.stats.RDS"))

# Percent of plots sampled with REMPER = 9, 10, and 11 years (rounded)
pct.remeas <- as.vector(analysis.stats$pct.remeas)

# Species scientific names
spp.names.use <- analysis.stats$gym.angio.dat

# Growth data for species overall, states, differences
grow.state <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/State_Growth_", CLIM.VAR.USE, ".csv"))
grow.all <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/All_Growth_", CLIM.VAR.USE, ".csv"))
grow.diff <- read_rds(paste0(save.loc.fcn(1), "Processed_6Domain_Data.RDS"))

# Mortality data for species overall and states
mort.state <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/State_Mortality_", CLIM.VAR.USE, ".csv"))
mort.all <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/All_Mortality_", CLIM.VAR.USE, ".csv"))
mort.diff <- read_rds(paste0(save.loc.fcn(2), "Processed_6Domain_Data.RDS"))

# Significant findings, Growth
grow.n.AvB <- nrow(grow.diff$AvB2)
grow.n.A <- nrow(grow.diff$A_LMH2)
grow.n.B <- nrow(grow.diff$B_LMH2)

grow.AvB.sig <- length(grow.diff$AvB2$significant[grow.diff$AvB2$significant == TRUE])
grow.AvB.signeg <- length(grow.diff$AvB2$significant[grow.diff$AvB2$significant == TRUE & grow.diff$AvB2$Means < 0])
grow.A.sig <- length(grow.diff$A_LMH2$significant[grow.diff$A_LMH2$significant == TRUE])
grow.A.signeg <- length(grow.diff$A_LMH2$significant[grow.diff$A_LMH2$significant == TRUE & grow.diff$A_LMH2$Means < 0])
grow.B.sig <- length(grow.diff$B_LMH2$significant[grow.diff$B_LMH2$significant == TRUE])
grow.B.signeg <- length(grow.diff$B_LMH2$significant[grow.diff$B_LMH2$significant == TRUE & grow.diff$B_LMH2$Means < 0])

# Significant findings, Mort
mort.n.AvB <- nrow(mort.diff$AvB2)
mort.n.A <- nrow(mort.diff$A_LMH2)
mort.n.B <- nrow(mort.diff$B_LMH2)

mort.AvB.sig <- length(mort.diff$AvB2$significant[mort.diff$AvB2$significant == TRUE])
mort.AvB.signeg <- length(mort.diff$AvB2$significant[mort.diff$AvB2$significant == TRUE & mort.diff$AvB2$Means < 0])
mort.A.sig <- length(mort.diff$A_LMH2$significant[mort.diff$A_LMH2$significant == TRUE])
mort.A.signeg <- length(mort.diff$A_LMH2$significant[mort.diff$A_LMH2$significant == TRUE & mort.diff$A_LMH2$Means < 0])
mort.B.sig <- length(mort.diff$B_LMH2$significant[mort.diff$B_LMH2$significant == TRUE])
mort.B.signeg <- length(mort.diff$B_LMH2$significant[mort.diff$B_LMH2$significant == TRUE & mort.diff$B_LMH2$Means < 0])

# Summary of significant Growth and Mortality findings for table creation.
datasets.grow <- list(grow.diff$AvB2, grow.diff$A_LMH2, grow.diff$B_LMH2)
datasets.mort <- list(mort.diff$AvB2, mort.diff$A_LMH2, mort.diff$B_LMH2)
datatypes <- list("Above vs. Below", "Above", "Below")
n.comp.grow <- list(grow.n.AvB, grow.n.A, grow.n.B)
n.comp.mort <- list(mort.n.AvB, mort.n.A, mort.n.B)

growth.summ <- pmap(list(datasets.grow, datatypes, n.comp.grow), summ.fig.fcn) %>% bind_rows()
mortality.summ <- pmap(list(datasets.mort, datatypes, n.comp.mort), summ.fig.fcn) %>% bind_rows()

```

```{r table-creation}
# State growth
s.g.dat <- state.table.fcn(grow.state, grow.all, 1)

growth.ft <- flextable(s.g.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
    italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = TRUE)
min.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = TRUE)

# State mortality
s.m.dat <- state.table.fcn(mort.state, mort.all, 2)

mort.ft <- flextable(s.m.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",  
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = FALSE, digits = 2)
min.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = FALSE, digits = 2)

# CWD amounts
min.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.min.mean]
min.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.min.mean]
max.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.max.mean]
max.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.max.mean]

max.q95.diff <- analysis.stats$diff.cwd.stats$q95[analysis.stats$diff.cwd.95max]
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]
max.range.diff <- analysis.stats$diff.cwd.90range.max
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]

# Adding a section identifier to each table
growth.summ2 <- growth.summ %>%
  mutate(Type = "Growth", .before = 1)

mortality.summ2 <- mortality.summ %>%
  mutate(Type = "Mortality", .before = 1)

# Combine the tables
grow.mort.summ <- bind_rows(growth.summ2, mortality.summ2)

# Create flextable
g.m.summ.ft <- flextable(grow.mort.summ) %>%
  merge_v(j = "Type") %>%  # Merge cells in Section column
  valign(j = "Type", valign = "top") %>%  # Top align section labels
  flextable::align(j = "Type", align = "left") %>%   # Left align section labels
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  set_table_properties(layout = "autofit", width = 0) %>%
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
    flextable::align(j = 3:7, align = "center", part = "all")

```

```{r set_captions}

# 
fig_nums <- captioner::captioner(prefix = "")
tab_nums <- captioner::captioner(prefix = "Table")

f.ref <- function(x) {
 caption_text <- fig_nums(x)
  # Remove any leading numbers and extract just "Figure X"
  stringr::str_extract(caption_text, "\\d+")
}

t.ref <- function(x)  {
  stringr::str_extract(tab_nums(x), "[^:]*")
}

tab.tbl_growth <- tab_nums(name = "tbl_growth", 
                        caption = "Decadal growth estimates by state and overall for analysis species (cm\u00B2/decade). Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots.")

tab.tbl_mort <- tab_nums(name = "tbl_mort", 
                        caption = "Decadal mortality estimates by state and overall for analysis species. Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots.")

fig.cwdmap_cap <- fig_nums(name = "fig_cwdmap", 
                        caption = "Fuzzed FIA plot locations displayed with initial visit climatic water deficit (CWD) values (A) and CWD change between visits (C) for the U.S. states, north to south, of Washington, Oregon, and California (see inset map B for the location of these states within the USA).")

fig.domains <- fig_nums(name = "fig_domains",
                        caption = "Decadal (A) growth and (B) mortality domain estimates for *Calocedrus decurrens*.  The six domains are defined as FIA plots that fall [A]bove or [B]elow the CWD change threshold with [L]ow, [M]edium, or [H]igh initial CWD values (see text for details). E.g., 'AL' is Above CWD change threshold, Low initial CWD value. The distribution of plot CWD change and initial values (C) is presented as well, with the blue horizontal line at a change in CWD of 5 mm representing the change threshold and the green vertical lines representing the quantile boundaries between Low, Medium, and High initial CWD values." )

fig.diff_grow <- fig_nums(name = "fig_diffgrow", 
                          caption = paste0("Differences in mean growth rates (cm\u00B2/decade) between domains by species.  Domain abbreviations are described in the legend for ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are coniferous; those below are deciduous."))

fig.diff_mort <- fig_nums(name = "fig_diffmort", 
                          caption = paste0("Differences in mean decadal mortality rates between domains by species.  Domain abbreviations are described in the legend for ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are coniferous; those below are deciduous."))

tab.summ_grow_mort <- tab_nums(name = "tbl_summ_grow_mort",
                               caption = paste0("Summary of number of comparisons with 95% CIs that do not overlap with zero ([T]otal, [P]ositive-, and [N]egative-valued estimates) for Figures ", fig_nums("fig_diffgrow", display = "num"), " and ", fig_nums("fig_diffmort", display = "num"), " by analysis (T)ype (Growth or Mortality) and type of (C)omparison.  The column Domains lists the domains compared; All is the sum of values within a Comparison. See text for other domain labels.  'n' is the total number of differences examined within a Comparison."))

```

# Abstract

# Introduction

# Methods

## FIA data

We used U.S. National Forest Inventory datasets for our analyses of forestland in California, Oregon, and Washington, USA.  Forestland is defined as an area of 4,050 m$^2$ that is at least 10 percent stocked or potentially stocked with tree species, excluding urban and agricultural land uses (Reams et al., 2005).  Additionally, as our study relied on plot revisit data, we only examined plots that were considered forestland during both visits.

FIA plots consist of four points where all trees with stem diameters greater than or equal to 12.7 cm (for most species measured at 1.37 m above the ground) were tallied in a 7.32 m radius circular subplot (total area 672.5 m$^2$) with additional trees sampled in an anulus around each subplot (macroplot) around each subplot, with a total radius of 18 m from the center point and a total area of 1012 m$^2$.  We did not include data from FIA microplots.  Bechtold and Scott (2005) provide plot design details and measurement protocols.

Since 2000, FIA plots have been selected via a spatially balanced probability sample (McRoberts 2005, Reams et al., 2005) with a target plot density of one ground plot every 24 km$^2$. Some regions and ownerships intensify the plot density; we excluded these additional plots to simplify stratification estimation, relying on FIA-defined strata and using the density of FIA Phase 1 sampling points within and across strata to determine strata weights.

Plots in the three states are sampled every 10 years with a target of 1/10 of plots sampled in a panel every year (Thompson 2015). Our analysis depends on examining the fate and growth of trees from two sequential visits (referred to as the first and second plot visit), ideally from a complete sample panel remeasure.  We obtained our FIA data from the USDA FIA DataMart web portal (https://research.fs.usda.gov/products/dataandtools/fia-datamart; downloaded May 5, 2025).  They included plot data sampled from 2001 to 2023.

Individual trees within plots were retained in the analysis if they were alive at the initial visit and met the stem diameter criterion during that visit (i.e., plot ingrowth was not included).  The growth analysis additionally restricted trees considered to those that were alive at both visits.  The mortality analysis excluded trees killed through silvicultural practices.  Both analyses excluded trees that shrank in diameter, physically moved due to landslides or other event, were erroneously tallied at either visit, were missed due to data collection procedural changes, or existed at any time in a nonsampled area.  We used only FIA plots sampled specifically for tree mortality and growth.

## Climate data

We obtained plot-level CWD data from ClimateNA (Wang et al. 2016) for June, July, and August (summer months) between 1980 and 2023 (Bryce: I just averaged the CWD values for these months.  It wasn’t available seasonally, was it? Also, can you provide more info on how CWD is calculated?  I know it is PET-AET, but I’m not sure how to interpret a monthly/seasonal value). Latitude and longitude were based on FIA-fuzzed coordinates (citation?  Bryce – should we tighten this up by running the analysis on non-fuzzed coordinates?).  We found the 20-year mean of summer CWD values for each plot prior to and including the first visit.  We refer this this value as the “initial CWD value”. We then found the mean of summer CWD values between visits and including the second visit year.  The difference between this second value and initial CWD value is the “CWD change value” (maybe ‘difference’ really is a better, less loaded term).  Both the initial CWD and CWD change values are mapped in Figure 1.

## Analysis
Estimation approach: Following Section 4.3.4 in Scott et al. (2005) we obtained point estimates of species growth and mortality using a ratio estimators (Eqn 4.16).  For mortality, the numerator is the sum of stratum-weighted means of trees per plot of a given species that died in plots.  The number of sampled trees in each plot is expanded by the area sampled (subplot or macroplot area) adjusted by the proportion of the plot area available for sampling and the number of years between visits.    The denominator is similar but is calculated as the sum of stratum-weighted means of all trees per plot of a given species.

The numerator of species-specific growth estimates is the sum of stratum-weighted means of the increase in basal area (cm$^2$) per plot while the denominator is the same as was used for the mortality estimate. This produces a point estimate for a species’ mean growth.  Mortality and growth values were standardized to a decadal period.

The analysis finds species growth and mortality estimates for six domains based on CWD values (Figure 2C).   The initial CWD values are divided into three quantiles: values below the first quartile (Low), between the first and third quartiles (Medium), and above the third quartile (High).  These initial CWD value divisions allowed for species-specific assessments.  For all species we defined an arbitrary boundary of 5 mm to define plots with a greater change in CWD (Above [the boundary]) and those with less CWD change (Below).  This was done because species did not differ greatly in CWD change (Table S2).  The category Below includes CWD changes that are negative (i.e., mean CWD was larger prior to the initial visit than between visits).  We also report species mortality and growth estimates for all plots across and within Oregon, Washington, and California irrespective of CWD values.  We limited the number of species examined to those with more than 400 plots.  We report domain estimates for all CWD-based domains with 10 or more plots.

We obtained empirical means and 95% confidence intervals via bootstrap. For every species we sampled plots with replacement and obtained point estimates.  For each domain we calculated a mean and upper and lower 95% CIs with the 2.5% and 97.5% quantiles from 1000 sampling iterations.  This approach allows for nonsymmetrical CIs around a mean and prevents CIs from including negative values.  We estimated differences between pairs of CWD domains by subtracting within-iteration domain point estimates.

We

[permutation analysis]

# Results

Our study tracked the fate of `r comma(analysis.stats$tree.num)` trees from `r length(sel.spp)` species in `r comma(analysis.stats$plot.num)` plots that were visited twice by FIA crews. The two visits for `r pct.remeas[2]`% of plots were 10 years apart (`r pct.remeas[1]`% were 9 years, `r pct.remeas[3]`% were 11 years). Plots were distributed in `r analysis.stats$strat.num` strata.  Of the `r length(sel.spp)` species, `r analysis.stats$n.gym.spp` are gymnosperms.  Mean overall growth rates for species varied between `r min.grow.spp$mean` (*`r min.grow.spp$spp`*) and `r max.grow.spp$mean` (*`r max.grow.spp$spp`*, `r t.ref("tbl_growth")`) cm$^2$/decade.  Decadal mean mortality rates varied between `r min.mort.spp$mean` (*`r min.mort.spp$spp`*) and `r max.mort.spp$mean` (*`r max.mort.spp$spp`*, `r t.ref("tbl_mort")`). Growth and mortality rates differed among states (Tables `r tab_nums("tbl_growth", display = "num")` and `r tab_nums("tbl_mort", display = "num")`).

Forested plot 10-year differences in CWD do not visually coincide with initial CWD values (Figure `r fig_nums("fig_cwdmap", display = "num")`); however, `r analysis.stats$cwd.gt0.pct`% of plots increased in their CWD values to some degree during that time period. Tree species exhibited a variety of initial-visit CWD ranges, with means ranging from `r min.mean.init` (*`r min.mean.init.spp`*) to `r max.mean.init` (*`r max.mean.init.spp`*; Table S1).  The CWD difference tended to be comparatively small, with a maximum 95th percentile value of `r max.q95.diff` (*`r max.q95.diff.spp`*, Table S2). The distribution of individual species plot domain values for CWD and change in CWD can be explored on the dashboard (https://tinyurl.com/FIAgrowmort).

The confidence intervals for 10-year mortality and growth rate domains did not necessarily follow the same patterns nor did all confidence intervals for a species' mortality or growth estimates overlap (e.g., Figures `r fig_nums("fig_domains", display = "num")`A and `r fig_nums("fig_domains", display = "num")`B). The relationship between initial CWD values and plot-level change in CWD were not reliably linear (e.g., Figure `r fig_nums("fig_domains", display = "num")`C) and varied by species (see dashboard).

We examined differences in 10-year growth estimates (Figure `r fig_nums("fig_diffgrow", display = "num")`) and mortality estimates (Figure `r fig_nums("fig_diffmort", display = "num")`) among domains.  The column "Above vs. Below Threshold" compares the domains above the CWD change threshold with those below it (e.g., for *Abies lasiocarpa*, subtracting the estimates for columns BL, BM, and BH from AL, AM, AH, respectively; Figure `r fig_nums("fig_domains", display = "num")`A). Negative estimates for the "Above vs. Below Threshold" in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")`  indicate that the 10-year growth or mortality rate was greater for a species in plots below the change estimates, respectively.  `r t.ref("tbl_summ_grow_mort")` summarizes the number of domain differences in each of the columns in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")` where the 95% CI does not include zero as well as the number of comparisons per figure column.

Growth domain comparisons (`r fig_nums("fig_diffgrow")`, `r t.ref("tbl_summ_grow_mort")`) did not appear to be produce substantially more "significant" negative vs. positive estimates, and all three columns of `r fig_nums("fig_diffgrow")` had similar numbers of "significant" comparisons. Growth may have been susceptible (negatively or positively) to a change in CWD and differed by initial CWD values, and the responses are species-specific. *Abies amabilis*, *Abies grandis*, *Calocedrus decurrens*, *Pinus ponderosa*, *Pseudotsuga menziesii*, *Tsuga mertensiana*, *Acer macrophyllum*, and *Alnus rubra* exhibited greater growth in the Low domains than the Medium and/or High domains.  Except for *Alnus rubra*, these results predominantly occurred in both the Below domain (2) or the Above and Below domains (5).  *Thuja plicata* had higher growth in the Above and Below Medium domains than either High or Low domains. *Abies concolor*, *Larix occidentalis*, *Picea engelmannii*, *Pinus contorta*, *Pinus lambertiana*, *Arbutus menziesii*, and *Quercus kelloggii* had greater growth in the more water stressed High domains than the Medium and/or Low domains. Except for *Larix occidentalis*, these results predominantly occurred in both the Above domain (3) or the Above and Below domains (3). Other species exhibited no 'significant' findings or displayed a mixture of results.

Mortality domain comparisons are similar in certain regards to the growth comparisons: results for the three types of domain comparisons did not appear to be predominantly positive or negative (`r fig_nums("fig_diffmort")`, `r t.ref("tbl_summ_grow_mort")`).  There were fewer mortality estimates with "significant" findings than growth. There were fewer differences for "Above vs. Below Threshold" comparisons (`r t.ref("tbl_summ_grow_mort")`) than for the comparisons Above or Below the threshold, indicating that mortality may not be as strongly associated with the change in CWD as by the starting value.  *Larix occidentalis*, *Picea engelmannii*,	*Pinus contorta*,	*Pinus monticola*, *Thuja plicata*,	and *Arbutus menziesii* had greater mortality in the lower water deficit Low domains than Medium and/or High domains. These results occurred in Above (3) or both Above and Below (3) domains. Mortality rates for *Abies lasiocarpa*,	*Calocedrus decurrens*,	*Pinus lambertiana*, *Pinus ponderosa*,	*Pseudotsuga menziesii*,	*Tsuga heterophylla*,	*Tsuga mertensiana*,	*Acer macrophyllum*,	and *Notolithocarpus densiflorus* were greater in the High domains than the Medium and/or Low domains.  These species were approximately evenly split between differences being present in the Above (4) or Below (5) domains, but none with 'significant' findings in both the Above and Below domains.

Mortality: 
Picea engelmannii: Highest upper mortality = low. 
Abies lasiocarpa: Lowest mortaity in Low. 
- delta CWD lower for PIEN than for PILA.  Mean of High Above = 7.5 vs. 9ish.  
- Similar patterns of delta CWD vs. CWD
- ABLA = Least mortality in Low, in OLY penninsula, NW Cascades of WA.  High mortality overall (0.22)
- PIEN = most mort in NE WA. High mortality overall (0.28)
- PILA = most mort in mid/southern Sierras.  Seems like high mortality (0.20)

ABLA and PIEN are conspecifics / codominants. Perret et al. 2023 estimated rangewide mortality of ABLA of 25% +/-2.7, mort concentrated in eastern OR cascades (and ID and CO). 83% of mort due to fire or bio disturbance.  Conspecifics mostly affected by beetles, ABLA by fire.  PIEN = higher survival and longevity (used to be; they estimate around a high of 1% annual mortality for fir, below 0.5% annually in spruce, PIEN), ABLA = faster growing, more prolific regeneration (Andrus 2018).  We found that PIEN had higher mortality in the Above Low and Medium domains.  We note that the Above domains represents a proportionally large increase for Low and Medium, with CWD means of approximately 50 and 80 mm, respectively.  The same is true for ABLA, however, where the mean for Low is approximately 40 mm yet PIEN has less mortality in the Low domains and PIEN has more mortality in the Above Low domain.  The difference may be geographical: ABLA low-mortality plots are primarily on the Olympic Peninsula of Washington and the northwestern Cascades, where rainfall is abundant.  PIEN is generally absent from these locations.  However, PIEN has the least mortality in the Above domain in the Blue Mountains of Oregon and Washington while ABLA does not show a similar finding for the same geographical region (their AH domains largely overlap and exist only in that mountain range). PIEN mortality in N&NE WA may have been driven by spruce beetles (WADNR 2023)

Pinus contorta: Greatest mort in Low, least in High
Pinus ponderosa: Greatest mort in High
- In southern Sierra I see that they co-occur but have opposite mortality signals. Also a scattering in the Blue Mtns of opposite signals domains. Different fire strategies; Keely et al. 2011 may be useful.

Growth: 
PIEN: greater growth in High, above and below.  So, greater growth where least mortality.  
ABLA: not really any difference in growth by domain.  Supposed to be faster growing than PIEN, but not seeing it.

Lowest growth rate is for PIEN BL, at 15 in2.  Highest growth for ABLA is 12.5 in2. ABLA is supposed to be a faster grower (height) but maybe not DIA?  Diameter growth may be slow (trees 10-20" may be 150-200 yrs; Alexander et al. 1990).  On other hand, PIEN appears to have the capability of achieving greater diameters relatively sooner than ABLA (Parish et al. 1999).  PIEN growth in the High domains may be related to more growing days relative to other domains.

Communities:
Pinus lambertiana PILA
Pinus monticola PIMO
(Pinus ponderosa) PIPO
(Pinus contorta) PICO

These two are interesting because the lowest mortality domain for PIMO (Above, High) coincides geographically in the southern Sierra with the highest mortality domain for PILA (and for much of the highest-mortality domain for PIPO).  PILA has the highest growth in AH whereas there isn't much of a growth story for PIMA.  PIPO and PICO have great range overlap.  PICO has the highest mortality in Low (NE WA), particularly AL and lowest mortality in BH (Sierras, OR Cascades).  Note that the mortality rate for PICO in WA was the highest of any species in any state, and more than twice as high as the mortality rate for WA and 5 times higher than CA.  PICO growth is lowest in AL.  PIPO has the highest growth in Low.  PIPO in NE WA has much less acreage of pine beetle damage in 2023 than PICO, not really a factor for PIMO.

What is happening in the southern Sierras?  N/NE wA?

Pseudotsuga menziesii PSME
So much data.  Over 4500 plots.  Except for High, where mortality of Above and Below was approximately equivalent, we see that mortality was high in AL and AM.  BL, BM had the lowest mortality.  These coincided with moister plots.  Growth was highest for Below except for BH.  PSME mortality appears to be drought related - see Cline et al. 2025. 
-Tree mortality increased 2020-2023, prolongued drought and warming.  Soil moisture, temperature, and VPD key predictors of old tree (125-500yr) mortality. 
-strongest signal at mid-elevation west of Cascade crest, OR
-Strongest for suppressed/intermediate, non-dominant or codominant individuals, smaller diameter (competition exacerbated the effect of climate variables on smaller-tree mortality at drier sites). 
-insect mortality more prevalent following drought. 
- lots of relevant citations
- Western hemlock (Tsuga heterophylla) in the S/I cohort also died more frequently.  ACMA also did at one site.  
- PNW heat dome of 2021 caused 11% mortality of C/D trees at Falls Creek site

Acer macrophyllum ACMA
We found that mortality was higher for BH than BL, BM. Maybe more interesting, growth was higher for Low, then Medium, then High irrespective of Above, Below. We cannot explain outright the high mortality for ACMA in California.  However, Betzen et al. (2021) found widespread evidence in Washington of ACMA decline (crown dieback, thinning, discoloration, reduced leaf size) and reduced growth and mortality associated with higher temperatures and water limitations. These findings are technically in alignment with the overall pattern in growth and mortality we observed, but does not specifically address the very high mortality rate of CA relative to OR and WA.

# Discussion

The initial CWD values coincide fairly well with mean annual precipitation [NOAA Climate.gov](https://www.climate.gov/news-features/featured-images/new-maps-annual-average-temperature-and-precipitation-us-climate).

# Conclusion

# Acknowledgments

# References

::: {#refs}

Betzen, J.J., Ramsey, A., Omdal, D., Ettl, G.J. and Tobin, P.C., 2021. Bigleaf maple, Acer macrophyllum Pursh, decline in western Washington, USA. Forest Ecology and Management, 501, p.119681.

Wang, T., Hamann, A. Spittlehouse, D.L. and Carroll, C. 2016. Locally downscaled and spatially customizable climate data for historical and future periods for North America. PLoS One 11: e0156720

Alexander, R.R., Shearer, R.C. and Shepperd, W.D., 1990. Abies lasiocarpa (Hook.) Nutt. subalpine fir. Silvics of North America, 1, pp.60-70.

Washington State Department of Natural Resources (DNR) Forest Reliance Division. 2024. Forest health highlights in Washington: A summary of insect, disease, and other disturbance conditions affecting Washington’s forests.

Washington State Department of Natural Resources (WADNR). 2024. Forest Health Highlights in Washington – 2024. Washington State Department of Natural Resources, Forest Resilience Division, Olympia, WA. https://dnr.wa.gov/sites/default/files/2025-03/rp_2023_fhh_report_final_wr.pdf, accessed 9/18/2025

:::

<!---BLOCK_LANDSCAPE_START--->

```{r tbl-growth, tab.cap=tab.tbl_growth}

growth.ft

```

```{r tbl-mort, tab.cap = tab.tbl_mort}

mort.ft

```

<!---BLOCK_LANDSCAPE_STOP--->

```{r mapfig, fig.height = 7,  fig.dpi=300, fig.cap=fig.cwdmap_cap, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "cwd_maps_figure.png"))

```

```{r domainsfig, fig.width = 6,  fig.dpi=300, fig.cap=fig.domains, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "Example_spp_figure.png"))

```

```{r diff_grow_fig, fig.height = 6, width = 6, fig.dpi=300, fig.cap=fig.diff_grow, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Growth_figs_cwd/Growth_Panel_Plot.png"))

```

```{r diff_mort_fig, fig.height = 6, width = 6, fig.dpi=300, fig.cap=fig.diff_mort, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Mort_figs_cwd/Mort_Panel_Plot.png"))

```

\newpage

```{r tbl-summ-grow-mort, tab.cap = tab.summ_grow_mort}

g.m.summ.ft

```