### 2025-10-06 Supplemental Material.Rmd
The permutation analysis may be in its final form.  The figures 3 and 4 were altered to show MC permutation p-value symbols, and tables were added to the supplemental material to display the actual permuatation p-values.

---
title: "Supplemental material for Tree growth, mortality, and climatic stress in west-coast states, USA"
date: "`r Sys.Date()`"
author: "Jeremiah D. Groom"
output: 
  officedown::rdocx_document:
    mapstyles:
      Normal: ['First Paragraph']
    reference_docx: "template_use2.docx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(officedown)
library(officer)
library(kableExtra)   # Enhanced table formatting
library(scales)       # Scale functions for visualization
library(flextable)    # Professional tables for Word output
library(rvg)
library(captioner) # For figure captions.

fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)

set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  theme_fun = "theme_vanilla"
)

```

```{r load_data}

source('Global.R')
source(paste0(CODE.LOC, "Functions.R"))
# Some general analysis info + plots
analysis.stats <- read_rds(paste0(RESULTS.OTHER, "analysis.stats.RDS"))

```

```{r cwd_summary_tables}
cwd.init <- analysis.stats$init.cwd.stats 
cwd.diff <- analysis.stats$diff.cwd.stats

init.ft <- flextable(cwd.init) %>%
  fontsize(size = 10, part = "body") %>%
  fontsize(size = 11, part = "header") %>%
  italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  set_table_properties(layout = "autofit", width = 0) %>%
  border_remove() %>%
  hline_top(part = "all") %>%
  hline_bottom(part = "all")

diff.ft <- flextable(cwd.diff) %>%
  fontsize(size = 10, part = "body") %>%
  fontsize(size = 11, part = "header") %>%
  italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  set_table_properties(layout = "autofit", width = 0) %>%
  border_remove() %>%
  hline_top(part = "all") %>%
  hline_bottom(part = "all")

```

```{r domain_pvals}

psig.dat2 <- analysis.stats$psig.dat2
psig.g <- psig.dat2[, 1:7]
psig.m <- psig.dat2[, c(1, 8:13)]

psig.ft.g <- psig.ft.fcn(psig.g)
psig.ft.m <- psig.ft.fcn(psig.m)

```

```{r table-legends}
tab_nums <- captioner::captioner(prefix = "Table S")

t.ref <- function(x)  {
  stringr::str_extract(tab_nums(x), "[^:]*")
}

tbl_cwd_init <- tab_nums(name = "tbl_cwdinit", 
                         caption = "Summary statistics for initial-visit plot CWD values by species. The number of species' plots (n), minimum, mean, median, and maximum CWD values are provided, along with the 5%, 25%, 75%, and 95% quantiles (q).")

tbl_cwd_diff <- tab_nums(name = "tbl_cwddiff", 
                         caption = "Summary statistics for the difference in plot CWD values between visits by species. The number of species' plots (n), minimum, mean, median, and maximum CWD values are provided, along with the 5%, 25%, 75%, and 95% quantiles (q).")

tbl_psig_growth <- tab_nums(name = "tbl_psiggrowth", 
                            caption = "Monte Carlo permutation p-values for tree species growth domains. Values correspond to Manuscript Figure 3 values.  Values of zero represented no permutation results with confidence intervals equal to or greater than those observed in the data.")

tbl_psig_mort <- tab_nums(name = "tbl_psigmort", 
                            caption = "Monte Carlo permutation p-values for tree species mortality domains. Values correspond to Manuscript Figure 4 values. Values of zero represented no permutation results with confidence intervals equal to or greater than those observed in the data.")

```

\newpage

<!---BLOCK_LANDSCAPE_START--->

```{r table-cwd-init, tab.cap=tbl_cwd_init}

init.ft

```

\newpage

```{r table-cwd-diff, tab.cap=tbl_cwd_diff}

diff.ft

```

\newpage

```{r table-psig-growth, tab.cap=tbl_psig_growth}

psig.ft.g

```

\newpage

```{r table-psig-mortality, tab.cap=tbl_psig_mort}

psig.ft.m

```

<!---BLOCK_LANDSCAPE_STOP--->