### 2025-10-13 Manuscript.Rmd
Updates to MS.  Added current Intro and Discussion material, cleaned up figure legends for Figs 3 & 4

---
title: "Tree growth, mortality, and climatic stress in west-coast states, USA"
author:
  - name: "Jeremiah D. Groom"
    affiliations:
      - name: "Groom Analytics LLC"
        address: "1975 SE Crystal Lake Dr., Unit 173"
        city: "Corvallis"
        state: "OR"
        postal-code: "97333"
    orcid: "0000-0002-0110-1036"
    email: "jeremy@groomanalytics.com"
  - name: "Bryce Frank"
    affiliations:
      - name: "USDA Forest Service"
        department: "Pacific Northwest Field Station"
        address: "3625 93rd Ave SW"
        city: "Olympia"
        state: "WA"
        postal-code: "98512"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  officedown::rdocx_document:
    reference_docx: "template_use2.docx"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(kableExtra)   # Enhanced table formatting
library(scales)       # Scale functions for visualization
library(flextable)    # Professional tables for Word output
library(officedown)
library(officer)
library(rvg)
#devtools::install_github("adletaw/captioner")
library(captioner) # For figure captions.

```

```{r other_settings}
fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)
```

```{r constants}

source("Global.R")
source(paste0(CODE.LOC, "Functions.R"))

# Define any custom functions
format_p_value <- function(p) {
  case_when(
    p < 0.001 ~ "< 0.001",
    p < 0.01 ~ sprintf("= %.3f", p),
    TRUE ~ sprintf("= %.2f", p)
  )
}

set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  theme_fun = "theme_vanilla"
)

centeredP <- fp_par(text.align = "center")

```

```{r load-data}
# Some general analysis info + plots
analysis.stats <- read_rds(paste0(RESULTS.OTHER, "analysis.stats.RDS"))

# Percent of plots sampled with REMPER = 9, 10, and 11 years (rounded)
pct.remeas <- as.vector(analysis.stats$pct.remeas)

# Species scientific names
spp.names.use <- analysis.stats$gym.angio.dat

# Growth data for species overall, states, differences
grow.state <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/State_Growth_", CLIM.VAR.USE, ".csv"))
grow.all <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/All_Growth_", CLIM.VAR.USE, ".csv"))
grow.diff <- read_rds(paste0(save.loc.fcn(1), "Processed_6Domain_Data.RDS"))

# Mortality data for species overall and states
mort.state <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/State_Mortality_", CLIM.VAR.USE, ".csv"))
mort.all <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/All_Mortality_", CLIM.VAR.USE, ".csv"))
mort.diff <- read_rds(paste0(save.loc.fcn(2), "Processed_6Domain_Data.RDS"))

# Significant findings, Growth
grow.n.AvB <- nrow(grow.diff$AvB2)
grow.n.A <- nrow(grow.diff$A_LMH2)
grow.n.B <- nrow(grow.diff$B_LMH2)

grow.AvB.sig <- length(grow.diff$AvB2$significant[grow.diff$AvB2$significant == TRUE])
grow.AvB.signeg <- length(grow.diff$AvB2$significant[grow.diff$AvB2$significant == TRUE & grow.diff$AvB2$Means < 0])
grow.A.sig <- length(grow.diff$A_LMH2$significant[grow.diff$A_LMH2$significant == TRUE])
grow.A.signeg <- length(grow.diff$A_LMH2$significant[grow.diff$A_LMH2$significant == TRUE & grow.diff$A_LMH2$Means < 0])
grow.B.sig <- length(grow.diff$B_LMH2$significant[grow.diff$B_LMH2$significant == TRUE])
grow.B.signeg <- length(grow.diff$B_LMH2$significant[grow.diff$B_LMH2$significant == TRUE & grow.diff$B_LMH2$Means < 0])

# Significant findings, Mort
mort.n.AvB <- nrow(mort.diff$AvB2)
mort.n.A <- nrow(mort.diff$A_LMH2)
mort.n.B <- nrow(mort.diff$B_LMH2)

mort.AvB.sig <- length(mort.diff$AvB2$significant[mort.diff$AvB2$significant == TRUE])
mort.AvB.signeg <- length(mort.diff$AvB2$significant[mort.diff$AvB2$significant == TRUE & mort.diff$AvB2$Means < 0])
mort.A.sig <- length(mort.diff$A_LMH2$significant[mort.diff$A_LMH2$significant == TRUE])
mort.A.signeg <- length(mort.diff$A_LMH2$significant[mort.diff$A_LMH2$significant == TRUE & mort.diff$A_LMH2$Means < 0])
mort.B.sig <- length(mort.diff$B_LMH2$significant[mort.diff$B_LMH2$significant == TRUE])
mort.B.signeg <- length(mort.diff$B_LMH2$significant[mort.diff$B_LMH2$significant == TRUE & mort.diff$B_LMH2$Means < 0])

# Summary of significant Growth and Mortality findings for table creation.
datasets.grow <- list(grow.diff$AvB2, grow.diff$A_LMH2, grow.diff$B_LMH2)
datasets.mort <- list(mort.diff$AvB2, mort.diff$A_LMH2, mort.diff$B_LMH2)
datatypes <- list("Above vs. Below", "Above", "Below")
n.comp.grow <- list(grow.n.AvB, grow.n.A, grow.n.B)
n.comp.mort <- list(mort.n.AvB, mort.n.A, mort.n.B)

growth.summ <- pmap(list(datasets.grow, datatypes, n.comp.grow), summ.fig.fcn) %>% bind_rows()
mortality.summ <- pmap(list(datasets.mort, datatypes, n.comp.mort), summ.fig.fcn) %>% bind_rows()

```

```{r table-creation}
# State growth
grow.state.cm <- cm2.fcn(1, grow.state)
grow.all.cm <- cm2.fcn(1, grow.all)
s.g.dat <- state.table.fcn(grow.state.cm, grow.all.cm, 1)

growth.ft <- flextable(s.g.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
    italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = TRUE)
min.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = TRUE)

# State mortality
s.m.dat <- state.table.fcn(mort.state, mort.all, 2)

mort.ft <- flextable(s.m.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",  
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = FALSE, digits = 2)
min.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = FALSE, digits = 2)

# CWD amounts
min.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.min.mean]
min.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.min.mean]
max.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.max.mean]
max.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.max.mean]

max.q95.diff <- analysis.stats$diff.cwd.stats$q95[analysis.stats$diff.cwd.95max]
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]
max.range.diff <- analysis.stats$diff.cwd.90range.max
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]

# Adding a section identifier to each table
growth.summ2 <- growth.summ %>%
  mutate(Type = "Growth", .before = 1)

mortality.summ2 <- mortality.summ %>%
  mutate(Type = "Mortality", .before = 1)

# Combine the tables
grow.mort.summ <- bind_rows(growth.summ2, mortality.summ2)

# Create flextable
g.m.summ.ft <- flextable(grow.mort.summ) %>%
  merge_v(j = "Type") %>%  # Merge cells in Section column
  valign(j = "Type", valign = "top") %>%  # Top align section labels
  flextable::align(j = "Type", align = "left") %>%   # Left align section labels
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  set_table_properties(layout = "autofit", width = 0) %>%
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
    flextable::align(j = 3:7, align = "center", part = "all")

```

```{r set_captions}

# 
fig_nums <- captioner::captioner(prefix = "")
tab_nums <- captioner::captioner(prefix = "Table")

f.ref <- function(x) {
 caption_text <- fig_nums(x)
  # Remove any leading numbers and extract just "Figure X"
  stringr::str_extract(caption_text, "\\d+")
}

t.ref <- function(x)  {
  stringr::str_extract(tab_nums(x), "[^:]*")
}

tab.tbl_growth <- tab_nums(name = "tbl_growth", 
                        caption = "Decadal growth estimates by state and overall for analysis species (cm\u00B2/decade). Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots.")

tab.tbl_mort <- tab_nums(name = "tbl_mort", 
                        caption = "Decadal mortality estimates by state and overall for analysis species. Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots.")

fig.cwdmap_cap <- fig_nums(name = "fig_cwdmap", 
                        caption = "Fuzzed FIA plot locations displayed with initial visit climatic water deficit (CWD) values (A) and CWD change between visits (C) for the U.S. states, north to south, of Washington, Oregon, and California (see inset map B for the location of these states within the USA).")

fig.domains <- fig_nums(name = "fig_domains",
                        caption = "Decadal (A) growth and (B) mortality domain estimates for *Calocedrus decurrens*.  The six domains are defined as FIA plots that fall [A]bove or [B]elow the CWD change threshold with [L]ow, [M]edium, or [H]igh initial CWD values (see text for details). E.g., 'AL' is Above CWD change threshold, Low initial CWD value. The distribution of plot CWD change and initial values (C) is presented as well, with the blue horizontal line at a change in CWD of 5 mm representing the change threshold and the green vertical lines representing the quantile boundaries between Low, Medium, and High initial CWD values." )

fig.diff_grow <- fig_nums(name = "fig_diffgrow", 
                          caption = paste0("Differences in mean growth rates (cm\u00B2/decade) between domains by species.  Domain abbreviations are described in the legend for ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are coniferous; those below are deciduous. Empirical MC permutation p-values are provided for filled-circle contrasts, with \u2013\u2013 corresponding to p-values of > 0.05, while one to four asterisks correspond to p > 0.01, > 0.001, > 0.0001 and < 0.0001, respectively."))

fig.diff_mort <- fig_nums(name = "fig_diffmort", 
                          caption = paste0("Differences in mean decadal mortality rates between domains by species.  Domain abbreviations are described in the legend for ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are coniferous; those below are deciduous.  Empirical MC permutation p-values are provided for filled-circle contrasts, with \u2013\u2013 corresponding to p-values of > 0.05 while one to four asterisks correspond to p > 0.01, > 0.001, > 0.0001 and < 0.0001, respectively. "))

tab.summ_grow_mort <- tab_nums(name = "tbl_summ_grow_mort",
                               caption = paste0("Summary of number of comparisons with 95% CIs that do not overlap with zero ([T]otal, [P]ositive-, and [N]egative-valued estimates) for Figures ", fig_nums("fig_diffgrow", display = "num"), " and ", fig_nums("fig_diffmort", display = "num"), " by analysis (T)ype (Growth or Mortality) and type of (C)omparison.  The column Domains lists the domains compared; All is the sum of values within a Comparison. See text for other domain labels.  'n' is the total number of differences examined within a Comparison."))

```

# Abstract

# Introduction

There  are signs that tree species in western North America are in regional decline.  Johnston et al. (2025) document a 29% 10-year mortality rate of 300+ yr old trees in Eastern Oregon.  High-severity fires in California in 2020 and 2021 killed 13%-19% of giant sequoia (Sequoiadendron gianteum) with >1.07-m bole diameters (Stephenson and Brigham 2021, Shive et al. 2021); recruitment in these burns was depressed (Soderberg et al. 2022).   Five of eight common species in western North America exhibit declines in relative density  as measured by relative stand density (Stanke et al. 2021).  Shriver et al. (2022) found that four of five juniper and piñon pine species were declining in portions of their range, comparing mortality to recruitment rates.

Climate change as manifest through increases or intensification of drought and aridity can affect tree mortality directly through drought and indirectly by increasing wildfire frequency and severity and promoting insect and disease outbreak.  Hammond et al. (2022) found globally that the intensity of drought and heat events are directly associated with widespread forest die-off events.  Wildfire frequency and area burned in the western United States has increased, in concert with the longer fire seasons, hotter summer temperatures, and earlier snowmelt increasingly experienced in the western United States (Abatzoglou and Williams 2016, Westerling 2016, Juang et al. 2022).  There has also been a regional increase in the area of bark beetle outbreaks (Meddens et al. 2012, Hicke et al. 2016).  Many of these climate-influence mortality causes may be attributable to the physical condition of trees following abnormal or prolonged exposure to water stress.

Drought-induced mortality can be driven by carbon starvation and hydraulic failure (see reviews by Adams et al. 2010 and McDowell et al. 2008).  Carbon starvation occurs when isohydric species (e.g., the piñon pine group), or those which mainly rely on stomatal closure to avoid excessive water loss through evapotranspiration (ET) receive insufficient atmospheric carbon dioxide for photosynthesis.  Anisohydric plants, or those who do not strongly regulate water loss through stomatal closure, risk hydraulic failure if sufficient xylem cavitation occurs due to insufficient water uptake needed to meet ET demands (e.g., as seen in Populus tremuloides; Anderegg et al. 2015). Adams et al. (2017) performed a review of drought-induced mortality across 23 species and concluded that hydraulic failure occurred broadly across taxa while non-structural carbohydrate reserves at mortality differed across taxa.  Regardless of mechanism, drought (specifically, water deficit) appears to be an important direct driver of mortality for tree taxa (Hammond et al. 2022. Repetitive citation. Others?).

Drought and physiological stress from drought can indirectly increase mortality as well.  Drought stress can increase tree susceptibility to wildfire, insects, and disease.  Wildfire fire area (Abatzoglou and Williams, 2016; Parks et al. 2025) and number (Westerling 2016) are positively associated with increasing aridity in the western US.   Bark beetle outbreaks in semi-arid areas of the western US may be positively affected by or associated with acute intense drought and water-stressed trees (Negrón 2009, Gaylord et al. 2015) but chronically stressed trees may not provide like bark beetle populations with the resources sufficient to promote population expansion (Huberty and Denno, 2004).  Norway spruce (Picea abies) saplings inoculated with the necrotrophic pathogenic fungi Heterobasidion annosum sensu stricto and exposed to drought conditions were more likely to die than control saplings or saplings exposed to either drought conditions or inoculation, although Scots pine (Pinus sylvestris) seedlings did not show a similar relationship (Gomez-Gallego et al. 2022). Other insects and disease may be promoted or inhibited by drought, depending on the circumstances and due to conflicting factors (Kolb et al. 2016).

# Methods

## FIA data

We used U.S. National Forest Inventory datasets for our analyses of forestland in California, Oregon, and Washington, USA.  Forestland is defined as an area of 4,050 m$^2$ that is at least 10 percent stocked or potentially stocked with tree species, excluding urban and agricultural land uses (Reams et al., 2005).  Additionally, as our study relied on plot revisit data, we only examined plots that were considered forestland during both visits.

FIA plots consist of four points where all trees with stem diameters greater than or equal to 12.7 cm (for most species measured at 1.37 m above the ground) were tallied in a 7.32 m radius circular subplot (total area 672.5 m$^2$) with additional trees sampled in an anulus around each subplot (macroplot) around each subplot, with a total radius of 18 m from the center point and a total area of 1012 m$^2$.  We did not include data from FIA microplots.  Bechtold and Scott (2005) provide plot design details and measurement protocols.

Since 2000, FIA plots have been selected via a spatially balanced probability sample (McRoberts 2005, Reams et al., 2005) with a target plot density of one ground plot every 24 km$^2$. Some regions and ownerships intensify the plot density; we excluded these additional plots to simplify stratification estimation, relying on FIA-defined strata and using the density of FIA Phase 1 sampling points within and across strata to determine strata weights.

Plots in the three states are sampled every 10 years with a target of 1/10 of plots sampled in a panel every year (Thompson 2015). Our analysis depends on examining the fate and growth of trees from two sequential visits (referred to as the first and second plot visit), ideally from a complete sample panel remeasure.  We obtained our FIA data from the USDA FIA DataMart web portal (https://research.fs.usda.gov/products/dataandtools/fia-datamart; downloaded May 5, 2025).  They included plot data sampled from 2001 to 2023.

Individual trees within plots were retained in the analysis if they were alive at the initial visit and met the stem diameter criterion during that visit (i.e., plot ingrowth was not included).  The growth analysis additionally restricted trees considered to those that were alive at both visits.  The mortality analysis excluded trees killed through silvicultural practices.  Both analyses excluded trees that shrank in diameter, physically moved due to landslides or other event, were erroneously tallied at either visit, were missed due to data collection procedural changes, or existed at any time in a nonsampled area.  We used only FIA plots sampled specifically for tree mortality and growth.

## Climate data

We obtained plot-level CWD data from ClimateNA (Wang et al. 2016) for June, July, and August (summer months) between 1980 and 2023 (Bryce: I just averaged the CWD values for these months.  It wasn’t available seasonally, was it? Also, can you provide more info on how CWD is calculated?  I know it is PET-AET, but I’m not sure how to interpret a monthly/seasonal value). Latitude and longitude were based on FIA-fuzzed coordinates (citation?  Bryce – should we tighten this up by running the analysis on non-fuzzed coordinates?).  We found the 20-year mean of summer CWD values for each plot prior to and including the first visit.  We refer this this value as the “initial CWD value”. We then found the mean of summer CWD values between visits and including the second visit year.  The difference between this second value and initial CWD value is the “CWD change value” (maybe ‘difference’ really is a better, less loaded term).  Both the initial CWD and CWD change values are mapped in Figure 1.

## Analysis
Estimation approach: Following Section 4.3.4 in Scott et al. (2005) we obtained point estimates of species growth and mortality using a ratio estimators (Eqn 4.16).  For mortality, the numerator is the sum of stratum-weighted means of trees per plot of a given species that died in plots.  The number of sampled trees in each plot is expanded by the area sampled (subplot or macroplot area) adjusted by the proportion of the plot area available for sampling and the number of years between visits.    The denominator is similar but is calculated as the sum of stratum-weighted means of all trees per plot of a given species.

The numerator of species-specific growth estimates is the sum of stratum-weighted means of the increase in basal area (cm$^2$) per plot while the denominator is the same as was used for the mortality estimate. This produces a point estimate for a species’ mean growth.  Mortality and growth values were standardized to a decadal period.

The analysis finds species growth and mortality estimates for six domains based on CWD values (Figure 2C).   The initial CWD values are divided into three quantiles: values below the first quartile (Low), between the first and third quartiles (Medium), and above the third quartile (High).  These initial CWD value divisions allowed for species-specific assessments.  For all species we defined an arbitrary boundary of 5 mm to define plots with a greater change in CWD (Above [the boundary]) and those with less CWD change (Below).  This was done because species did not differ greatly in CWD change (Table S2).  The category Below includes CWD changes that are negative (i.e., mean CWD was larger prior to the initial visit than between visits).  We also report species mortality and growth estimates for all plots across and within Oregon, Washington, and California irrespective of CWD values.  We limited the number of species examined to those with more than 400 plots.  We report domain estimates for all CWD-based domains with 10 or more plots.

We obtained empirical means and 95% confidence intervals via bootstrap. For each of the 1000 bootstrap iterations we sampled FIA plots with replacement and obtained point estimates for each species' domain.  For each domain we calculated a mean and upper and lower 95% CIs with the 2.5% and 97.5% quantiles from the  sampling iterations.  This approach allows for nonsymmetrical CIs around means.  We estimated contrasts between pairs of CWD domains by subtracting within-iteration domain point estimates.

Our analysis approach results in many comparisons.  We conducted a Monte Carlo (MC) permutation analysis to provide species-level false positive rates (p-values). The MC permutation analysis had 1000 iterations; each iteration relied on the bootstrap confidence interval estimation described above. The domain assignments for each species was sampled without replacement.  The distance of MC permutation contrast CI bounds from zero were compared against test CI bounds. The MC permutation procedure recorded a null significant value if the absolute values of the permutation iteration CI bounds were at least as far from zero as observed bounds.

# Results

Our study tracked the fate of `r comma(analysis.stats$tree.num)` trees from `r length(sel.spp)` species in `r comma(analysis.stats$plot.num)` plots that were visited twice by FIA crews. The two visits for `r pct.remeas[2]`% of plots were 10 years apart (`r pct.remeas[1]`% were 9 years, `r pct.remeas[3]`% were 11 years). Plots were distributed in `r analysis.stats$strat.num` strata.  Of the `r length(sel.spp)` species, `r analysis.stats$n.gym.spp` are gymnosperms.  Mean overall growth rates for species varied between `r min.grow.spp$mean` (*`r min.grow.spp$spp`*) and `r max.grow.spp$mean` (*`r max.grow.spp$spp`*, `r t.ref("tbl_growth")`) cm$^2$/decade.  Decadal mean mortality rates varied between `r min.mort.spp$mean` (*`r min.mort.spp$spp`*) and `r max.mort.spp$mean` (*`r max.mort.spp$spp`*, `r t.ref("tbl_mort")`). Growth and mortality rates differed among states (Tables `r tab_nums("tbl_growth", display = "num")` and `r tab_nums("tbl_mort", display = "num")`).

Forested plot 10-year differences in CWD do not visually coincide with initial CWD values (Figure `r fig_nums("fig_cwdmap", display = "num")`); however, `r analysis.stats$cwd.gt0.pct`% of plots increased in their CWD values to some degree during that time period. Tree species exhibited a variety of initial-visit CWD ranges, with means ranging from `r min.mean.init` (*`r min.mean.init.spp`*) to `r max.mean.init` (*`r max.mean.init.spp`*; Table S1).  The CWD difference tended to be comparatively small, with a maximum 95th percentile value of `r max.q95.diff` (*`r max.q95.diff.spp`*, Table S2). The distribution of individual species plot domain values for CWD and change in CWD can be explored on the dashboard (https://tinyurl.com/FIAgrowmort).

The confidence intervals for 10-year mortality and growth rate domains did not necessarily follow the same patterns nor did all confidence intervals for a species' mortality or growth estimates overlap (e.g., Figures `r fig_nums("fig_domains", display = "num")`A and `r fig_nums("fig_domains", display = "num")`B). The relationship between initial CWD values and plot-level change in CWD were not reliably linear (e.g., Figure `r fig_nums("fig_domains", display = "num")`C) and varied by species (see dashboard).

We examined differences in 10-year growth estimates (Figure `r fig_nums("fig_diffgrow", display = "num")`) and mortality estimates (Figure `r fig_nums("fig_diffmort", display = "num")`) among domains.  The column "Above vs. Below Threshold" compares the domains above the CWD change threshold with those below it (e.g., for *Abies lasiocarpa*, subtracting the estimates for columns BL, BM, and BH from AL, AM, AH, respectively; Figure `r fig_nums("fig_domains", display = "num")`A). Negative estimates for the "Above vs. Below Threshold" in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")`  indicate that the 10-year growth or mortality rate was greater for a species in plots below the change estimates, respectively.  `r t.ref("tbl_summ_grow_mort")` summarizes the number of domain differences in each of the columns in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")` where the 95% CI does not include zero as well as the number of comparisons per figure column.

Growth domain comparisons (Figure `r fig_nums("fig_diffgrow", display = "num")`, `r t.ref("tbl_summ_grow_mort")`) did not appear to be produce substantially more "significant" negative vs. positive estimates, and all three columns of Figure `r fig_nums("fig_diffgrow", display = "num")` had similar numbers of "significant" comparisons. Growth may have been susceptible (negatively or positively) to a change in CWD and differed by initial CWD values, and the responses are species-specific. *Abies amabilis*, *Abies grandis*, *Calocedrus decurrens*, *Pinus ponderosa*, *Pseudotsuga menziesii*, *Tsuga mertensiana*, *Acer macrophyllum*, and *Alnus rubra* exhibited greater growth in the Low domains than the Medium and/or High domains.  Except for *Alnus rubra*, these results predominantly occurred in both the Below domain (2) or the Above and Below domains (5).  *Thuja plicata* had higher growth in the Above and Below Medium domains than either High or Low domains. *Abies concolor*, *Larix occidentalis*, *Picea engelmannii*, *Pinus contorta*, *Pinus lambertiana*, *Arbutus menziesii*, and *Quercus kelloggii* had greater growth in the more water stressed High domains than the Medium and/or Low domains. Except for *Larix occidentalis*, these results predominantly occurred in both the Above domain (3) or the Above and Below domains (3). Other species exhibited no 'significant' findings or displayed a mixture of results.

Mortality domain comparisons are similar in certain regards to the growth comparisons: results for the three types of domain comparisons did not appear to be predominantly positive or negative (Figure `r fig_nums("fig_diffmort", display = "num")`, `r t.ref("tbl_summ_grow_mort")`).  There were fewer mortality estimates with "significant" findings than growth. There were fewer differences for "Above vs. Below Threshold" comparisons (`r t.ref("tbl_summ_grow_mort")`) than for the comparisons Above or Below the threshold, indicating that mortality may not be as strongly associated with the change in CWD as by the starting value.  *Larix occidentalis*, *Picea engelmannii*,	*Pinus contorta*,	*Pinus monticola*, *Thuja plicata*,	and *Arbutus menziesii* had greater mortality in the lower water deficit Low domains than Medium and/or High domains. These results occurred in Above (3) or both Above and Below (3) domains. Mortality rates for *Abies lasiocarpa*,	*Calocedrus decurrens*,	*Pinus lambertiana*, *Pinus ponderosa*,	*Pseudotsuga menziesii*,	*Tsuga heterophylla*,	*Tsuga mertensiana*,	*Acer macrophyllum*,	and *Notolithocarpus densiflorus* were greater in the High domains than the Medium and/or Low domains.  These species were approximately evenly split between differences being present in the Above (4) or Below (5) domains, but none with 'significant' findings in both the Above and Below domains.

(Section on geography?  Growth highest for Abies grandis in BL, which coincides with coast range for OR, WA, CA plus Olympics.  CA really only has trees in BL; these probably are the ones that grew the most.  Then, mention some mortality patterns observed below, especially N/NE WA and Sierra Nevadas)

The MC permutation analysis provided empirical p-values for observing contrast confidence intervals with bounds at least as far from zero as those observed (Supplemental figures S3 and S4), with values ranging from $p$ = 0.072 to $p$ < 0.0001 (41 instances).

# Discussion 
We detected numerous species for whom growth and/or mortality rates differed by CWD category.  We also found species for whom growth and mortality differed by state.  Our observed changes in CWD were not large, but regionally, from southern California to the Olympic Peninsula in Washington, CWD increased in 9 out of 10 plots.

Our regional probabilistically sampled design-based results provide a potentially valuable perspective on species status.  Acer macrophyllum growth was negatively associated with initial CWD, and its growth was lower and mortality almost three times higher in California than Washington and Oregon.  We cannot find primary literature that identifies an increase in mortality for this species in California; the degree of mortality observed seems severe.  However, Betzen et al. (2021) found widespread evidence in Washington of A. macrophyllum decline (crown dieback, thinning, discoloration, reduced leaf size) and reduced growth and increased mortality in areas associated with higher temperatures and water limitations.  Although these results align with our CWD findings for growth, they are not reflective (yet) of our state findings and are not apparent in our CWD findings for mortality.  We encourage a more intensive assessment of A. macrophyllum’s status and causes for decline in California.

A converse example is the mortality rate of Thuja plicata, where researchers and state agencies are concerned about die-off but our analysis did not find a strong mortality signal.  Andrus et al. (2023) sampled T. plicata at nine locations in Washington, northern Oregon, and Idaho that were without canopy dieback, exhibited canopy dieback, or had died. They found that the mortality was associated with exceptionally warm and dry climate conditions. A collaboration of state and federal agencies is currently reaching out to the public to identify dieback sites for this species (website: Goodrich et al. 2025).  Our analysis found that T. plicata exhibited the lowest decadal mortality rate and third highest growth rate of any species we examined and the greatest growth in Medium domains and some evidence that mortality in the Above Low domain (Olympic peninsula, western Washington Cascades) was greater than Above High domain (NE Washington, southern Oregon Cascades).  Our analysis may not be temporally sensitive enough to capture the same recent findings as Andrus et al., given that our FIA plot revisit data represented plot visits between 2011 to 2023.  However, our analysis does indicate that numerous species are exhibiting mortality rates at a state level or overall that are higher and are potentially unsustainable (e.g., Abies grandis, Abies lasiocarpa, Picea engelmannii, Pinus contorta, Pinus monticola, and Alnus rubra have state or overall decadal mortality rates greater than 0.25).

Abies lasiocarpa and Picea engelmannii are conspecifics and frequently codominant. Perret et al. (2023) estimated a rangewide mortality of A. lasiocarpa of 25% +/-2.7, with mortality concentrated in the eastern Oregon cascades as well as in Idaho and Colorado. 83% of the mortality was due to fire or biogenic disturbance.  These mortality rates are comparable to the state-level ones we found. Andrus et al. (2018) found that P. engelmannii exhibited higher survival and longevity in undisturbed plots in Colorado, with firs exhibiting mortalities of up to 1% annually while P. engelmannii had mortality rates below 0.5% annually.  We found that P. engelmannii had higher mortality in the Above Low domain.  We note that the Above domains represent a proportionally large increase for Low and Medium given initial CWD values. The same is true for A. lasiocarpa, yet P. engelmannii has greater mortality in the Low domains and A. lasiocarpa has less mortality in the Below Low domain although the MC permutation p-values were not very small.  The difference may be geographical: A. lasiocarpa low-mortality plots are primarily on the Olympic Peninsula of Washington and the northwestern Cascades, where rainfall is abundant.  P. engelmannii is generally absent from these locations.  However, P. engelmannii has the least mortality in the Above domain in the Blue Mountains of Oregon and Washington while A. lasiocarpa does not show a similar finding for the same geographical region. P. engelmannii mortality in northern and northeastern Washington, WA may have been driven by spruce beetles (WADNR 2023).
Our mortality estimate for Pinus contorta in Washington was twice that of Oregon’s and five times greater than Calfornia’s, while the growth confidence intervals for Washignton were lower than California’s and did not overlap. Mortality was highest in the Low CWD domains and growth lowest in the Above Low domain.  The Above Low domain aligns well with plots for northern and northeastern Washington.  Pine bark beetles (Dendroctonus ponderosae, Dendroctonus brevicomis, and Ips spp.) have affected this region for the past decade (WADNR 2023).  This area corresponds with the portions of P. contorta’s range in the lowest forest stability index category for Stanke et al. (2021) while California’s Sierra Nevadas and the Oregon and Washington Cascade Range correspond to positive stability index values.  Stanke et al. also found that insects (followed by fire) were a major driver of the percent change in relative tree density.

Our analysis centered around initial CWD and delta CWD as a variable of interest relative to tree growth and mortality.  There is ample evidence in the literature supporting the relationship or correlation of CWD or its cousin VPD to tree growth (Dyer 2004, Restaino et al. 2016) and mortality (Anderegg et al. 2015, Das et al. 2013, Hammond et al. 2022).  However, there are some challenges in relying on it in a design-based regional analysis.  We did not detect sizable changes in CWD between visits although CWD disproportionately increased between visits across sites.  The changes we observed may have represented greater proportional change for some species than others, with species such as Abies amablis, Alnus rubra, and Juniperus occidentalis exhibiting a low initial CWD range.

Interpreting CWD results at a regional scale becomes difficult because water stress alone can kill trees (Anderegg et al. 2015, Maherali et al.  2004, and van Nieuwstadt and Sheil 2005), but it can also increase the probability of insect (Hart et al. 2014, Preisler et al. 2012 , examples and counterexamples in Huberty and Denno 2004) and wildfire-caused mortality (Abatzoglou and Williams 2016). With reference to Manion’s (1991) “death spiral” (see also Bennett et al. 2023 for a CWD-focused version), teasing apart the cause of mortality from predisposing, inciting, and contributing factors is difficult at best, as these factors can interact (Gaylord et al. 2015).   Interrelated predisposing conditions for individual tree mortality such as canopy openness (Lemmo et al. 2024), structural complexity (Ma et al. 2023), tree height (Stovall et al. 2019), and within-stand competition (Ford et al. 2017) would be missed by our regional evaluation.

# Discussion

The initial CWD values coincide fairly well with mean annual precipitation [NOAA Climate.gov](https://www.climate.gov/news-features/featured-images/new-maps-annual-average-temperature-and-precipitation-us-climate).

# Conclusion

# Acknowledgments

# References

::: {#refs}

Betzen, J.J., Ramsey, A., Omdal, D., Ettl, G.J. and Tobin, P.C., 2021. Bigleaf maple, Acer macrophyllum Pursh, decline in western Washington, USA. Forest Ecology and Management, 501, p.119681.

Wang, T., Hamann, A. Spittlehouse, D.L. and Carroll, C. 2016. Locally downscaled and spatially customizable climate data for historical and future periods for North America. PLoS One 11: e0156720

Alexander, R.R., Shearer, R.C. and Shepperd, W.D., 1990. Abies lasiocarpa (Hook.) Nutt. subalpine fir. Silvics of North America, 1, pp.60-70.

Washington State Department of Natural Resources (DNR) Forest Reliance Division. 2024. Forest health highlights in Washington: A summary of insect, disease, and other disturbance conditions affecting Washington’s forests.

Washington State Department of Natural Resources (WADNR). 2024. Forest Health Highlights in Washington – 2024. Washington State Department of Natural Resources, Forest Resilience Division, Olympia, WA. https://dnr.wa.gov/sites/default/files/2025-03/rp_2023_fhh_report_final_wr.pdf, accessed 9/18/2025

:::

<!---BLOCK_LANDSCAPE_START--->

```{r tbl-growth, tab.cap=tab.tbl_growth}

growth.ft

```

```{r tbl-mort, tab.cap = tab.tbl_mort}

mort.ft

```

<!---BLOCK_LANDSCAPE_STOP--->

```{r mapfig, fig.height = 7,  fig.dpi=300, fig.cap=fig.cwdmap_cap, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "cwd_maps_figure.png"))

```

```{r domainsfig, fig.width = 6,  fig.dpi=300, fig.cap=fig.domains, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "Example_spp_figure.png"))

```

```{r diff_grow_fig, fig.height = 6.5, width = 6.5, fig.dpi=300, fig.cap=fig.diff_grow, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Growth_figs_cwd/Growth_Panel_Plot.png"))

```

```{r diff_mort_fig, fig.height = 6.5, width = 6.5, fig.dpi=300, fig.cap=fig.diff_mort, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Mort_figs_cwd/Mort_Panel_Plot.png"))

```

\newpage

```{r tbl-summ-grow-mort, tab.cap = tab.summ_grow_mort}

g.m.summ.ft

```