### 2025-11-03 Manuscript.Rmd
Latest version of manuscript, added CWD analysis to supplemental materials.

---
title: "Tree growth, mortality, and climatic stress in west-coast states, USA"
author:
  - name: "Jeremiah D. Groom"
    affiliations:
      - name: "Groom Analytics LLC"
        address: "1975 SE Crystal Lake Dr., Unit 173"
        city: "Corvallis"
        state: "OR"
        postal-code: "97333"
    orcid: "0000-0002-0110-1036"
    email: "jeremy@groomanalytics.com"
  - name: "Bryce Frank"
    affiliations:
      - name: "USDA Forest Service"
        department: "Pacific Northwest Field Station"
        address: "3625 93rd Ave SW"
        city: "Olympia"
        state: "WA"
        postal-code: "98512"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  officedown::rdocx_document:
    reference_docx: "template_use2.docx"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(kableExtra)   # Enhanced table formatting
library(scales)       # Scale functions for visualization
library(flextable)    # Professional tables for Word output
library(officedown)
library(officer)
library(rvg)
#devtools::install_github("adletaw/captioner")
library(captioner) # For figure captions.

```

```{r other_settings}
fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)
```

```{r constants}

source("Global.R")
source(paste0(CODE.LOC, "Functions.R"))

# Define any custom functions
format_p_value <- function(p) {
  case_when(
    p < 0.001 ~ "< 0.001",
    p < 0.01 ~ sprintf("= %.3f", p),
    TRUE ~ sprintf("= %.2f", p)
  )
}

set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  theme_fun = "theme_vanilla"
)

centeredP <- fp_par(text.align = "center")

```

```{r load-data}
# Some general analysis info + plots
analysis.stats <- read_rds(paste0(RESULTS.OTHER, "analysis.stats.RDS"))

# Percent of plots sampled with REMPER = 9, 10, and 11 years (rounded)
pct.remeas <- as.vector(analysis.stats$pct.remeas)

# Species scientific names
spp.names.use <- analysis.stats$gym.angio.dat

# Correlation/fits for Initial CWD vs Delta CWD:
CWD.rels <- read_rds(paste0(RESULTS.OTHER, "CWD_deltaCWD_relationship.rds"))
CWD.lin.cor <- CWD.rels$lm1$r.squared^0.5 
CWD.dAIC <- round(CWD.rels$aic.cwd$AIC[1] - CWD.rels$aic.cwd$AIC[2], 0)

# Growth data for species overall, states, differences
grow.state <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/State_Growth_", CLIM.VAR.USE, ".csv"))
grow.all <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/All_Growth_", CLIM.VAR.USE, ".csv"))
grow.diff <- read_rds(paste0(save.loc.fcn(1), "Processed_6Domain_Data.RDS"))

# Mortality data for species overall and states
mort.state <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/State_Mortality_", CLIM.VAR.USE, ".csv"))
mort.all <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/All_Mortality_", CLIM.VAR.USE, ".csv"))
mort.diff <- read_rds(paste0(save.loc.fcn(2), "Processed_6Domain_Data.RDS"))

# Significant findings, Growth
grow.n.AvB <- nrow(grow.diff$AvB2)
grow.n.A <- nrow(grow.diff$A_LMH2)
grow.n.B <- nrow(grow.diff$B_LMH2)

grow.AvB.sig <- length(grow.diff$AvB2$significant[grow.diff$AvB2$significant == TRUE])
grow.AvB.signeg <- length(grow.diff$AvB2$significant[grow.diff$AvB2$significant == TRUE & grow.diff$AvB2$Means < 0])
grow.A.sig <- length(grow.diff$A_LMH2$significant[grow.diff$A_LMH2$significant == TRUE])
grow.A.signeg <- length(grow.diff$A_LMH2$significant[grow.diff$A_LMH2$significant == TRUE & grow.diff$A_LMH2$Means < 0])
grow.B.sig <- length(grow.diff$B_LMH2$significant[grow.diff$B_LMH2$significant == TRUE])
grow.B.signeg <- length(grow.diff$B_LMH2$significant[grow.diff$B_LMH2$significant == TRUE & grow.diff$B_LMH2$Means < 0])

grow.tot.sig <- grow.AvB.sig + grow.A.sig + grow.B.sig

# Significant findings, Mort
mort.n.AvB <- nrow(mort.diff$AvB2)
mort.n.A <- nrow(mort.diff$A_LMH2)
mort.n.B <- nrow(mort.diff$B_LMH2)

mort.AvB.sig <- length(mort.diff$AvB2$significant[mort.diff$AvB2$significant == TRUE])
mort.AvB.signeg <- length(mort.diff$AvB2$significant[mort.diff$AvB2$significant == TRUE & mort.diff$AvB2$Means < 0])
mort.A.sig <- length(mort.diff$A_LMH2$significant[mort.diff$A_LMH2$significant == TRUE])
mort.A.signeg <- length(mort.diff$A_LMH2$significant[mort.diff$A_LMH2$significant == TRUE & mort.diff$A_LMH2$Means < 0])
mort.B.sig <- length(mort.diff$B_LMH2$significant[mort.diff$B_LMH2$significant == TRUE])
mort.B.signeg <- length(mort.diff$B_LMH2$significant[mort.diff$B_LMH2$significant == TRUE & mort.diff$B_LMH2$Means < 0])

mort.tot.sig <- mort.AvB.sig + mort.A.sig + mort.B.sig

# Summary of significant Growth and Mortality findings for table creation.
datasets.grow <- list(grow.diff$AvB2, grow.diff$A_LMH2, grow.diff$B_LMH2)
datasets.mort <- list(mort.diff$AvB2, mort.diff$A_LMH2, mort.diff$B_LMH2)
datatypes <- list("Above vs. Below", "Above", "Below")
n.comp.grow <- list(grow.n.AvB, grow.n.A, grow.n.B)
n.comp.mort <- list(mort.n.AvB, mort.n.A, mort.n.B)

growth.summ <- pmap(list(datasets.grow, datatypes, n.comp.grow), summ.fig.fcn) %>% bind_rows()
mortality.summ <- pmap(list(datasets.mort, datatypes, n.comp.mort), summ.fig.fcn) %>% bind_rows()

```

```{r table-creation}
# State growth
grow.state.cm <- cm2.fcn(1, grow.state)
grow.all.cm <- cm2.fcn(1, grow.all)
s.g.dat <- state.table.fcn(grow.state.cm, grow.all.cm, 1)

growth.ft <- flextable(s.g.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
    italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = TRUE)
min.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = TRUE)

# State mortality
s.m.dat <- state.table.fcn(mort.state, mort.all, 2)

mort.ft <- flextable(s.m.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",  
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = FALSE, digits = 2)
min.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = FALSE, digits = 2)

# CWD amounts
min.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.min.mean]
min.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.min.mean]
max.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.max.mean]
max.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.max.mean]

max.q95.diff <- analysis.stats$diff.cwd.stats$q95[analysis.stats$diff.cwd.95max]
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]
max.range.diff <- analysis.stats$diff.cwd.90range.max
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]

# Adding a section identifier to each table
growth.summ2 <- growth.summ %>%
  mutate(Type = "Growth", .before = 1)

mortality.summ2 <- mortality.summ %>%
  mutate(Type = "Mortality", .before = 1)

# Combine the tables
grow.mort.summ <- bind_rows(growth.summ2, mortality.summ2)

# Create flextable
g.m.summ.ft <- flextable(grow.mort.summ) %>%
  merge_v(j = "Type") %>%  # Merge cells in Section column
  valign(j = "Type", valign = "top") %>%  # Top align section labels
  flextable::align(j = "Type", align = "left") %>%   # Left align section labels
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  set_table_properties(layout = "autofit", width = 0) %>%
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
    flextable::align(j = 3:7, align = "center", part = "all")

```

```{r set_captions}

# 
fig_nums <- captioner::captioner(prefix = "")
tab_nums <- captioner::captioner(prefix = "Table")

f.ref <- function(x) {
 caption_text <- fig_nums(x)
  # Remove any leading numbers and extract just "Figure X"
  stringr::str_extract(caption_text, "\\d+")
}

t.ref <- function(x)  {
  stringr::str_extract(tab_nums(x), "[^:]*")
}

tab.tbl_growth <- tab_nums(name = "tbl_growth", 
                        caption = "Decadal growth estimates by state and overall for analysis species (cm\u00B2/decade). Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots.")

tab.tbl_mort <- tab_nums(name = "tbl_mort", 
                        caption = "Decadal mortality estimates by state and overall for analysis species. Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots.")

fig.cwdmap_cap <- fig_nums(name = "fig_cwdmap", 
                        caption = "Fuzzed FIA plot locations displayed with initial visit CWD values (A) and change in CWD between visits (C) for the U.S. states, north to south, of Washington, Oregon, and California (see inset map B for the location of these states within the USA).")

fig.domains <- fig_nums(name = "fig_domains",
                        caption = "Decadal (A) growth and (B) mortality domain estimates for *Calocedrus decurrens*.  The six domains are defined as FIA plots that fall [A]bove or [B]elow the CWD change threshold with [L]ow, [M]edium, or [H]igh initial CWD values (see text for details). E.g., 'AL' is Above CWD change threshold, Low initial CWD value. The distribution of plot CWD change and initial values (C) is presented as well, with the blue horizontal line indicating the CWD change threshold and the orange vertical lines representing the quantile boundaries between Low, Medium, and High initial CWD values." )

fig.diff_grow <- fig_nums(name = "fig_diffgrow", 
                          caption = paste0("Differences in mean growth rates (cm\u00B2/decade) between domains by species.  Domain abbreviations are described in the legend for Figure ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are coniferous; those below are deciduous. Empirical MC permutation p-values are provided for filled-circle contrasts, with \u2013\u2013 corresponding to p-values of > 0.05, while one to four asterisks correspond to p > 0.01, > 0.001, > 0.0001 and < 0.0001, respectively."))

fig.diff_mort <- fig_nums(name = "fig_diffmort", 
                          caption = paste0("Differences in mean decadal mortality rates between domains by species.  Domain abbreviations are described in the legend for Figure ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are coniferous; those below are deciduous.  Empirical MC permutation p-values are provided for filled-circle contrasts, with \u2013\u2013 corresponding to p-values of > 0.05 while one to four asterisks correspond to p > 0.01, > 0.001, > 0.0001 and < 0.0001, respectively. "))

tab.summ_grow_mort <- tab_nums(name = "tbl_summ_grow_mort",
                               caption = paste0("Summary of number of contrasts with 95% CIs that do not overlap with zero ([T]otal, [P]ositive-, and [N]egative-valued estimates) for Figures ", fig_nums("fig_diffgrow", display = "num"), " and ", fig_nums("fig_diffmort", display = "num"), " by analysis (T)ype (Growth or Mortality) and type of (C)omparison.  The column Domains lists the domains compared; All is the sum of values within a Comparison. See text for other domain labels.  'n' is the total number of differences examined within a Comparison."))

```

# Abstract

# Introduction

There  are signs that tree species in western North America are in regional decline.  Johnston et al. (2025) document a 29% 10-year mortality rate of 300+ yr old trees in Eastern Oregon.  Peret et al. (2022) estimated a decadal range-wide mortality rate of 25% for Subalpine fir (*Picea engelmannii*).  High-severity fires in California in 2020 and 2021 killed 13%-19% of giant sequoia (*Sequoiadendron gianteum*) with >1.07-m bole diameters (Stephenson and Brigham 2021, Shive et al. 2021); recruitment in these burns was depressed (Soderberg et al. 2022).   Five of eight common species in western North America exhibit declines as measured by relative stand density (Stanke et al. 2021).  Shriver et al. (2022) found that four of five juniper (*Juniperus* spp.) and piñon pine (*Pinus* spp.) species were declining in portions of their range, comparing mortality to recruitment rates.

Climate change as manifest through increases or intensification in drought and aridity can affect tree mortality directly through physiological damage and indirectly by increasing wildfire frequency and severity and by promoting insect and disease outbreaks.  Hammond et al. (2022) found globally that the intensity of drought and heat events are directly associated with widespread forest die-off events.  Wildfire frequency and area burned in the western United States has increased, in concert with the longer fire seasons, hotter summer temperatures, and earlier snowmelt (Abatzoglou and Williams 2016, Westerling 2016, Juang et al. 2022).  There has also been a regional increase in the area of bark beetle (subfamily *Scolytinae* spp.) outbreaks (Meddens et al. 2012, Hicke et al. 2016).  Many of these indirect climate-influenced mortality causes may be attributable to the physical condition of trees following abnormal or prolonged exposure to water stress.

Drought-induced mortality can be driven by carbon starvation and hydraulic failure (see reviews by Adams et al. 2010 and McDowell et al. 2008).  Carbon starvation occurs when isohydric species (e.g., piñon pine, *Pinus edulus*; Breshears et al. 2009), or those which mainly rely on stomatal closure to avoid excessive water loss through evapotranspiration (ET), receive insufficient atmospheric carbon dioxide for photosynthesis.  Anisohydric plants, or those who do not strongly regulate water loss through stomatal closure, risk hydraulic failure if sufficient xylem cavitation occurs due to insufficient water uptake needed to meet ET demands (e.g., as seen in *Populus tremuloides*; Anderegg et al. 2015). Adams et al. (2017) performed a review of drought-induced mortality across 26 species and concluded that hydraulic failure occurred broadly across taxa while non-structural carbohydrate reserves at mortality differed across taxa.  Regardless of mechanism, drought (specifically, water deficit) appears to be an important direct driver of mortality for tree taxa (Hammond et al. 2022).

Drought stress can increase tree susceptibility to wildfire, insects, and disease.  Wildfire area (Abatzoglou and Williams, 2016; Parks et al. 2025) and number (Westerling 2016) are positively associated with increasing aridity in the western US.   Bark beetle outbreaks in semi-arid areas of the western US may be positively affected by or associated with acute intense drought and water-stressed trees (Negrón 2009, Gaylord et al. 2015) but chronically stressed trees may not provide bark beetle populations with the resources sufficient to promote population expansion (Huberty and Denno, 2004).  Norway spruce (*Picea abies*) saplings inoculated with the necrotrophic pathogenic fungi *Heterobasidion annosum* sensu stricto and exposed to drought conditions were more likely to die than control saplings or saplings exposed to either drought conditions or inoculation, although Scots pine (*Pinus sylvestris*) seedlings did not show a similar relationship (Gomez-Gallego et al. 2022). Other insects and disease may be promoted or inhibited by drought, depending on the circumstances and due to conflicting factors (Kolb et al. 2016).

One regionally useful metric of water stress is Climatic Water Deficit (CWD; Stephenson 1998) which appears to be a reasonable covariate for examining tree growth and mortality.  Stephenson (1990) defines CWD as evaporative demand not met by available water, or the difference between plant water demand (actual evapotranspiration, AET), and plant water supply (potential evapotranspiration, PET); these variables are driven by temperature and precipitation, respectively.  CWD is geographically readily available from downscaled PRISM data (Daly et al. 2008, Wang et al. 2016).  The Anderegg  et al. (2015) tree mortality study relied on CWD to capture drought occurrence.  Das et al. (2013) found low-elevation California tree mortality best modeled using CWD.  Restaino et al. (2016) found decreased Douglass fir growth with an increase in CWD.

Data from the U.S. National Forest Inventory, collected by the US Forest Service Forest Inventory and Analysis Program (FIA), is well suited for evaluating regional species mortality and growth.   Since 2001 the FIA program has gathered a variety of forestry data at plots along the west coast of the USA using a spatially balanced probability sample (McRoberts 2005, Reams et al. 2005).  The sample allows for a relatively unbaised design-based inference of population characteristics across a large geographical region.  Data from plot revisits and individual tree data permit direct estimation of change features such as mortality and growth estimation.

Monleon and Linz (2015) evaluated the evidence of tree species range shifts by subtracting the design-based mean estimates of annual FIA plot temperatures for seedlings from those of trees. FIA plot revisit data were mostly incomplete at the time of their analysis.  Groom and Monleon (2023) used the revisit data to compare plot climate variable values (vapor pressure deficit [VPD], temperature, and precipitation) to estimate shifts in plot occupancy and density.  Both studies used a single climate variable value associated with each plot for estimation; e.g., Groom and Monleon (2023) examined shifts in species density by estimating the difference in VPD means between plots with more trees of a given species and plots with fewer trees.  They documented, however, that the initial climate variable values could (and often did) change between plot visits, with species often exhibiting positive or negative correlations between the initial visit and between-visits change value of the climate variable.  They identified that their design-based analysis could potentially improve by including both initial visit and change values for the climate variable.

The purpose of this analysis is to obtain design-based estimates of tree species’ growth and mortality across the western U.S. states of California, Oregon, and Washington as they relate to plot-level initial-visit water stress (Climatic Water Deficit, initial CWD) and the change in water stress (delta CWD) over time.  We approached this bivariate analysis by dividing species’ plot data into domains based on initial and delta CWD values, finding growth and mortality estimates for each domain, and estimating domain contrasts to establish differences between domains.  Our secondary goal includes providing decadal growth and mortality estimates for our analysis species overall and by state to establish species status and to inform our CWD analysis findings.

# Methods

## FIA data

Our analysis relies on FIA data for forestland in California, Oregon, and Washington, USA.  Forestland is defined as an area of 4,050 m$^2$ that is at least 10 percent stocked or potentially stocked with tree species, excluding urban and agricultural land uses (Reams et al., 2005).  We constrained our analysis to only include remeasurements on areas considered forestland during both visits.

FIA plots consist of four points where all trees with stem diameters greater than or equal to 12.7 cm (for most species measured at 1.37 m above the ground) were tallied in a 7.32 m radius circular subplot (total area 672.5 m$^2$) with additional trees sampled in an anulus around each subplot (macroplot), with a total radius 18 m from the center point and a total area of 1012 m$^2$.  We did not include data from FIA microplots.  Bechtold and Scott (2005) provide plot design details and measurement protocols.

Since 2000, FIA plots have been selected via spatially-balanced probability sampling (McRoberts 2005, Reams et al., 2005) with a target plot density of one ground plot every 24 km$^2$. Some regions and ownerships intensify the plot density; we excluded these additional plots to simplify stratification estimation, relying on FIA-defined strata and using the density of FIA Phase 1 sampling points within and across strata to determine strata weights.

Plots in the three states are sampled every 10 years with a target of 1/10th of plots sampled in a panel every year (Thompson 2015). Our analysis depends on examining the fate and growth of trees from two sequential visits (referred to as the first and second plot visit), ideally from a complete sample panel remeasure.  We obtained our FIA data from the USDA FIA DataMart web portal (https://research.fs.usda.gov/products/dataandtools/fia-datamart; downloaded May 5, 2025).  The database included plot data sampled from 2001 to 2023.

Individual trees within plots were retained in the analysis if they were alive at the initial visit and met the stem diameter criterion during that visit (i.e., plot ingrowth was not included).  The growth analysis additionally restricted trees considered to those that were alive at both visits.  The mortality analysis excluded trees killed through silvicultural practices.  Both analyses excluded trees that shrank in diameter, physically moved due to landslides or other event, were erroneously tallied at either visit, were missed due to data collection procedural changes, or existed at any time in a nonsampled area.  We used only FIA plots sampled specifically for tree mortality and growth.

## Climate data

We obtained plot-level CWD data from the model TerraClimate (Abatzoglou et al. 2018) for June, July, and August (summer months) between 1980 and 2023. Latitude and longitude were based on FIA-fuzzed coordinates.  We found the 20-year mean of summer CWD values for each plot prior to and including the first visit year.  We refer this this value as the “initial CWD value”. We then found the mean of summer CWD values between visits and including the second visit year.  The difference between this second value and initial CWD value is the “delta CWD value”.  Both the initial CWD and delta CWD values for FIA plots are mapped in Figure 1.

## Analysis
Following Section 4.3.4 in Scott et al. (2005) we obtained point estimates of species growth and mortality using ratio estimators (Eqn 4.16).  The numerator for species-specific mortality is the sum of stratum-weighted means of the number of trees that died between visits per plot.  The number of sampled trees in each plot is expanded by the area sampled (subplot or macroplot area) adjusted by the proportion of the plot area available for sampling and the number of years between visits.    The denominator for mortality is similar but is calculated as the sum of stratum-weighted means of all trees per plot of a given species.

The numerator of species-specific growth estimates is the sum of stratum-weighted means of the increase in basal area (cm$^2$) between visits per plot while the denominator is the same as was used for the mortality estimate. This produces a point estimate for a species’ mean growth.  Mortality and growth values were standardized to a decadal time period.

The analysis finds species growth and mortality estimates for six domains based on CWD values (Figure 2C).   The initial CWD values are divided into three species-specific quantiles: values below the first quartile (Low), between the first and third quartiles (Medium), and above the third quartile (High).  Delta CWD plot values for all species were divided into these above a 5mm CWD change threshold (Above) and those below the threshold (Below). The 5mm threshold was selected because it is approximately the mean delta CWD value across species and because species' delta CWD value distributions did not differ greatly (Table S2).  The category Below includes negative delta CWD values (i.e., mean CWD was larger prior to the initial visit than between visits).  The domain designations AL, AM, and AH correspond to Above Low, Above Medium, and Above High. The Below domains are similarly labeled (BL, BM, BH). We also report species mortality and growth estimates for all plots across and within Oregon, Washington, and California irrespective of CWD values.  We limited the number of species examined to those with more than 400 occupied plots across the three states.  We calculated domain estimates for all CWD-based domains with 10 or more plots.

We examined contrasts in 10-year growth estimates (Figure `r fig_nums("fig_diffgrow", display = "num")`) and mortality estimates (Figure `r fig_nums("fig_diffmort", display = "num")`) among domains.  The figure column "Above vs. Below Threshold" displays the mean point estimate and 95% CI for the contrast between corresponding Above and Below domains (e.g., for *Abies lasiocarpa*, subtracting the estimates for columns BL, BM, and BH from AL, AM, AH, respectively; Figure `r fig_nums("fig_domains", display = "num")`A). Negative estimates for the "Above vs. Below Threshold" in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")`  indicate that the 10-year growth or mortality rate was greater for a species in plots below the change estimates.  `r t.ref("tbl_summ_grow_mort")` summarizes the number of domain differences in each of the columns in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")` where the 95% CI does not include zero (termed "significant" findings) as well as the number of contrasts per figure column.

We obtained empirical means and 95% confidence intervals via bootstrap. For each of the 1000 bootstrap iterations we sampled FIA plots with replacement and obtained point estimates for each species' domain.  For each domain we calculated a mean, upper, and lower 95% CI with the 2.5% and 97.5% quantiles from the  sampling iterations.  This approach allows for nonsymmetrical CIs around means.  We estimated contrasts between pairs of CWD domains by subtracting within-iteration domain point estimates.

Our analysis approach results in many comparisons.  We conducted a Monte Carlo (MC) permutation analysis to provide species-level false positive rates (p-values) for "significant" contrasts. The MC permutation analysis had 1000 iterations; each iteration estimated each species' bootstrap confidence interval as described above. The plot domain assignments for each species were sampled without replacement.  The distance of MC permutation contrast CI bounds from zero were compared against test CI bounds. The MC permutation procedure recorded a null significant value if the absolute values of the permutation iteration CI bounds were at least as far from zero as observed bounds.

## Dashboard
We developed a data dashboard (https://tinyurl.com/FIAgrowmort) to allow readers to examine, for individual species, a species' domain estimates, the distribution of initial and delta CWD values, and the geographical distribution of plots and plot domains.  All analysis files, manuscript markdowns, dashboard code, and ancillary data sources are available at the GitHub repository https://github.com/jeremygroom/Tree_Growth_Death.

# Results

Our study tracked the fate of `r comma(analysis.stats$tree.num)` trees from `r length(sel.spp)` species in `r comma(analysis.stats$plot.num)` plots that were visited twice by FIA crews. Of the `r length(sel.spp)` species, `r analysis.stats$n.gym.spp` are gymnosperms.  The two visits were 10 years apart for `r pct.remeas[2]`% of plots (`r pct.remeas[1]`% were 9 years, `r pct.remeas[3]`% were 11 years). Plots were distributed in `r analysis.stats$strat.num` strata.

Mean overall growth rates for species varied between `r min.grow.spp$mean` (*`r min.grow.spp$spp`*) and `r max.grow.spp$mean` (*`r max.grow.spp$spp`*, `r t.ref("tbl_growth")`) cm$^2$/decade.  Six species had greater growth (non-overlapping CIs) in California than in Oregon and/or Washington: *Abies concolor*, *Abies grandis*, *Pinus ponderosa*, *Tsuga heterophylla*, *Arbutus menziesii*, and *Notolithocarpus densiflorus*.  Three species had greater growth in Oregon (*Calocedrus decurrens*, *Picea engelmanii*, and *Pseudotsuga menziesii*) and two species had greater growth in Washington (*Abies amablis* and *Pinus monticola*).  Decadal mean mortality rates varied between `r min.mort.spp$mean` (*`r min.mort.spp$spp`*) and `r max.mort.spp$mean` (*`r max.mort.spp$spp`*, `r t.ref("tbl_mort")`). California had four species with greater mortality rates than Oregon and/or Washington: *Abies concolor*, *Pinus ponderosa*, *Acer macrophyllum*, and *Quercus kelloggii*.  Oregon had one species with a greater mortality rate (*Tsuga mertensiana*) and Washington had four (*Abies grandis*, *Larix occidentalis*, *Picea engelmannii*, and *Pseudotsuga menziesii*).

Delta CWD increased at `r analysis.stats$cwd.gt0.pct`% of plots. Initial and delta CWD values were positively linearly associated (correlation = `r CWD.lin.cor`); however, the two variables were better modeled by a 3rd order polynomial ($\Delta AIC = $ `r CWD.dAIC`, Figure S1, see model fits in Tables S1 and S2), and for most species delta CWD neither linearly related to nor monotonically increased with initial CWD values (Figure S2; also, see dashboard).  Tree species exhibited a variety of initial-visit CWD ranges, with means ranging from `r min.mean.init` (*`r min.mean.init.spp`*) to `r max.mean.init` (*`r max.mean.init.spp`*; Table S3).  Delta CWD values tended to be comparatively small, with a maximum 95th percentile value of `r max.q95.diff` (*`r max.q95.diff.spp`*, Table S4).

Individual species growth and mortality domain patterns were not always similar (e.g., Figures `r fig_nums("fig_domains", display = "num")`A and `r fig_nums("fig_domains", display = "num")`B). Growth domain contrasts (Figure `r fig_nums("fig_diffgrow", display = "num")`, `r t.ref("tbl_summ_grow_mort")`) did not appear to result in substantially more "significant" negative or positive estimates, and all three columns of Figure `r fig_nums("fig_diffgrow", display = "num")` had over 20 "significant" contrasts. Growth may have been susceptible (negatively or positively) to a change in CWD and differed by initial CWD values, and the responses are species-specific. *Abies amabilis*, *Abies grandis*, *Calocedrus decurrens*, *Pinus ponderosa*, *Pseudotsuga menziesii*, *Tsuga mertensiana*, *Acer macrophyllum*, and *Alnus rubra* exhibited greater growth in the Low domains than the Medium and/or High domains.  Except for *Alnus rubra*, these results predominantly occurred in both the Below domain (2 spp.) or the Above and Below domains (5 spp.).  *Thuja plicata* had greater growth in the Above and Below Medium domains than either High or Low domains. *Abies concolor*, *Larix occidentalis*, *Picea engelmannii*, *Pinus contorta*, *Pinus lambertiana*, *Arbutus menziesii*, and *Quercus kelloggii* had greater growth in the more water stressed High domains than the Medium and/or Low domains. Except for *Larix occidentalis*, these results predominantly occurred in both the Above domain (3 spp.) or the Above and Below domains (3 spp.). Other species exhibited no 'significant' findings (4 spp.) or displayed other result patterns (3 spp.).

Mortality domain contrasts are similar in certain regards to the growth contrasts: results for the three types of domain contrasts did not appear to be predominantly positive or negative (Figure `r fig_nums("fig_diffmort", display = "num")`, `r t.ref("tbl_summ_grow_mort")`).  There were fewer mortality estimates with "significant" findings than growth (`r mort.tot.sig` vs `r grow.tot.sig`. There were fewer differences for "Above vs. Below Threshold" contrasts (`r mort.n.AvB`, `r t.ref("tbl_summ_grow_mort")`) than for Above (`r mort.A.sig`) or Below (`r mort.B.sig`), indicating that mortality may not be as strongly associated with delta CWD as by the initial CWD value.  *Larix occidentalis*, *Picea engelmannii*,	*Pinus contorta*,	*Pinus monticola*, *Thuja plicata*,	and *Arbutus menziesii* had greater mortality in the lower water deficit Low domains than Medium and/or High domains. These results occurred in Above (3 spp.) or both Above and Below (3 spp.) domains. Mortality rates for *Abies lasiocarpa*,	*Calocedrus decurrens*,	*Pinus lambertiana*, *Pinus ponderosa*,	*Pseudotsuga menziesii*,	*Tsuga heterophylla*,	*Tsuga mertensiana*,	*Acer macrophyllum*,	and *Notolithocarpus densiflorus* were greater in the High domains than the Medium and/or Low domains.  These species were approximately evenly split between differences being present in the Above (4 spp.) or Below (5 spp.) domains, but none with 'significant' findings in both the Above and Below domains.  Five species had no significant contrasts.

The MC permutation analysis provided empirical p-values for observing contrast confidence intervals with bounds at least as far from zero as those observed (Supplemental figures S5 and S6), with values ranging from a maximum of $p$ = 0.072 to a minimum $p$ < 0.0001 (41 instances).  Except where explicitly stated, we only discuss individual species' domain findings if the MC permutation p-values are < 0.001 (indicated by three asterisks in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")`)

# Discussion

Across the 23 species examined we found no clear pattern of response to initial and/or delta CWD.  Responses were species-specific, potentially driven in some cases by local phenomena and captured by a domain category or categories.  We did find more “significant” CWD contrasts for species growth than mortality, possibly indicating that growth differences were more responsive as an indicator than mortality over a decade, although growth, although growth and mortality findings were not necessarily negatively related.  Delta CWD increased almost everywhere between plot revisits although the magnitude of change was generally not large relative to initial CWD values.  We additionally found a range of decadal mortality rates by state and region, some of which were concerningly high, such as those for *Picea engelmannii*, *Pinus contorta* in Washington, *Acer macrophyllum* in California, *Abies lasiocarpa*, and *Abies grandis* in Washington.

We believe the value of our analysis lies in its application to a relatively large number of tree species and its approach for investigating life history conditions such as growth and mortality that require an estimate of condition change.  The availability of revisit data for FIA datasets is a relatively recent development.  Prior to the availability of the revisit data, Monleon and Lintz (2015) found design-based estimates for the difference in mean sapling and mature tree plot variable values for (“initial”) plot temperature values to assess tree species’ range shifts.  Groom and Monleon (2023), also using design-based estimation approaches, examined FIA revisit data and found evidence of tree species density shifts along initial values of climatic variables but documented evidence of interactions between those variables’ initial and change values. We are not aware of other design-based estimates of revisit data that examine more than one variable of interest simultaneously. Our approach allows us to capitalize on the amenability of the data to design-based estimation while accounting for the need, when examining revisit conditions, of capturing the initial and change value conditions.  Examples of model-based analyses of mortality and growth using FIA revisit data include Shriver et al. (2022), Ogana et al. (2025), and Brandeis and Brandeis (2021).  Examples of growth and mortality design-based estimation that rely on FIA revisit data, similar to our state and overall analyses, include Vickers and Goff (2025), Goeking and Windmuller-Campione (2021), and Perret et al. (2023). Stanke et al. (2021) found estimates of condition indices.  Our bivariate examination of plot CWD shares some dimensional similarities with Dobrowski et al. (2013)  who examined the velocity of CWD (change in distance over time) along with actual evapotranspiration (AET) and minimum temperatures.

Our analysis relied on plot-level initial CWD and delta CWD values as variables of interest for examining tree growth and mortality.  CWD is a reasonable variable to use for examining regional water stress as it integrates climatic and soil water availability through considering precipitation and air temperatures (Stephenson 1990, 1998).  There is ample evidence in the literature supporting the relationship or correlation between CWD and tree growth (Dyer 2004, Restaino et al. 2016) and mortality (Anderegg et al. 2015, Das et al. 2013, Hammond et al. 2022).  However, there are some challenges in relying on CWD or any other climate variable in a design-based regional analysis.  We did not detect sizable changes in CWD between visits, potentially signalling that 10 years is too soon to observe CWD-related species responses (perhaps moreso for mortality than growth).  The changes we observed may have represented greater proportional change for some species than others, with species such as *Abies amablis*, *Alnus rubra*, and *Juniperus occidentalis* exhibiting narrow and low water stress initial CWD ranges.  CWD effects can manifest in direct and indirect ways, potentially differing with species by geographic location and the biotic and abiotic impacts to those locations.  It also cannot easily convey differences in impacts to trees within the same stands.

Interpreting CWD results at a regional scale is difficult because evidence of water stress can manifest in multiple overlapping ways.  Water stress alone can kill trees (Anderegg et al. 2015, Maherali et al.  2004, and van Nieuwstadt and Sheil 2005), but it can also increase the probability of insect (Hart et al. 2014, Preisler et al. 2012 , examples and counterexamples in Huberty and Denno 2004) and wildfire-caused mortality (Abatzoglou and Williams 2016). With reference to Manion’s (1991) “death spiral” (see also Bennett et al. 2023 for a CWD-related version), teasing apart the cause of mortality from predisposing, inciting, and contributing factors is difficult at best, as these factors can interact (Gaylord et al. 2015).   Interrelated predisposing conditions for individual tree mortality such as canopy openness (Lemmo et al. 2024), structural complexity (Ma et al. 2023), tree height (Stovall et al. 2019), and within-stand competition (Ford et al. 2017) would be missed by our regional evaluation.

Although our current analysis offers a coarse examination of species status across a large area, our bivariate approach for assessing tree species growth and mortality by domains could be refined.  Our climate domains could be divided further into sub-domains: trees could be assigned to tree size, stand competition, plot aspect, or plot elevation categories within the climate domain categories.  This approach has its limits: only estimates for certain widespread tree species would be reasonable to report, and domains could not be divided too finely into sub-domains, limiting the number of variables that could be considered simultaneously.  However, this approach would permit design-based estimation at a resolution better attuned to capturing factors affecting growth and mortality.

We found some suggestion that growth patterns with MC permutation p-values less than < 0.001 were more frequently positive than negative for contrasts of Above or Below initial CWD values.  This may indicate that more tree species are exhibiting greater growth at plots with less initial water stress (e.g., *Pseudotsuga menziesii*, *Pinus ponderosa*, and *Acer macrophyllum*) with counterexamples (e.g., *Pinus contorta*).  We found fewer such contrasts for the decadal mortality analysis of CWD relative to decadal growth, with a fairly even number of significant contrasts for both Above and Below initial CWD values. Tree species could exhibit higher mortality at higher or lower initial water stress values.

The number of "significant" Above vs. Below threshold contrasts for growth and mortality were approximately evenly distributed between positive and negative contrast estimates, with a positive difference indicating that some species exhibited higher growth or greater mortality in an Above domain (e.g., *Pinus lambertiana* for both growth and mortality) versus a Below domain (*Abies grandis* and *Pinus contorta*, respectively, for growth and mortality).  This lack of a pattern across species may indicate that the distinction of Above vs. Below is generally not important, or that ‘significant’ findings are due to confounding factors such as geographical locations with different ecological pressures.

The interpretation of contrast results may be more useful at an individual species level and when considering overall mortality and growth estimates.  Regional analyses detected several species with relatively high mortality rates.  *Acer macrophyllum* growth was negatively associated with initial CWD values, and its growth was lower and mortality almost three times higher in California than in Washington and Oregon.  We cannot find primary literature that corroborates an increase in mortality for this species in California; the degree of mortality observed seems severe. McNellis et al. (2021) estimated a recent *Acer macrophyllum* overall mortality rate of 0.45%/yr (4.5%/decade).  However, Betzen et al. (2021) found widespread evidence in Washington of *Acer macrophyllum* decline (crown dieback, thinning, discoloration, reduced leaf size) and reduced growth and increased mortality in locations with higher temperatures and water limitations.  Although these results align with our regional CWD findings for growth, as displayed in our dashboard, they are not reflective (yet) of our state mortality findings for Washington.  We encourage a more intensive assessment of *Acer macrophyllum*’s status and causes for decline in California.

Our regional analyses found several other species with mortality rates that are relatively high and are potentially unsustainable (e.g., *Abies grandis*, *Abies lasiocarpa*, *Picea engelmannii*, *Pinus contorta*, *Pinus monticola*, and *Alnus rubra* have state or overall decadal mortality rates greater than 0.25).  We do not believe our results are concerning for *Alnus rubra*.  Our estimates for this short-lived tree species align with reported background mortality rates (e.g., Brown and Courtin 2018).

Conspecifics *Abies lasiocarpa* and *Picea engelmannii* are frequently codominant and exhibit high mortality rates in our study, likely due to wildfire and insects. Andrus et al. (2018) found that *Picea engelmannii* exhibited higher survival and longevity compared to *Abies lasiocarpa* in undisturbed plots in Colorado, with firs exhibiting mortalities of up to 1% annually while *Picea engelmannii* had mortality rates below 0.5% annually, while our overall estimated mortality rates were substantially higher (although our mortality estimate for *Picea engelmannii* in Oregon was similar).  Perret et al. (2023) estimated a rangewide mortality of *Abies lasiocarpa* of 25% +/-2.7, with mortality concentrated in the eastern Oregon Cascades as well as in Idaho and Colorado. They found that 83% of the mortality was due to fire or biogenic disturbance.  These mortality rates are comparable to the state-level ones we found.  We found that *Picea engelmannii* had lower mortality in the Above High domain, corresponding to stands in the Blue Mountains of Oregon and Washington, and higher mortality in the Above Low and Above High domains, largely coinciding with northern and northeastern Washington, where spruce beetle were elevated (WADNR 2023).

Our mortality estimate for *Pinus contorta* in Washington was twice that of Oregon’s and five times greater than Calfornia’s, while the growth confidence intervals for Washignton were lower than California’s and did not overlap. Mortality was highest in the Low CWD domains and growth lowest in the Above Low domain.  The Above Low domain aligns well with plots for northern and northeastern Washington.  Pine bark beetles (*Dendroctonus ponderosae*, *Dendroctonus brevicomis*, and *Ips* spp.) have affected this region for the past decade (WADNR 2023).  This area corresponds with the portions of *Pinus contorta*’s range in the lowest forest stability index category for Stanke et al. (2021) while California’s Sierra Nevadas and the Oregon and Washington Cascade Range correspond to positive stability index values.  Stanke et al. also found, from FIA plot data, that insects (followed by fire) were related to the percent change in relative tree density.

*Abies grandis* exhibited higher mortality in Washington than Oregon or California. We found no evidence of a difference in domain mortality geographically.  This result may be due to an intersection in a fire suppression history promoting a densification of shade-tolerant but drought-intolerant firs in the arid climates of eastern Oregon and Washington, drought, and fir engraver beetle mortality (*Scolytus ventralis*; Washington DNR 2023, Merschel et al. 2021).  Johnson et al. (2025) report mortality rates of *Abies grandis* in protected areas in the Malheur National Forest of Oregon that exceed ours, primarily attributing fire as the cause.

Our dashboard reveals some instances of domains representing geographical mortality differences.  State estimates for *Pinus monticola* mortality do not appear to differ. Mortality for the Above High domain was low (see dashboard; MC permutation test results only found p < 0.01 when comparing Above High to Above Low) and predominantly represented by plots in the southern Sierra Nevada mountains.  *Pinus lambertiana* and *Calocedrus decurrens* exhibited high mortality in approximately the same region.  This region experienced an extreme drought that killed 170m trees between 2010 and 2021 (OEHHA 2022).  Dudney et al. (2020) also estimated lower mortality rates for *Pinus monticola* than for *Pinus lambertiana* in the Sequoia and Kings Canyon National Parks, which are within this region.  They attributed some of the difference in mortality to elevation, with mountain pine beetle (*Dendroctonus ponderosae*), fire, and white pine blister rust (*Cronartium ribicola*) more prevalent at lower elevations, corresponding to the range of *Pinus lambertiana*. Fettig et al. (2019) separately corroborated high tree mortality rates for *Pinus lambertiana* and *Calocedrus decurrens* in this region.

An example species for which our results did not corroborate with other researchers is *Thuja plicata*.  Researchers and state agencies are currently concerned about die-off but our analysis did not find a strong mortality signal.  Andrus et al. (2023) sampled *Thuja plicata* at nine locations in Washington, northern Oregon, and Idaho that were without canopy dieback, exhibited canopy dieback, or had died. They found that the mortality was associated with exceptionally warm and dry climate conditions. A collaboration of state and federal agencies is currently reaching out to the public to identify dieback sites for this species (website: Goodrich et al. 2025).  Our analysis found that *Thuja plicata* exhibited the lowest decadal mortality rate and third highest growth rate of any species we examined. It also had its greatest growth in Medium domains.  Our analysis may not be temporally sensitive enough to capture the same recent findings as Andrus et al., given that our FIA plot revisit data represented plot visits between 2011 to 2023.  If species conditions change quickly (on the order of 3-5 years) the FIA plot data may not reflect the change until a full or nearly full panel of plots is complete following the change in conditions.

# Conclusion

We hope our approach, results, and dashboard assist in refining future species inquiries.

# Acknowledgments

# References

::: {#refs}
Abatzoglou, J.T., Dobrowski, S.Z., Parks, S.A. and Hegewisch, K.C., 2018. TerraClimate, a high-resolution global dataset of monthly climate and climatic water balance from 1958–2015. Scientific data, 5(1), pp.1-12.

Alexander, R.R., Shearer, R.C. and Shepperd, W.D., 1990. Abies lasiocarpa (Hook.) Nutt. subalpine fir. Silvics of North America, 1, pp.60-70.

Betzen, J.J., Ramsey, A., Omdal, D., Ettl, G.J. and Tobin, P.C., 2021. Bigleaf maple, Acer macrophyllum Pursh, decline in western Washington, USA. Forest Ecology and Management, 501, p.119681.

Brandeis, T. and Brandeis, C., 2021. Assessing tree mortality probability in harvested hardwood stands using long-term forest inventory data. Forest Science, 67(2), pp.231-240.

Breshears, D.D., Myers, O.B., Meyer, C.W., Barnes, F.J., Zou, C.B., Allen, C.D., McDowell, N.G. and Pockman, W.T., 2009. Tree die‐off in response to global change‐type drought: mortality insights from a decade of plant water potential measurements. Frontiers in Ecology and the Environment, 7(4), pp.185-189.

Brown, K.R. and Courtin, P.J., 2018. Can phosphorus additions increase long-term growth and survival of red alder (Alnus rubra Bong.) on periodically dry sites?. Forest Ecology and Management, 430, pp.545-557.

Daly C, Halbleib M, Smith JI, Gibson WP, Doggett MK, Taylor GH, et al. Physiographically sensitive mapping of climatological temperature and precipitation across the conterminous United States. Int J Climatol. 2008; 28(15):2031–64. doi: 10.1002/joc.1688 PMID: WOS:000261810700004.

Goeking, S.A. and Windmuller-Campione, M.A., 2021. Comparative species assessments of five-needle pines throughout the western United States. Forest Ecology and Management, 496, p.119438.

McRoberts, R.E. 2005. The Enhanced Forest Inventory and Analysis Program, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Merschel, A.G., Beedlow, P.A., Shaw, D.C., Woodruff, D.R., Lee, E.H., Cline, S.P., Comeleo, R.L., Hagmann, R.K. and Reilly, M.J., 2021. An ecological perspective on living with fire in ponderosa pine forests of Oregon and Washington: Resistance, gone but not forgotten. Trees, Forests and People, 4, p.100074.

Ogana, F.N., Green, P.C. and Radtke, P., 2025. Modeling the growth and yield of natural hardwood stands in the southern United States using the Forest Inventory and Analysis data. Forest Ecology and Management, 586, p.122722.

Office of Environmental Health Hazard Assessment (OEHHA, 2022). Indicators of climate change in california, fourth edition, California Environmental Protection Agency, OEHHA.

Reams, G.A, Smith, W.D., Hansen, M.H., Bechtold, W.A., Roesch, F.A., Moisen, G.G. 2005. The forest inventory and analysis sampling frame, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Scott, C.T., W. A. Bechtold, G.A. Reams, W.D. Smith, J.A. Westfall, M.H. Hansen, Moisen, G.G. 2005. Sample-based estimators used by the Forest Inventory and Analysis
National Information Management System, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design
and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Stephenson, N.L., 1990. Climatic control of vegetation distribution: the role of the water balance. The American Naturalist, 135(5), pp.649-670.

Stephenson, N., 1998. Actual evapotranspiration and deficit: biologically meaningful correlates of vegetation distribution across spatial scales. Journal of biogeography, 25(5), pp.855-870.

Vickers, L.A. and Goff, T.C., White Oak Sapling Dynamics Across the Eastern Us: Quantifying a Bottleneck to Sustainability. Available at SSRN 5398050.

Washington State Department of Natural Resources (WADNR) Forest Reliance Division. 2024. Forest health highlights in Washington: A summary of insect, disease, and other disturbance conditions affecting Washington’s forests.  Washington State Department of Natural Resources, Forest Resilience Division, Olympia, WA. https://dnr.wa.gov/sites/default/files/2025-03/rp_2023_fhh_report_final_wr.pdf, accessed 9/18/2025

:::

<!---BLOCK_LANDSCAPE_START--->

```{r tbl-growth, tab.cap=tab.tbl_growth}

growth.ft

```

```{r tbl-mort, tab.cap = tab.tbl_mort}

mort.ft

```

<!---BLOCK_LANDSCAPE_STOP--->

```{r tbl-summ-grow-mort, tab.cap = tab.summ_grow_mort}

g.m.summ.ft

```

```{r mapfig, fig.height = 7,  fig.dpi=300, fig.cap=fig.cwdmap_cap, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "cwd_maps_figure.png"))

```

```{r domainsfig, fig.width = 6,  fig.dpi=300, fig.cap=fig.domains, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "Example_spp_figure.png"))

```

```{r diff_grow_fig, fig.height = 6.5, width = 6.5, fig.dpi=300, fig.cap=fig.diff_grow, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Growth_figs_cwd/Growth_Panel_Plot.png"))

```

```{r diff_mort_fig, fig.height = 6.5, width = 6.5, fig.dpi=300, fig.cap=fig.diff_mort, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Mort_figs_cwd/Mort_Panel_Plot.png"))

```