### 2025-11-03 Supplemental Material.Rmd

---
title: "Supplemental material for Tree growth, mortality, and climatic stress in west-coast states, USA"
date: "`r Sys.Date()`"
author: "Jeremiah D. Groom"
output: 
  officedown::rdocx_document:
    mapstyles:
      Normal: ['First Paragraph']
    reference_docx: "template_use2.docx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(officedown)
library(officer)
library(kableExtra)   # Enhanced table formatting
library(scales)       # Scale functions for visualization
library(flextable)    # Professional tables for Word output
library(rvg)
library(captioner) # For figure captions.

fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)

set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  theme_fun = "theme_vanilla"
)

```

```{r load_data}

source('Global.R')
source(paste0(CODE.LOC, "Functions.R"))
# Some general analysis info + plots
analysis.stats <- read_rds(paste0(RESULTS.OTHER, "analysis.stats.RDS"))
# Correlation/fits for Initial CWD vs Delta CWD:
CWD.rels <- read_rds(paste0(RESULTS.OTHER, "CWD_deltaCWD_relationship.rds"))

```

```{r cwd_summary_tables}
ft.lm1 <- CWD.rels$ft_lm1

cwd.init <- analysis.stats$init.cwd.stats 
cwd.diff <- analysis.stats$diff.cwd.stats

init.ft <- flextable(cwd.init) %>%
  fontsize(size = 10, part = "body") %>%
  fontsize(size = 11, part = "header") %>%
  italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  set_table_properties(layout = "autofit", width = 0) %>%
  border_remove() %>%
  hline_top(part = "all") %>%
  hline_bottom(part = "all")

diff.ft <- flextable(cwd.diff) %>%
  fontsize(size = 10, part = "body") %>%
  fontsize(size = 11, part = "header") %>%
  italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  set_table_properties(layout = "autofit", width = 0) %>%
  border_remove() %>%
  hline_top(part = "all") %>%
  hline_bottom(part = "all")

```

```{r domain_pvals}

psig.dat2 <- analysis.stats$psig.dat2
psig.g <- psig.dat2[, 1:7]
psig.m <- psig.dat2[, c(1, 8:13)]

psig.ft.g <- psig.ft.fcn(psig.g)
psig.ft.m <- psig.ft.fcn(psig.m)

```

```{r legends}
fig_nums <- captioner::captioner(prefix = " S")
tab_nums <- captioner::captioner(prefix = "Table S")

f.ref <- function(x) {
 caption_text <- fig_nums(x)
  # Remove any leading numbers and extract just "Figure X"
  stringr::str_extract(caption_text, "\\d+")
}

t.ref <- function(x)  {
  stringr::str_extract(tab_nums(x), "[^:]*")
}

tbl_cwd_lm1 <- tab_nums(name = "tbl_cwdlm1", 
                         caption = "Linear regression model output for the linear fit of initial and delta CWD plot data (green line in Figure S1). Variable name CWD refers to initial CWD value.")

tbl_cwd_lm2 <- tab_nums(name = "tbl_cwdlm2", 
                         caption = "Linear regression model output for the third-order polynomial fit of initial and delta CWD plot data (orange line in Figure S1). Variable name CWD refers to initial CWD value.")

tbl_cwd_init <- tab_nums(name = "tbl_cwdinit", 
                         caption = "Summary statistics for initial-visit plot CWD values by species. The number of species' plots (n), minimum, mean, median, and maximum CWD values are provided, along with the 5%, 25%, 75%, and 95% quantiles (q).")

tbl_cwd_diff <- tab_nums(name = "tbl_cwddiff", 
                         caption = "Summary statistics for the difference in plot CWD values between visits by species. The number of species' plots (n), minimum, mean, median, and maximum CWD values are provided, along with the 5%, 25%, 75%, and 95% quantiles (q).")

tbl_psig_growth <- tab_nums(name = "tbl_psiggrowth", 
                            caption = "Monte Carlo permutation p-values for tree species growth domains. Values correspond to Manuscript Figure 3 values.  Values of zero represented no permutation results with confidence intervals equal to or greater than those observed in the data.")

tbl_psig_mort <- tab_nums(name = "tbl_psigmort", 
                            caption = "Monte Carlo permutation p-values for tree species mortality domains. Values correspond to Manuscript Figure 4 values. Values of zero represented no permutation results with confidence intervals equal to or greater than those observed in the data.")

fig_cwd_all <- fig_nums(name = "fig_cwdall", 
                        caption = "Initial CWD and Delta CWD values for all FIA plots occupied by analysis species. The green line is a linear fit of the values while the orange line is a third-order polynomial fit of the points.  ")

fig_cwd_spp <- fig_nums(name = "fig_cwdspp", 
                        caption = "Third-order polynomial fits for initial CWD and delta CWD values for FIA plots occupied by individual species.")

```

<br>

<br>

```{r cwdfig-all, fig.width=6, fig.height=6,  fig.dpi=300, fig.cap=fig_cwd_all, fig.cap.sep=""}

# Output from Nplot_Climvar_analysis.R
knitr::include_graphics(paste0(RESULTS.OTHER, "CWD_deltaCWD.png"))

```

\newpage

```{r table-cwd-lm1, tab.cap=tbl_cwd_lm1}

# Output from Nplot_Climvar_analysis.R
#ft.lm1.1 <- ft.lm1 %>%
#  compose(i = 2, j = 1, value = as_paragraph("CWD"), part = "body")

ft.lm1.1 <- ft.lm1 %>%
     mk_par(i = 2, j = 1, value = as_paragraph("CWD"), part = "body")

ft.lm1.1

```

<br>

<br>

```{r table-cwd-lm2, tab.cap=tbl_cwd_lm2}

# Output from Nplot_Climvar_analysis.R
ft.lm2 <- CWD.rels$ft_lm2 %>%
 set_formatter(
      values = function(y) formatC(y, format = "e", digits = 2),
      part = "body"
    ) %>%
  mk_par(i = 2:4, j = 1, value = c(as_paragraph("CWD"), as_paragraph("CWD^2"), as_paragraph("CWD^3")), part = "body")

ft.lm2

```

\newpage

```{r cwdfig-spp, fig.width = 6, fig.height= 6,  fig.dpi=300, fig.cap=fig_cwd_spp, fig.cap.sep=""}

# Output from Nplot_Climvar_analysis.R
knitr::include_graphics(paste0(RESULTS.OTHER, "CWD_spp_plt.png"))

```

\newpage

<!---BLOCK_LANDSCAPE_START--->

```{r table-cwd-init, tab.cap=tbl_cwd_init}

init.ft

```

\newpage

```{r table-cwd-diff, tab.cap=tbl_cwd_diff}

diff.ft

```

\newpage

```{r table-psig-growth, tab.cap=tbl_psig_growth}

psig.ft.g

```

\newpage

```{r table-psig-mortality, tab.cap=tbl_psig_mort}

psig.ft.m

```

<!---BLOCK_LANDSCAPE_STOP--->