### 2025-11-19 Manuscript.Rmd
Latest manuscript changes to RMD file

---
title: "Tree growth, mortality, and climatic stress in west-coast states, USA"
author:
  - name: "Jeremiah D. Groom"
    affiliations:
      - name: "Groom Analytics LLC"
        address: "1975 SE Crystal Lake Dr., Unit 173"
        city: "Corvallis"
        state: "OR"
        postal-code: "97333"
    orcid: "0000-0002-0110-1036"
    email: "jeremy@groomanalytics.com"
  - name: "Bryce Frank"
    affiliations:
      - name: "USDA Forest Service"
        department: "Pacific Northwest Field Station"
        address: "3625 93rd Ave SW"
        city: "Olympia"
        state: "WA"
        postal-code: "98512"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  officedown::rdocx_document:
    reference_docx: "template_use2.docx"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(kableExtra)   # Enhanced table formatting
library(scales)       # Scale functions for visualization
library(flextable)    # Professional tables for Word output
library(officedown)
library(officer)
library(rvg)
#devtools::install_github("adletaw/captioner")
library(captioner) # For figure captions.

```

```{r other_settings}
fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)
```

```{r constants}

source("Global.R")
source(paste0(CODE.LOC, "Functions.R"))

# Define any custom functions
format_p_value <- function(p) {
  case_when(
    p < 0.001 ~ "< 0.001",
    p < 0.01 ~ sprintf("= %.3f", p),
    TRUE ~ sprintf("= %.2f", p)
  )
}

set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  theme_fun = "theme_vanilla"
)

centeredP <- fp_par(text.align = "center")

```

```{r load-data}
# Some general analysis info + plots
analysis.stats <- read_rds(paste0(RESULTS.OTHER, "analysis.stats.RDS"))

# Percent of plots sampled with REMPER = 9, 10, and 11 years (rounded)
pct.remeas <- as.vector(analysis.stats$pct.remeas)

# Species scientific names
spp.names.use <- analysis.stats$gym.angio.dat

# Correlation/fits for Initial CWD vs Delta CWD:
CWD.rels <- read_rds(paste0(RESULTS.OTHER, "CWD_deltaCWD_relationship.rds"))
CWD.lin.cor <- format(round(CWD.rels$lm1$r.squared^0.5, 2), nsmall = 2) # Last digit = 0, want to show.
CWD.dAIC <- round(CWD.rels$aic.cwd$AIC[1] - CWD.rels$aic.cwd$AIC[2], 0)

# Growth data for species overall, states, differences
grow.state <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/State_Growth_", CLIM.VAR.USE, ".csv"))
grow.all <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/All_Growth_", CLIM.VAR.USE, ".csv"))
grow.diff <- read_rds(paste0(save.loc.fcn(1), "Processed_6Domain_Data.RDS"))

# Mortality data for species overall and states
mort.state <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/State_Mortality_", CLIM.VAR.USE, ".csv"))
mort.all <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/All_Mortality_", CLIM.VAR.USE, ".csv"))
mort.diff <- read_rds(paste0(save.loc.fcn(2), "Processed_6Domain_Data.RDS"))

# Significant findings, Growth
grow.n.AvB <- nrow(grow.diff$AvB2)
grow.n.A <- nrow(grow.diff$A_LMH2)
grow.n.B <- nrow(grow.diff$B_LMH2)

grow.AvB.sig <- length(grow.diff$AvB2$significant[grow.diff$AvB2$significant == TRUE])
grow.AvB.signeg <- length(grow.diff$AvB2$significant[grow.diff$AvB2$significant == TRUE & grow.diff$AvB2$Means < 0])
grow.A.sig <- length(grow.diff$A_LMH2$significant[grow.diff$A_LMH2$significant == TRUE])
grow.A.signeg <- length(grow.diff$A_LMH2$significant[grow.diff$A_LMH2$significant == TRUE & grow.diff$A_LMH2$Means < 0])
grow.B.sig <- length(grow.diff$B_LMH2$significant[grow.diff$B_LMH2$significant == TRUE])
grow.B.signeg <- length(grow.diff$B_LMH2$significant[grow.diff$B_LMH2$significant == TRUE & grow.diff$B_LMH2$Means < 0])

grow.tot.sig <- grow.AvB.sig + grow.A.sig + grow.B.sig
grow.tot <- grow.n.AvB + grow.n.A + grow.n.B

# Significant findings, Mort
mort.n.AvB <- nrow(mort.diff$AvB2)
mort.n.A <- nrow(mort.diff$A_LMH2)
mort.n.B <- nrow(mort.diff$B_LMH2)

mort.AvB.sig <- length(mort.diff$AvB2$significant[mort.diff$AvB2$significant == TRUE])
mort.AvB.signeg <- length(mort.diff$AvB2$significant[mort.diff$AvB2$significant == TRUE & mort.diff$AvB2$Means < 0])
mort.A.sig <- length(mort.diff$A_LMH2$significant[mort.diff$A_LMH2$significant == TRUE])
mort.A.signeg <- length(mort.diff$A_LMH2$significant[mort.diff$A_LMH2$significant == TRUE & mort.diff$A_LMH2$Means < 0])
mort.B.sig <- length(mort.diff$B_LMH2$significant[mort.diff$B_LMH2$significant == TRUE])
mort.B.signeg <- length(mort.diff$B_LMH2$significant[mort.diff$B_LMH2$significant == TRUE & mort.diff$B_LMH2$Means < 0])

mort.tot.sig <- mort.AvB.sig + mort.A.sig + mort.B.sig
mort.tot <- mort.n.AvB + mort.n.A + mort.n.B

# Summary of significant Growth and Mortality findings for table creation.
datasets.grow <- list(grow.diff$AvB2, grow.diff$A_LMH2, grow.diff$B_LMH2)
datasets.mort <- list(mort.diff$AvB2, mort.diff$A_LMH2, mort.diff$B_LMH2)
datatypes <- list("Above vs. Below", "Above", "Below")
n.comp.grow <- list(grow.n.AvB, grow.n.A, grow.n.B)
n.comp.mort <- list(mort.n.AvB, mort.n.A, mort.n.B)

growth.summ <- pmap(list(datasets.grow, datatypes, n.comp.grow), summ.fig.fcn) %>% bind_rows()
mortality.summ <- pmap(list(datasets.mort, datatypes, n.comp.mort), summ.fig.fcn) %>% bind_rows()

```

```{r table-creation}
# State growth
grow.state.cm <- cm2.fcn(1, grow.state)
grow.all.cm <- cm2.fcn(1, grow.all)
s.g.dat <- state.table.fcn(grow.state.cm, grow.all.cm, 1)

growth.ft <- flextable(s.g.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
    italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = TRUE)
min.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = TRUE)

# State mortality
s.m.dat <- state.table.fcn(mort.state, mort.all, 2)

mort.ft <- flextable(s.m.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",  
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = FALSE, digits = 2)
min.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = FALSE, digits = 2)

# CWD amounts
min.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.min.mean]
min.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.min.mean]
max.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.max.mean]
max.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.max.mean]

max.q95.diff <- analysis.stats$diff.cwd.stats$q95[analysis.stats$diff.cwd.95max]
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]
max.range.diff <- analysis.stats$diff.cwd.90range.max
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]

# Adding a section identifier to each table
growth.summ2 <- growth.summ %>%
  mutate(Type = "Growth", .before = 1)

mortality.summ2 <- mortality.summ %>%
  mutate(Type = "Mortality", .before = 1)

# Combine the tables
grow.mort.summ <- bind_rows(growth.summ2, mortality.summ2)

# Create flextable
g.m.summ.ft <- flextable(grow.mort.summ) %>%
  merge_v(j = "Type") %>%  # Merge cells in Section column
  valign(j = "Type", valign = "top") %>%  # Top align section labels
  flextable::align(j = "Type", align = "left") %>%   # Left align section labels
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  set_table_properties(layout = "autofit", width = 0) %>%
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
    flextable::align(j = 3:7, align = "center", part = "all")

```

```{r set_captions}

# 
fig_nums <- captioner::captioner(prefix = "")
tab_nums <- captioner::captioner(prefix = "Table")

f.ref <- function(x) {
 caption_text <- fig_nums(x)
  # Remove any leading numbers and extract just "Figure X"
  stringr::str_extract(caption_text, "\\d+")
}

t.ref <- function(x)  {
  stringr::str_extract(tab_nums(x), "[^:]*")
}

tab.tbl_growth <- tab_nums(name = "tbl_growth", 
                        caption = "Decadal growth estimates by state and overall for analysis species (cm\u00B2/decade). Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots.")

tab.tbl_mort <- tab_nums(name = "tbl_mort", 
                        caption = "Decadal mortality estimates by state and overall for analysis species. Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots.")

fig.cwdmap_cap <- fig_nums(name = "fig_cwdmap", 
                        caption = "Fuzzed FIA plot locations displayed with initial visit CWD values (A) and change in CWD between visits (C) for the U.S. states, north to south, of Washington, Oregon, and California (see inset map B for the location of these states within the USA).")

fig.domains <- fig_nums(name = "fig_domains",
                        caption = "This figure illustrates our bivariate domain approach using *Calocedrus decurrens* as an example. Here we depict decadal (A) growth and (B) mortality domain estimates.  The six domains are defined as FIA plots that fall [A]bove or [B]elow the CWD change threshold with [L]ow, [M]edium, or [H]igh initial CWD values (see text for details). E.g., 'AL' is Above CWD change threshold, Low initial CWD value. The distribution of plot CWD change and initial values (C) is presented as well, with the blue horizontal line indicating the CWD change threshold and the orange vertical lines representing the quantile boundaries between Low, Medium, and High initial CWD values." )

fig.diff_grow <- fig_nums(name = "fig_diffgrow", 
                          caption = paste0("Differences in mean growth rates (cm\u00B2/decade) between domains by species.  Domain abbreviations are described in the legend for Figure ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are coniferous; those below are deciduous. Empirical MC permutation p-values are provided for filled-circle contrasts, with \u2013\u2013 corresponding to p-values of > 0.05, while one to four asterisks correspond to p > 0.01, > 0.001, > 0.0001 and < 0.0001, respectively."))

fig.diff_mort <- fig_nums(name = "fig_diffmort", 
                          caption = paste0("Differences in mean decadal mortality rates between domains by species.  Domain abbreviations are described in the legend for Figure ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are coniferous; those below are deciduous.  Empirical MC permutation p-values are provided for filled-circle contrasts, with \u2013\u2013 corresponding to p-values of > 0.05 while one to four asterisks correspond to p > 0.01, > 0.001, > 0.0001 and < 0.0001, respectively. "))

tab.summ_grow_mort <- tab_nums(name = "tbl_summ_grow_mort",
                               caption = paste0("Summary of number of contrasts with 95% CIs that do not overlap with zero ([T]otal, [P]ositive-, and [N]egative-valued estimates) for Figures ", fig_nums("fig_diffgrow", display = "num"), " and ", fig_nums("fig_diffmort", display = "num"), " by analysis (T)ype (Growth or Mortality) and type of (C)omparison.  The column Domains lists the domains compared; All is the sum of values within a Comparison. See text for other domain labels.  'n' is the total number of differences examined within a Comparison."))

```

# Abstract

Western North American forests are experiencing accelerating tree mortality rates in response to intensifying drought and water stress driven by climate change.  Tree mortality and growth reflect cumulative effects of direct physiological stress (xylem cavitation, carbon starvation), and indirect climate-mediated impacts (wildfire, insects, pathogens).  However, obtaining design-based estimates for these demographic rates requires addressing a methodological challenge: revisit data are necessary to measure tree fate and growth, but plots differ in their initial water stress conditions and the magnitude of change between visits, necessitating a bivariate rather than univariate estimation approach. We developed a bivariate design-based estimation framework that stratifies plots into domains defined by both baseline initial-visit climatic water deficit (CWD) and change in CWD between visits.  We applied this approach to `r comma(analysis.stats$plot.num)` National Forest Inventory plots spanning California, Oregon, and Washington, tracking `r comma(analysis.stats$tree.num)` trees from `r length(sel.spp)` species over decade-long intervals.  We estimated decadal growth and mortality overall, by state, and across six CWD domains per species, using bootstrap confidence intervals and Monte Carlo permutation tests to assess statistical significance.  Species responses were highly variable, with no universal pattern of increased mortality or reduced growth in domains experiencing elevated water stress. Several species exhibited overall mortality rates exceeding 0.20 per decade (20% decadal mortality), with state estimates of 0.34 for *Picea engelmannii* in Washington, 0.27 for *Acer macrophyllum* in California, and 0.37 for *Pinus contorta* in Washington. Domain-specific patterns aligned with known regional disturbances (bark beetles, drought, fire) for multiple species. Our bivariate framework provides a template for monitoring forest vital rates as ecosystem-level indicators of climate change response, while our species-specific findings identify or validate populations warranting immediate conservation assessment and inform forest management strategies.

# Introduction

Western North American forests are experiencing increases in tree mortality across multiple species in concert with changing climate conditions.  Species and tree groups have incurred extraordinary losses in specific regions.  Johnston et al. (2025) documented a 29% 10-year mortality rate of 300+ yr old trees in Eastern Oregon.  High-severity fires in California in 2020 and 2021 killed 13%-19% of extant giant sequoia (*Sequoiadendron gianteum*) with >1.07-m bole diameters (Stephenson and Brigham 2021, Shive et al. 2022); recruitment in these burns was depressed (Soderberg et al. 2022).  Species and genera have also seen their mortality rates increase overall. van Mantgem et al. (2009) modeled mortality rates for the genera *Abies* and *Pinus* in the western USA as increasing from less than 0.05% annually in 1960 to almost 0.2% by 2010. Perret et al. (2023) estimated a decadal range-wide mortality rate of 25% for Subalpine fir (*Picea engelmannii*).    Five of eight common tree species in western North America exhibit declines as measured by relative stand density (Stanke et al. 2021).  Shriver et al. (2022) found that four of five juniper (*Juniperus* spp.) and piñon pine (*Pinus* spp.) species were declining in portions of their range, comparing mortality to recruitment rates.

Climate change as manifest through increases or intensification in drought and aridity can affect tree mortality directly and indirectly.  Direct impacts include physiological damage while indirect causes include increasing wildfire, insect, and disease outbreak frequency and severity.  Hammond et al. (2022) found globally that the intensity of drought and heat events are directly associated with widespread forest die-off events.  Wildfire frequency and area burned in the western United States has increased in concert with the longer fire seasons, hotter summer temperatures, and earlier snowmelt (Abatzoglou and Williams 2016, Westerlin 2016, Juang et al. 2022).  There has also been a regional increase in the area of bark beetle (subfamily *Scolytinae* spp.) outbreaks (Meddens et al. 2012, Hicke et al. 2016).  Many of these indirect climate-influenced mortality causes may be attributable to the physical condition of trees following abnormal or prolonged exposure to water stress.

Carbon starvation and hydraulic failure can drive drought-induced mortality (see reviews by Adams et al. 2010, Sala et al. 2010, 2012, and McDowell et al. 2008).  Carbon starvation occurs when isohydric species (e.g., piñon pine, *Pinus edulus*; Breshears et al. 2009), or those which mainly rely on stomatal closure to avoid excessive water loss through evapotranspiration (ET), receive insufficient atmospheric carbon dioxide for photosynthesis.  Anisohydric plants, or those who do not strongly regulate water loss through stomatal closure, risk hydraulic failure if sufficient xylem cavitation occurs due to insufficient water uptake needed to meet ET demands (e.g., as seen in *Populus tremuloides*; Anderegg et al. 2015). Adams et al. (2017) performed a review of drought-induced mortality across 26 species and concluded that hydraulic failure occurred broadly across taxa while non-structural carbohydrate reserves at mortality differed across taxa.  Regardless of mechanism, drought (specifically, water deficit) appears to be an important direct driver of mortality for tree taxa (Hammond et al. 2022).

Drought stress can increase tree susceptibility to wildfire, insects, and disease.  Wildfire area (Abatzoglou and Williams, 2016; Parks et al. 2025) and number (Westerling 2016) are positively associated with increasing aridity in the western US.   Bark beetle outbreaks in semi-arid areas of the western US may be positively affected by or associated with acute intense drought and water-stressed trees (Negrón 2009, Gaylord et al. 2015) but chronically stressed trees may not provide bark beetle populations with the resources sufficient to promote population expansion (Huberty and Denno, 2004).  Norway spruce (*Picea abies*) saplings inoculated with the necrotrophic pathogenic fungi *Heterobasidion annosum* sensu stricto and exposed to drought conditions were more likely to die than control saplings or saplings exposed to either drought conditions or inoculation, although Scots pine (*Pinus sylvestris*) seedlings did not show a similar relationship (Gomez-Gallego et al. 2022). Other insects and disease may be promoted or inhibited by drought, depending on the circumstances and due to conflicting factors (Kolb et al. 2016).

Given the importance of water stress on tree mortality, selecting an appropriate metric is critical for rangewide assessments.  One regionally useful metric of water stress is Climatic Water Deficit (CWD; Stephenson 1998).  Stephenson (1990) defines CWD as evaporative demand not met by available water, or the difference between plant water demand (actual evapotranspiration, AET), and plant water supply (potential evapotranspiration); these variables are driven by temperature and precipitation, respectively.  The Anderegg  et al. (2015) tree mortality study relied on CWD to capture drought occurrence.  Das et al. (2013) determined that low-elevation California tree mortality was best modeled using CWD.  Restaino et al. (2016) found decreased Douglass fir growth with an increase in CWD.

Data from the U.S. National Forest Inventory, collected by the US Forest Service Forest Inventory and Analysis Program (FIA), is well suited for evaluating regional species mortality and growth.   Since 1999 the FIA program has gathered a variety of forestry data at plots nationwide using a spatially balanced probability sample (McRoberts 2005, Reams et al. 2005, Woudenberg et al. 2010).  The sample allows for relatively unbiased design-based inference of population characteristics across a large geographical region.  Data from plot revisits and individual tree data permit direct estimation of change features such as mortality and growth estimation.

Climatic data such as mean annual temperature and CWD can be combined with FIA data to obtain climate-related design-based estimates. Monleon and Linz (2015) evaluated the evidence of tree species range shifts by subtracting the design-based mean estimates of FIA plot temperatures for seedlings from those of trees. FIA plot revisit data were mostly incomplete at the time of their analysis.  Groom and Monleon (2023) used a mostly-completed revisit data set to compare plot climate variable values (vapor pressure deficit [VPD], temperature, and precipitation) to estimate shifts in plot occupancy and density.  Both studies used a single climate variable value associated with each plot for estimation; e.g., Groom and Monleon (2023) examined shifts in species density by estimating the difference in VPD means between plots that gained trees of a given species and plots that lost trees.

Groom et al. (2023) documented, however, that the initial climate variable values could (and often did) change between plot visits, with species' plots often exhibiting positive or negative correlations between the initial visit values and between-visits change value of the climate variable.  They identified that their design-based analysis would potentially improve if they could simultaneously include both initial visit and change values for the climate variable.

Our study has two objectives.  The first is to demonstrate, across 23 species, a bivariate design-based estimation approach for tree growth and mortality that simultaneously accounts for initial climatic conditions and change between measurements.  This addresses a gap in design-based analyses of forest change.  Second, we provide current growth and mortality estimates for our analysis species across our study area and by state to establish species status and to inform our climate analysis findings.

# Methods

## FIA data

Our analysis relies on FIA data for forestland in California, Oregon, and Washington, USA.  Forestland is defined as an area of 4,050 m$^2$ that is at least 10 percent stocked or potentially stocked with tree species, excluding urban and agricultural land uses (Reams et al. 2005).  We constrained our analysis to only include remeasurements for areas considered forestland during both visits.

FIA relies on three related sampling procedures.  Phase 1 involves the classification of distributed points of remote sensing data into (at least) the categories of forest and non-forest.  Phase 1 plots are useful for estimation stratification.  The phase 2 permanent ground plots are a subset of Phase 1 points.  They consist of four points where all trees with stem diameters greater than or equal to 12.7 cm (for most species measured at 1.37 m above the ground) were tallied in a 7.32 m radius circular subplot (total area 672.5 m$^2$) with additional trees sampled in an annulus around each subplot (macroplot), with a total radius 18 m from the center point and a total area of 1012 m$^2$.  We did not include data from microplots, where seedlings and saplings are measured, as the analysis was not intended to estimate species recruitment but rather to follow the trajectory of established trees. Phase 3 plots are an intensification of Phase 2 plots; we do not rely on their data here. Reams et al. (2005) and Bechtold and Scott (2005) provide plot design details and measurement protocols.

The FIA program selected plots with a spatially-balanced probability sampling design (McRoberts 2005, Reams et al., 2005) that used a target plot density of one ground plot every 24 km$^2$. Some regions and ownerships intensify their plot density; we excluded these additional plots to simplify stratification estimation, relying on FIA-defined strata and using the density of FIA Phase 1 sampling points within and across strata to determine stratum weights.

Plots in the three states are sampled every 10 years with a target of annually sampling a panel of 10% of plots every year (Thompson 2015). Our analysis depends on examining the fate and growth of trees measured during two sequential visits (referred to as the first and second plot visit), ideally from a complete sample panel remeasure. Tree growth is the observed increase in marked tree bole diameter between visits (cm$^2$) while mortality is captured from the change in tree condition. We obtained our FIA data from the USDA FIA DataMart web portal (https://research.fs.usda.gov/products/dataandtools/fia-datamart; downloaded May 5, 2025).  The database included plot data sampled from 2001 to 2023.

Individual trees within plots were retained in the analysis if they were alive at the initial visit and met the stem diameter criterion during that visit (i.e., plot ingrowth was not included).  The growth analysis additionally restricted trees considered to those that were alive at both visits.  The mortality analysis excluded trees killed through silvicultural practices.  Both analyses excluded trees that shrank in diameter, physically moved due to landslides or other event, were erroneously tallied at either visit, were missed due to data collection procedural changes, or existed at any time in a nonsampled area.  We used only FIA plots sampled specifically for tree mortality and growth.

## Climate data

We obtained plot-level CWD data from the model TerraClimate (Abatzoglou et al. 2018) for June, July, and August (summer months) between 1980 and 2023. Latitude and longitude were based on FIA-fuzzed coordinates.  We defined the 'initial CWD' values as the 20-year mean of summer CWD values for each plot prior to and including the first visit year (initial CWD = mean(summer CWD from visit year$_{-19}$ to year$_0$).  We then found the mean of summer CWD values between visits and including the second visit year (second visit CWD = mean(summer CWD from year$_1$ to year$_10$) .  We then defined delta CWD = second visit CWD - initial CWD.  Figure 1 presents maps of both the initial CWD and delta CWD values for forested and nonforested FIA plots.

## Analysis
We obtained point estimates of species growth and mortality using ratio estimators (Eqn 4.16 in Scott et al. 2005).  We expanded the number of sampled trees in each plot by the area sampled (subplot or macroplot area) adjusted by the proportion of the plot area available for sampling and the number of years between visits. The numerator for species-specific mortality is the sum of stratum-weighted means of the number of trees that died between visits per plot.  The denominator for mortality is similar but is calculated as the sum of stratum-weighted means of all trees per plot of a given species.

The numerator of species-specific growth estimates is the sum of stratum-weighted means of the increase in basal area (cm$^2$) between visits per plot while the denominator is the same as was used for the mortality estimate. This produces a point estimate for a species’ mean growth.  We standardized mortality and growth values to a decadal time period.

We estimated, for each species examined, six domains based on CWD values.  The domains allow us to parse the plot CWD values into three levels for initial CWD values and two levels for delta CWD values.  We created the domains based on initial CWD plot values by dividing the values into three species-specific quantiles: values below the first quartile (Low), between the first and third quartiles (Medium), and above the third quartile (High; Figure 2C).  We defined the two domains for delta CWD plot values for all species as those above a 5mm CWD change threshold (Above) and those below the threshold (Below). The 5mm threshold was selected because species' delta CWD value distributions did not differ greatly and because 5mm is approximately the mean delta CWD value across species (Table S2).  The Below domains may includes negative delta CWD values (i.e., mean CWD was larger prior to the initial visit than between visits).  The domain designations AL, AM, and AH correspond to Above Low, Above Medium, and Above High. The Below domains are similarly labeled (BL, BM, BH). Parsing the plot data into six domains allow for bivariate examinations of each species' data.  This bivariate approach allows us to ask, for example, whether trees in historically dry sites (High initial CWD) respond differently to increased water stress (Above threshold delta CWD) compared to trees in mesic sites (Low initial CWD) experiencing similar increases. We calculated estimates for all CWD-based domains with 10 or more plots.

We additionally report species mortality and growth estimates for all plots across and within Oregon, Washington, and California irrespective of CWD values.  We limited the number of species examined to those with more than 400 occupied plots across the three states.

For each species we examined contrasts among growth (Figure `r fig_nums("fig_diffgrow", display = "num")`) and mortality (Figure `r fig_nums("fig_diffmort", display = "num")`) CWD domain estimates.  The figure column "Above vs. Below Threshold" displays the mean point estimate and 95% CI for contrast between corresponding Above and Below domains (e.g., for *Abies lasiocarpa*, subtracting the estimates for columns BL, BM, and BH from AL, AM, AH, respectively; Figure `r fig_nums("fig_domains", display = "num")`A). Negative estimates for the "Above vs. Below Threshold" in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")`  indicate that the 10-year growth or mortality rate was greater for a species in plots below the change estimates.  `r t.ref("tbl_summ_grow_mort")` summarizes the number of domain differences in each of the columns in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")` where the 95% CI does not include zero (termed "significant" findings) as well as the number of contrasts per figure column.

We obtained empirical means and 95% confidence intervals via bootstrap. For each of the 1000 bootstrap iterations we sampled FIA plots with replacement and obtained point estimates for each species' domain.  For each domain we calculated a mean, upper, and lower 95% CI with the 2.5% and 97.5% quantiles from the  sampling iterations.  This approach allows for nonsymmetrical CIs around means.  We estimated contrasts between pairs of CWD domains by subtracting within-iteration domain point estimates.

Our analysis approach results in hundreds of comparisons.  We conducted a Monte Carlo (MC) permutation analysis to provide species-level false positive rates (p-values) for "significant" contrasts. For each of 1000 MC permutation analysis iterations the code estimated each all species' bootstrap confidence intervals as described above. The MC permutation analysis sampled each species' plot domain assignments without replacement.  Given this sampling approach, and since species distributions were not independent, the p-values are strictly interpretable at the species level (e.g., we cannot speak to the likelihood of observing a specific number of significant findings for a contrast type). The distance of MC permutation contrast CI bounds from zero were compared against test CI bounds. The MC permutation procedure recorded a null significant value if the absolute values of the permutation iteration CI bounds were at least as far from zero as observed bounds.

Given the large number of contrasts conducted (`r mort.tot` each for growth and mortality), we adopted a conservative approach to interpretation. Aside from discussing patterns of findings across species, we focus discussion on contrasts where (1) the 95% CI does not include zero, and (2) the MC permutation test yields p < 0.001, indicating that the observed pattern is unlikely under random domain assignment.  Contrasts with less stringent p-values are shown in figures but are generally not discussed individually as they may represent Type I errors.

## Dashboard
We developed a data dashboard (https://tinyurl.com/FIAgrowmort) to enable ourselves and readers to examine individual species domain estimates, the distribution of initial and delta CWD values, and the geographical distribution of plots and plot domains.  All analysis files, manuscript markdowns, dashboard code, and ancillary data sources are available at the GitHub repository https://github.com/jeremygroom/Tree_Growth_Death.

# Results

Our study tracked the fate of `r comma(analysis.stats$tree.num)` trees from `r length(sel.spp)` species in `r comma(analysis.stats$plot.num)` plots that were visited twice by FIA crews. Of the `r length(sel.spp)` species, `r analysis.stats$n.gym.spp` were gymnosperms.  The two visits were 10 years apart for `r pct.remeas[2]`% of plots (`r pct.remeas[1]`% were 9 years, `r pct.remeas[3]`% were 11 years). Plots were distributed across `r analysis.stats$strat.num` strata spanning forested regions in all three states, from coastal temperate rainforest to semi-arid interior woodlands.

Mean overall growth rates for species varied between `r min.grow.spp$mean` (*`r min.grow.spp$spp`*) and `r max.grow.spp$mean` (*`r max.grow.spp$spp`*, `r t.ref("tbl_growth")`) cm$^2$/decade.  Six species had greater growth (non-overlapping CIs) in California than in Oregon and/or Washington: *Abies concolor*, *Abies grandis*, *Pinus ponderosa*, *Tsuga heterophylla*, *Arbutus menziesii*, and *Notolithocarpus densiflorus*.  Three species had greater growth in Oregon (*Calocedrus decurrens*, *Picea engelmanii*, and *Pseudotsuga menziesii*) and two species had greater growth in Washington (*Abies amablis* and *Pinus monticola*).

Decadal mean mortality rates varied between `r min.mort.spp$mean` (*`r min.mort.spp$spp`*) and `r max.mort.spp$mean` (*`r max.mort.spp$spp`*, `r t.ref("tbl_mort")`). Four species in California exhibited greater mortality rates than in Oregon and/or Washington: *Abies concolor*, *Pinus ponderosa*, *Acer macrophyllum*, and *Quercus kelloggii*.  One species had a greater mortality rate in Oregon (*Tsuga mertensiana*) and Washington had five (*Abies grandis*, *Larix occidentalis*, *Picea engelmannii*, *Pinus contorta*, and *Pseudotsuga menziesii*).  Seven species had overall mortality rates $\geq$ 0.20 (*Abies concolor*, *Abies lasiocarpa*, *Picea engelmannii*, *Pinus contorta*, *Pinus lambertiana*, *Alnus rubra*, and *Quercus chryselopis*) while at a state level seven species had mortality rates $\geq$ 0.25 (California: *Acer macrophyllum* and *Alnus rubra*; Oregon: *Abies lasiocarpa*; Washington: *Abies grandis*, *Picea engelmannii*, *Pinus contorta*, and *Pinus monticola*).

Delta CWD increased at `r analysis.stats$cwd.gt0.pct`% of plots. Initial and delta CWD values were positively linearly associated (correlation = `r CWD.lin.cor`); however, the two variables were better modeled by a 3rd order polynomial ($\Delta AIC$ =  `r CWD.dAIC`, Figure S1, see model fits in Tables S1 and S2), and for most species delta CWD neither linearly related to nor monotonically increased with initial CWD values (Figure S2; also, see dashboard).  Tree species exhibited a variety of initial-visit CWD ranges, with means ranging from `r min.mean.init` (*`r min.mean.init.spp`*) to `r max.mean.init` (*`r max.mean.init.spp`*; Table S3).  Delta CWD values tended to be comparatively small, with a maximum 95th percentile value of `r max.q95.diff` (*`r max.q95.diff.spp`*, Table S4).

Individual species growth and mortality domain patterns could differ (e.g., Figures `r fig_nums("fig_domains", display = "num")`A and `r fig_nums("fig_domains", display = "num")`B). Growth domain contrasts (Figure `r fig_nums("fig_diffgrow", display = "num")`, `r t.ref("tbl_summ_grow_mort")`) did not appear to result in substantially more "significant" negative or positive estimates, indicating that species overall did not exhibit more or less growth in the Above, Low, or High domains.

All three columns of Figure `r fig_nums("fig_diffgrow", display = "num")` had over 20 "significant" contrasts. Depending on the species, growth may have been susceptible (negatively or positively) to a change in CWD and/or differed by initial CWD values. *Abies amabilis*, *Abies grandis*, *Calocedrus decurrens*, *Pinus ponderosa*, *Pseudotsuga menziesii*, *Tsuga mertensiana*, *Acer macrophyllum*, and *Alnus rubra* exhibited greater growth in the Low domains than the Medium and/or High domains.  Except for *Alnus rubra*, these results predominantly occurred in both the Below domain (2 spp.) or the Above and Below domains (5 spp.).  *Thuja plicata* had greater growth in the Above and Below Medium domains than either High or Low domains. *Abies concolor*, *Larix occidentalis*, *Picea engelmannii*, *Pinus contorta*, *Pinus lambertiana*, *Arbutus menziesii*, and *Quercus kelloggii* had greater growth in the more water stressed High domains than the Medium and/or Low domains. Except for *Larix occidentalis*, these results predominantly occurred in both the Above domain (3 spp.) or the Above and Below domains (3 spp.). Other species exhibited no 'significant' findings (4 spp.) or displayed other result patterns (3 spp.).

Mortality domain contrasts are similar in certain regards to the growth contrasts: results for the three types of domain contrasts did not appear to be predominantly positive or negative (Figure `r fig_nums("fig_diffmort", display = "num")`, `r t.ref("tbl_summ_grow_mort")`).  There were fewer mortality estimates with "significant" findings than growth (`r mort.tot.sig` vs `r grow.tot.sig`). There were fewer differences for "Above vs. Below Threshold" contrasts (`r mort.AvB.sig`, `r t.ref("tbl_summ_grow_mort")`) than for Above (`r mort.A.sig`) or Below (`r mort.B.sig`), indicating that mortality may not be as strongly associated with delta CWD as by the initial CWD value.

*Larix occidentalis*, *Picea engelmannii*,	*Pinus contorta*,	*Pinus monticola*, and *Thuja plicata* had greater mortality in the lower water deficit Low domains than Medium and/or High domains. These results occurred in Above (3 spp.) or both Above and Below (2 spp.) domains. Mortality rates for *Abies lasiocarpa*,	*Calocedrus decurrens*,	*Pinus lambertiana*, *Pinus ponderosa*,	*Pseudotsuga menziesii*,	*Tsuga heterophylla*,	*Tsuga mertensiana*,	*Acer macrophyllum*,	and *Notolithocarpus densiflorus* were greater in the High domains than the Medium and/or Low domains.  These species were approximately evenly split between differences being present in the Above (4 spp.) or Below (5 spp.) domains, but none with "significant" findings in both the Above and Below domains. *Arbutus menziesii* had greater mortality in Medium domains than the High domains. Five species had no significant contrasts.

The interactive dashboard revealed geographic patterns that were associated with specific domains while not necessarily captured by state-level estimates.  For the Above and Below Low domains *Abies lasiocarpa* had less mortality in northeastern Washington and *Tsuga mertensiana* exhibited greater growth in the same region.  *Pinus contorta* exhibited less growth and greater mortality in northern Washington (Above Low).  In the southern Sierra Nevada *Pinus lambertiana* had greater growth (Above Medium and High) and greater mortality (Above High), *Calocedrus decurrens* had greater mortality (Above High), and *Pinus monticola* had less mortality (Above High). *Acer macrophyllum* exhibited less growth in the Sierra Nevada and southern Klamath Mountains (Above High) and greater growth in northwestern Washington in the Coast and Cascade ranges (Above and Below Low).

The MC permutation analysis provided empirical p-values for observing contrast confidence intervals with bounds at least as far from zero as those observed (Supplemental figures S5 and S6), with values ranging from a maximum of $p$ = 0.072 to a minimum $p$ < 0.0001 (41 instances).  Except where explicitly stated, we only discuss individual species' domain findings if the species has one or more relevant MC permutation p-values < 0.001 (indicated by three or four asterisks in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")`).

# Discussion

Analysis of 23 tree species revealed diverse, species-specific associations with climatic water deficit, with different species exhibiting contrasting relationships among growth, mortality, and water stress. Responses were potentially driven by differences in drought tolerance strategies, local disturbance regimes, and biotic interactions. We found more “significant” CWD contrasts for species growth than mortality, possibly indicating that growth differences by CWD domain were more readily detectable than mortality over a decade, although growth and mortality findings were not necessarily negatively related.  Delta CWD increased almost everywhere between plot revisits although the magnitude of change was generally not large relative to initial CWD values.  We additionally found concerningly high state and/or overall decadal mortality rates for some species.

Contrast results may be more informative at an individual species level and when considering state-level mortality and growth estimates.  Regional analyses detected several species with relatively high mortality rates including *Acer macrophyllum*, *Picea engelmannii*, *Abies lasiocarpa*, *Pinus contorta*, and *Alnus rubra*.  We do not believe our results are concerning for *Alnus rubra*.  Our estimates for this short-lived tree species align with reported background mortality rates (e.g., Brown and Courtin 2018).   However, the mortality rate for *Acer macrophyllum* in California appears to be both unreported elsewhere and substantially higher in California than in Oregon or Washington.

*Acer macrophyllum* growth was negatively associated with initial CWD values, and its growth was lower and mortality almost three times higher in California than in Washington and Oregon.  We cannot find primary literature that corroborates an increase in mortality for this species in California. McNellis et al. (2021) estimated a recent *Acer macrophyllum* overall mortality rate of 0.45%/yr (4.5%/decade) while our estimates for Washington and Oregon were twice this level, and California's estimate was six times as large.  However, Betzen et al. (2021) found widespread evidence in Washington of *Acer macrophyllum* decline (crown dieback, thinning, discoloration, reduced leaf size) and reduced growth and increased mortality in locations with higher temperatures and water limitations.  Although these results align with our contrast findings for growth, as displayed in our dashboard, they are not reflective (yet) of our state mortality findings for Washington.  We encourage a more intensive assessment of *Acer macrophyllum*’s status in California.

Conspecifics *Abies lasiocarpa* and *Picea engelmannii* are frequently codominant and exhibit high mortality rates. Andrus et al. (2018) found that *Picea engelmannii* exhibited higher survival and longevity compared to *Abies lasiocarpa* in undisturbed plots in Colorado, with firs exhibiting a mortality rates of up to 1% annually while *Picea engelmannii* had mortality rates below 0.5% annually.  Mortality estimates for *Abies lasiocarpa* from Perret et al. (2023) approximated our state and overall estimates at 25% +/-2.7; they found the highest level of mortality in the eastern Oregon Cascades as well as in Idaho and Colorado. They attributed 83% of the mortality to fire or biogenic disturbance.  We found that the pressures on the two species appear to differ by CWD values and location.  We found that *Picea engelmannii* had lower mortality in the Above High domain, corresponding to stands in the Blue Mountains of Oregon and Washington, and higher mortality in the Above Low and Above High domains, largely coinciding with northern and northeastern Washington, where spruce beetle infestations were elevated (WADNR 2023).  Its mortality rate in Washington was almost 5 times higher than in Oregon.  *Abies lasiocarpa* did not have substantial mortality contrast findings (some evidence of BH - BL difference, but the MC permutation p-value was greater than 0.001) and its mortality was high but similar across all three states.

Our findings for *Pinus contorta* in Washington likely reflected insect pressures on the species.  Our Washington mortality estimate was twice that of Oregon’s and five times greater than Calfornia’s, while the growth confidence intervals for Washington were lower than California’s and did not overlap. The domains with the highest mortality and lowest growth align well with Washington regions that have experienced water stress (Raymond et al. 2023).  Pine bark beetles (*Dendroctonus ponderosae*, *Dendroctonus brevicomis*, and *Ips* spp.) have affected *Pinus contorta* for the past decade in this region (WADNR 2023).  This area also corresponds, at the level of ecoregion subsections, with the portions of *Pinus contorta*’s range in the lowest forest stability index category for Stanke et al. (2021) while California’s Sierra Nevada and the Oregon and Washington Cascade Range correspond to positive stability index values.  Stanke et al. also found, from FIA plot data, that insects (followed by fire) were related to the percent change in relative tree density.

The absence of consistent contrast patterns across species is itself informative, indicating that tree responses to water stress conditions are potentially mediated by local adaptation to conditions, species-specific physiological traits, and local environmental conditions rather than following a general response curve.  The number of "significant" Above vs. Below threshold contrasts for growth and mortality were approximately evenly distributed between positive and negative contrast estimates, with a positive difference indicating that some species exhibited higher growth or greater mortality in an Above domain (e.g., *Pinus lambertiana* for both growth and mortality) versus a Below domain (*Abies grandis* and *Pinus contorta*, respectively, for growth and mortality).  This lack of a pattern across species may indicate that the distinction of Above vs. Below is not generalizable across species, or that "significant" findings are due to confounding factors such as geographical locations with different ecological pressures.

*Abies grandis* exhibited higher mortality in Washington than Oregon or California. We found no evidence of a difference in domain mortality geographically, although the greatest growth domain was Below Low, which corresponded to the more mesic plots west of the Cascade Crest of Washington and Oregon.  The mortality result may be due to an intersection in a fire suppression history promoting a densification of shade-tolerant but drought-intolerant firs in the arid climates of eastern Oregon and Washington, drought, and fir engraver beetle mortality (*Scolytus ventralis*; Washington DNR 2023, Merschel et al. 2021).  Johnson et al. (2025) report mortality rates of *Abies grandis* in protected areas in the Malheur National Forest of Oregon that exceed ours, primarily attributing the cause to fire.

Our dashboard reveals some instances of domains representing geographical mortality differences.  State estimates for *Pinus monticola* mortality do not appear to differ. Mortality for the Above High domain was low (see dashboard; MC permutation test results only found p < 0.01 when comparing Above High to Above Low) and predominantly represented by plots in the southern Sierra Nevada mountains of California.  *Pinus lambertiana* and *Calocedrus decurrens* exhibited high mortality in approximately the same region, a result supported by Fettig et al. (2019).  This region experienced an extreme drought that killed 170m trees between 2010 and 2021 (OEHHA 2022).  Dudney et al. (2020) also estimated lower mortality rates for *Pinus monticola* than for *Pinus lambertiana* in the Sequoia and Kings Canyon National Parks, which are within this region.  They attributed some of the difference in mortality to elevation, with mountain pine beetle (*Dendroctonus ponderosae*), fire, and white pine blister rust (*Cronartium ribicola*) more prevalent at lower elevations, corresponding to the range of *Pinus lambertiana*.

Our results for *Thuja plicata* did not corroborate with those from other researchers who are currently concerned about mortality in this species. Andrus et al. (2023) sampled *Thuja plicata* at nine locations in Washington, northern Oregon, and Idaho that were without canopy dieback, exhibited canopy dieback, or had died. They found that the mortality was associated with exceptionally warm and dry climate conditions. A collaboration of state and federal agencies is currently reaching out to the public to identify dieback sites for this species (website: Goodrich et al. 2025).  Our analysis found that *Thuja plicata* exhibited the lowest decadal mortality rate and third highest growth rate of any species we examined. It also had its greatest growth in Medium domains.  Our analysis may not be temporally sensitive enough to capture the same recent findings as Andrus et al., given that our FIA plot revisit data represented plot visits between 2011 to 2023.  If species conditions change quickly (on the order of 3-5 years) the FIA plot data may not reflect the change until a full or nearly full panel of plots is complete following the change in conditions.  Another non-exclusive possibility is that our study serves as one example of many broad-scale design-based analyses of spatially-balanced samples that offer a relatively unbiased assessment of species condition that can assist managers in identifying species of concern and prioritizing restoration actions.  *Thuja plicata* may indeed be senescing in many regions, but other species may be faring worse.

We believe the value of our analysis lies in its application to a relatively large number of tree species and its approach for investigating life history conditions such as growth and mortality that require the documentation of tree condition change.  The availability of full-panel FIA revisit data is a relatively recent development.  Prior to the availability of the revisit data, Monleon and Lintz (2015) assessed tree species’ range shifts using the single-visit data set, comparing mean plot temperature values for sapling and mature trees.  Groom and Monleon (2023) accessed to the more complete FIA revisit data set and found evidence of tree species density shifts along initial values of climatic variables but documented evidence of interactions between those variables’ initial and change values.   The literature of model- and design-based estimation of FIA revisit data is growing.  Examples of model-based analyses of mortality and growth using FIA revisit data include Shriver et al. (2022), Ogana et al. (2025), and Brandeis and Brandeis (2021).  Examples of growth and mortality design-based estimation that rely on FIA revisit data, similar to our state and overall analyses, include Vickers and Goff (2025), Goeking and Windmuller-Campione (2021), and Perret et al. (2023). Stanke et al. (2021) found estimates of condition indices.  Our bivariate examination of plot CWD shares some dimensional similarities with Dobrowski et al. (2013)  who examined the velocity of CWD (change in distance over time) along with AET and minimum temperatures. However, we are not aware of other design-based estimates of revisit data that examine more than one variable of interest simultaneously. Our approach allows us to capitalize on the amenability of the data to design-based estimation while accounting for the need, when examining revisit conditions, of capturing the initial and change value conditions.

CWD is a defensible variable to use for examining regional water stress as it integrates climatic and soil water availability through considering precipitation and air temperatures (Stephenson 1990, 1998).  CWD correlates well with tree growth (Dyer 2004, Restaino et al. 2016) and mortality (Anderegg et al. 2015, Das et al. 2013, Hammond et al. 2022).  However, there are some challenges in relying on CWD or any other climate variable in a design-based regional analysis such as this one. The challenges stem from the remeasurement timeframe, using a fixed delta CWD domain threshold, the broad, fixed scope of our analysis, and the lack of incorporation of local-scale conditions.

We did not detect sizable changes in CWD between visits, potentially signalling that 10 years is too soon to observe CWD-related species responses (perhaps more true for mortality than growth).  The CWD changes we observed may have represented greater proportional change for some species than others, with species such as *Abies amablis*, *Alnus rubra*, and *Juniperus occidentalis* exhibiting narrow and low water stress initial CWD ranges. CWD effects can manifest in direct and indirect ways, potentially differing with species by geographic location and the biotic and abiotic impacts to those locations.  It also cannot easily convey differences in impacts to trees within the same stands.

The scale of our analysis and the lack of incorporation of other variables likely limited our analysis because the evidence of water stress can manifest in multiple overlapping ways.  Water stress alone can kill trees (Anderegg et al. 2015, Maherali et al.  2004, and van Nieuwstadt and Sheil 2005), but it can also increase the probability of insect (Hart et al. 2014, Preisler et al. 2012 , examples and counterexamples in Huberty and Denno 2004) and wildfire-caused mortality (Abatzoglou and Williams 2016). With reference to Manion’s (1991) “death spiral” (see also Bennett et al. 2023 for a CWD-related version), teasing apart the cause of mortality from predisposing, inciting, and contributing factors is difficult at best, as these factors can interact (Gaylord et al. 2015).  Our regional analysis necessarily aggregates across local-scale variation in interrelated predisposing conditions for individual tree mortality such as canopy openness (Lemmo et al. 2024), structural complexity (Ma et al. 2023), tree height (Stovall et al. 2019), and within-stand competition (Ford et al. 2017). For management applications, particularly those for individual stands, our regional and domain estimates should be viewed as general results that may be modified by these local factors.

Although our current analysis offers a coarse examination of species status across a large area, future analyses could refine our bivariate approach for assessing tree species growth and mortality by domains.  Our climate domains could be divided further into sub-domains: trees could be assigned to tree size, stand competition, plot aspect, or plot elevation categories within the climate domain categories.  This approach has its limits: only estimates for certain widespread tree species would be reasonable to report, and domains could not be divided too finely into sub-domains, limiting the number of variables that could be considered simultaneously.  However, this approach would permit design-based estimation at a resolution better attuned to capturing factors affecting growth and mortality.

Our analysis may overall capture significant changes in mortality and growth for the tree species examined, with some trees exhibiting increases in mortality and decreases in growth either in portions of their ranges or across their entire range.  For some species in semi-arid environments such as *Pinus ponderosa*, *Pseudotsuga menziesii*, *Pinus contorta*, *Picea engelmannii*, and *Abies lasiocarpa*, we may be observing ecosystem shifts with a change in water stress combined with disturbances such as wildfire and insect mortality (Halofsky et al. 2020, Parks et al. 2019, Raymond et al. 2023).  Our hope is that our bivariate domain-based estimation approach is useful for others, while our results for individual species serve to bolster or direct others' inquiries into the status of the species we examined.  For managers, the results may be used to inform action for resiliency.   With the shift in ecosystems, landowners can assist with species migration when planting or replanting (Raymond et al. 2023), and forest thinning can help reduce water stress and the impacts of fire and insects (Bernal et al. 2023).

# Conclusion

Our analysis serves as an example of bivariate design-based estimation, where we use estimation domains based on initial CWD and delta CWD values of plots.  For the 23 species we examined, we found no general patterns in domains with higher or lower tree growth or mortality; the results appeared species specific.  We believe our analysis approach can be refined for individual species analysis, with the potential for integrating a third variable that could introduce information besides a condition variable and change variable.  Our analysis and the analyses of others provide valuable and reliable insights into the condition of several tree species that exhibit abnormally high mortality (and low growth).  Our hope is that these findings will inform management and encourage focused monitoring efforts for these species in the rapidly changing landscape of western North America.

# Acknowledgments

# References

::: {#refs}
Abatzoglou, J.T. and Williams, A.P., 2016. Impact of anthropogenic climate change on wildfire across western US forests. Proceedings of the National Academy of sciences, 113(42), pp.11770-11775.

Abatzoglou, J.T., Dobrowski, S.Z., Parks, S.A. and Hegewisch, K.C., 2018. TerraClimate, a high-resolution global dataset of monthly climate and climatic water balance from 1958–2015. Scientific data, 5(1), pp.1-12.

Adams, H.D., Macalady, A.K., Breshears, D.D., Allen, C.D., Stephenson, N.L., Saleska, S.R., Huxman, T.E. and McDowell, N.G., 2010. Climate‐induced tree mortality: Earth system consequences. Eos, Transactions American Geophysical Union, 91(17), pp.153-154.

Adams, H.D., Zeppel, M.J., Anderegg, W.R., Hartmann, H., Landhäusser, S.M., Tissue, D.T., Huxman, T.E., Hudson, P.J., Franz, T.E., Allen, C.D. and Anderegg, L.D., 2017. A multi-species synthesis of physiological mechanisms in drought-induced tree mortality. Nature ecology & evolution, 1(9), pp.1285-1291.

Alexander, R.R., Shearer, R.C. and Shepperd, W.D., 1990. Abies lasiocarpa (Hook.) Nutt. subalpine fir. Silvics of North America, 1, pp.60-70.

Anderegg, W.R., Flint, A., Huang, C.Y., Flint, L., Berry, J.A., Davis, F.W., Sperry, J.S. and Field, C.B., 2015. Tree mortality predicted from drought-induced vascular damage. Nature geoscience, 8(5), pp.367-371.

Andrus, R.A., Harvey, B.J., Chai, R.K. and Veblen, T.T., 2018. Different vital rates of Engelmann spruce and subalpine fir explain discordance in understory and overstory dominance. Canadian Journal of Forest Research, 48(12), pp.1554-1562.

Bechtold, W.A., Scott, C.T. 2005. The Forest Inventory and Analysis plot design, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Bennett, M., Shaw, D.C. and Lowrey, L., 2023. Recent Douglas-fir mortality in the Klamath Mountains ecoregion of Oregon: Evidence for a decline spiral. Journal of Forestry, 121(3), pp.246-261.

Betzen, J.J., Ramsey, A., Omdal, D., Ettl, G.J. and Tobin, P.C., 2021. Bigleaf maple, Acer macrophyllum Pursh, decline in western Washington, USA. Forest Ecology and Management, 501, p.119681.

Brandeis, T. and Brandeis, C., 2021. Assessing tree mortality probability in harvested hardwood stands using long-term forest inventory data. Forest Science, 67(2), pp.231-240.

Breshears, D.D., Myers, O.B., Meyer, C.W., Barnes, F.J., Zou, C.B., Allen, C.D., McDowell, N.G. and Pockman, W.T., 2009. Tree die‐off in response to global change‐type drought: mortality insights from a decade of plant water potential measurements. Frontiers in Ecology and the Environment, 7(4), pp.185-189.

Brown, K.R. and Courtin, P.J., 2018. Can phosphorus additions increase long-term growth and survival of red alder (Alnus rubra Bong.) on periodically dry sites?. Forest Ecology and Management, 430, pp.545-557.

Daly C, Halbleib M, Smith JI, Gibson WP, Doggett MK, Taylor GH, et al. Physiographically sensitive mapping of climatological temperature and precipitation across the conterminous United States. Int J Climatol. 2008; 28(15):2031–64. doi: 10.1002/joc.1688 PMID: WOS:000261810700004.

Das, A.J., Stephenson, N.L., Flint, A., Das, T. and van Mantgem, P.J., 2013. Climatic correlates of tree mortality in water-and energy-limited forests. PLoS One, 8(7), p.e69917.

Dobrowski, S.Z., Abatzoglou, J., Swanson, A.K., Greenberg, J.A., Mynsberge, A.R., Holden, Z.A. and Schwartz, M.K., 2013. The climate velocity of the contiguous United States during the 20th century. Global change biology, 19(1), pp.241-251.

Dudney, J.C., Nesmith, J.C., Cahill, M.C., Cribbs, J.E., Duriscoe, D.M., Das, A.J., Stephenson, N.L. and Battles, J.J., 2020. Compounding effects of white pine blister rust, mountain pine beetle, and fire threaten four white pine species. Ecosphere, 11(10), p.e03263.

Dyer, J.M., 2004. A water budget approach to predicting tree species growth and abundance, utilizing paleoclimatology sources. Climate Research, 28(1), pp.1-10.

Fettig, C.J., Mortenson, L.A., Bulaon, B.M. and Foulk, P.B., 2019. Tree mortality following drought in the central and southern Sierra Nevada, California, US. Forest Ecology and Management, 432, pp.164-178.

Ford, K.R., Breckheimer, I.K., Franklin, J.F., Freund, J.A., Kroiss, S.J., Larson, A.J., Theobald, E.J. and HilleRisLambers, J., 2017. Competition alters tree growth responses to climate at individual and stand scales. Canadian Journal of Forest Research, 47(1), pp.53-62.

Gaylord, M.L., Kolb, T.E. and McDowell, N.G., 2015. Mechanisms of piñon pine mortality after severe drought: a retrospective study of mature trees. Tree physiology, 35(8), pp.806-816.

Goeking, S.A. and Windmuller-Campione, M.A., 2021. Comparative species assessments of five-needle pines throughout the western United States. Forest Ecology and Management, 496, p.119438.

Gomez‐Gallego, M., Galiano, L., Martínez‐Vilalta, J., Stenlid, J., Capador‐Barreto, H.D., Elfstrand, M., Camarero, J.J. & Oliva, J. (2022) Interaction of drought‐ and pathogen‐induced mortality in Norway spruce and Scots pine. Plant, Cell & Environment, 45, 2292–2305. https://doi.org/10.1111/pce.14360

Goodrich, B., Fischer, M., Buhl, C.  2025. Western redcedar dieback.  https://storymaps.arcgis.com/stories/1405dab5f59246aa83849ec43f72b15a. Accessed 11 November 2025.

Groom, J.D. and Monleon, V.J., 2023. West coast US tree species show stronger evidence for shifts in density than range. Forest Ecology and Management, 545, p.121262. 
Hammond, W.M., Williams, A.P., Abatzoglou, J.T., Adams, H.D., Klein, T., López, R., Sáenz-Romero, C., Hartmann, H., Breshears, D.D. and Allen, C.D., 2022. Global field observations of tree die-off reveal hotter-drought fingerprint for Earth’s forests. Nature communications, 13(1), p.1761. https://doi.org/10.1038/s41467-022-29289-2

Hart, S.J., Veblen, T.T., Eisenhart, K.S., Jarvis, D. and Kulakowski, D., 2014. Drought induces spruce beetle (Dendroctonus rufipennis) outbreaks across northwestern Colorado. Ecology, 95(4), pp.930-939.

Hicke, J.A., Meddens, A.J. and Kolden, C.A., 2016. Recent tree mortality in the western United States from bark beetles and forest fires. Forest Science, 62(2), pp.141-153.

Huberty, A.F. and Denno, R.F., 2004. Plant water stress and its consequences for herbivorous insects: a new synthesis. Ecology, 85(5), pp.1383-1398.

Johnston, J.D., Stephens, P., Schmidt, M., Jeronimo, S., Brackett, A., Wuenschel, A., Ulrich, D.E., Paine, R., Duloisy, L., Greenler, S., Larson, A.J., and Churchill, D.J.  2025. Significant Mortality of Old Trees Across a Dry Forest Landscape, Oregon, USA. Forest Ecology and Management, Volume 593, 122907, ISSN 0378-1127, https://doi.org/10.1016/j.foreco.2025.122907.

Juang, C.S., Williams, A.P., Abatzoglou, J.T., Balch, J.K., Hurteau, M.D. and Moritz, M.A., 2022. Rapid growth of large forest fires drives the exponential response of annual forest‐fire area to aridity in the western United States. Geophysical Research Letters, 49(5), p.e2021GL097131.

Kolb, T.E., Fettig, C.J., Ayres, M.P., Bentz, B.J., Hicke, J.A., Mathiasen, R., Stewart, J.E. and Weed, A.S., 2016. Observed and anticipated impacts of drought on forest insects and diseases in the United States. Forest Ecology and Management, 380, pp.321-334.

Lemmo, S.L., Kerhoulas, L.P., Sherriff, R.L. and Beckmann, J.J., 2024. Drought effects on tree mortality and regeneration in northern California. Forest Ecology and Management, 563, p.121969.

Ma, Q., Su, Y., Niu, C., Ma, Q., Hu, T., Luo, X., Tai, X., Qiu, T., Zhang, Y., Bales, R.C. and Liu, L., 2023. Tree mortality during long-term droughts is lower in structurally complex forest stands. Nature communications, 14(1), p.7467.

Maherali, H., Pockman, W.T. and Jackson, R.B., 2004. Adaptive variation in the vulnerability of woody plants to xylem cavitation. Ecology, 85(8), pp.2184-2199.

Manion, P.D. 1991. Tree Disease Concepts, 2nd ed, 402. Englewood Cliffs, NJ: Prentice-Hall.

McDowell, N., Pockman, W.T., Allen, C.D., Breshears, D.D., Cobb, N., Kolb, T., Plaut, J., Sperry, J., West, A., Williams, D.G. and Yepez, E.A., 2008. Mechanisms of plant survival and mortality during drought: why do some plants survive while others succumb to drought?. New phytologist, 178(4), pp.719-739.

McNellis, B.E., Smith, A.M., Hudak, A.T. and Strand, E.K., 2021. Tree mortality in western US forests forecasted using forest inventory and Random Forest classification.  Ecosphere, 12(3), p.e03419.

McRoberts, R.E. 2005. The Enhanced Forest Inventory and Analysis Program, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Meddens, A.J., Hicke, J.A. and Ferguson, C.A., 2012. Spatiotemporal patterns of observed bark beetle‐caused tree mortality in British Columbia and the western United States. Ecological Applications, 22(7), pp.1876-1891.

Merschel, A.G., Beedlow, P.A., Shaw, D.C., Woodruff, D.R., Lee, E.H., Cline, S.P., Comeleo, R.L., Hagmann, R.K. and Reilly, M.J., 2021. An ecological perspective on living with fire in ponderosa pine forests of Oregon and Washington: Resistance, gone but not forgotten. Trees, Forests and People, 4, p.100074.

Monleon, V.J., Lintz, H.E., Delzon, S., 2015. Evidence of tree species’ range shifts in a complex landscape. PLoS One 10 (1), e0118069.

Negrón, J.F., McMillin, J.D., Anhold, J.A. and Coulson, D., 2009. Bark beetle-caused mortality in a drought-affected ponderosa pine landscape in Arizona, USA. Forest Ecology and Management, 257(4), pp.1353-1362.

Office of Environmental Health Hazard Assessment (OEHHA, 2022). Indicators of climate change in California, fourth edition, California Environmental Protection Agency, OEHHA.

Ogana, F.N., Green, P.C. and Radtke, P., 2025. Modeling the growth and yield of natural hardwood stands in the southern United States using the Forest Inventory and Analysis data. Forest Ecology and Management, 586, p.122722.

Parks, S.A., Coop, J.D. and Davis, K.T., 2025. Intensifying fire season aridity portends ongoing expansion of severe wildfire in western US forests. Global Change Biology, 31(8), p.e70429.

Perret, D.L., Bell, D.M., Gray, A.N., Shaw, J.D. and Zald, H.S., 2023. Range-wide population assessments for subalpine fir indicate widespread disturbance-driven decline. Forest Ecology and Management, 542, p.121128.

Preisler, H.K., Hicke, J.A., Ager, A.A. and Hayes, J.L., 2012. Climate and weather influences on spatial temporal patterns of mountain pine beetle populations in Washington and Oregon. Ecology, 93(11), pp.2421-2434.

Reams, G.A, Smith, W.D., Hansen, M.H., Bechtold, W.A., Roesch, F.A., Moisen, G.G. 2005. The forest inventory and analysis sampling frame, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Restaino, C.M., Peterson, D.L. and Littell, J., 2016. Increased water deficit decreases Douglas fir growth throughout western US forests. Proceedings of the National academy of Sciences, 113(34), pp.9557-9562.

Scott, C.T., W. A. Bechtold, G.A. Reams, W.D. Smith, J.A. Westfall, M.H. Hansen, Moisen, G.G. 2005. Sample-based estimators used by the Forest Inventory and Analysis National Information Management System, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Shive, K.L., Wuenschel, A., Hardlund, L.J., Morris, S., Meyer, M.D. and Hood, S.M., 2022. Ancient trees and modern wildfires: Declining resilience to wildfire in the highly fire-adapted giant sequoia. Forest Ecology and Management, 511, p.120110.

Shriver, R.K., Yackulic, C.B., Bell, D.M., & Bradford, J.B. (2022). Dry forest decline is driven by both declining recruitment and increasing mortality in response to warm, dry conditions. Global Ecology and Biogeography, 31, 2259–2269. https://doi.org/10.1111/geb.13582

Soderberg, D.N., A.J. Das, N. L. Stephenson, M.D. Meyer, C.A. Brigham, and J. Flickinger. 2024. “Assessing giant sequoia Mortality and regeneration following high-severity wildfire.” Ecosphere 15(3): e4789. https://doi.org/10.1002/ecs2.4789

Stanke, H., Finley, A.O., Domke, G.M., Weed, A.S., MacFarlane, D.W., 2021. Over half of western United States’ most abundant tree species in decline. Nat. Commun. 12, 451. https://doi.org/10.1038/s41467-020-20678-z.

Stephenson, N.L., and C. Brigham. 2021. Preliminary estimates of sequoia mortality in the 2020 Castle Fire. National Park Service, Sequoia and Kings Canyon National Parks.

Stephenson, N.L., 1998. Actual evapotranspiration and deficit: biologically meaningful correlates of vegetation distribution across spatial scales. Journal of biogeography, 25(5), pp.855-870.

Stephenson, N.L., 1990. Climatic control of vegetation distribution: the role of the water balance. The American Naturalist, 135(5), pp.649-670.     
Sala, A., Piper, F. and Hoch, G., 2010. Physiological mechanisms of drought-induced tree mortality are far from being resolved. The New Phytologist, 186(2), pp.274-281.

Sala, A., Woodruff, D.R. and Meinzer, F.C., 2012. Carbon dynamics in trees: feast or famine?. Tree physiology, 32(6), pp.764-775.

Stovall, A.E., Shugart, H. and Yang, X., 2019. Tree height explains mortality risk during an intense drought. Nature communications, 10(1), p.4385.
Thompson, J. 2015. PNW-FIADB Users Manual: A data dictionary and user guide for the PNW-FIADB database. Forest Inventory and Analysis, Pacific Northwest Research Station, Portland, Oregon.

Van Mantgem, P.J., Stephenson, N.L., Byrne, J.C., Daniels, L.D., Franklin, J.F., Fulé, P.Z., Harmon, M.E., Larson, A.J., Smith, J.M., Taylor, A.H. and Veblen, T.T., 2009. Widespread increase of tree mortality rates in the western United States. Science, 323(5913), pp.521-524. DOI: 10.1126/science.1165000

Van Nieuwstadt, M.G. and Sheil, D., 2005. Drought, fire and tree survival in a Borneo rain forest, East Kalimantan, Indonesia. Journal of ecology, 93(1), pp.191-201.
Vickers, L.A. and Goff, T.C., 2025. White oak sapling dynamics across the eastern US: Quantifying a bottleneck to sustainability. Forest Ecology and Management, 598, p.123232.

Washington State Department of Natural Resources (WADNR) Forest Reliance Division. 2024. Forest health highlights in Washington: A summary of insect, disease, and other disturbance conditions affecting Washington’s forests.  Washington State Department of Natural Resources, Forest Resilience Division, Olympia, WA. https://dnr.wa.gov/sites/default/files/2025-03/rp_2023_fhh_report_final_wr.pdf, accessed 9/18/2025

Westerling ALR. 2016 Increasing western US forest wildfire activity: sensitivity to changes in the timing of spring. Phil. Trans. R. Soc. B 371: 20150178. http://dx.doi.org/10.1098/rstb.2015.0178

Woudenberg, S.W., Conkling, B.L., O'Connell, B.M., LaPoint, E.B., Turner, J.A. and Waddell, K.L., 2010. The Forest Inventory and Analysis Database: Database description and users manual version 4.0 for Phase 2. Gen. Tech. Rep. RMRS-GTR-245. Fort Collins, CO: US Department of Agriculture, Forest Service, Rocky Mountain Research Station. 336 p.

:::

<!---BLOCK_LANDSCAPE_START--->

```{r tbl-growth, tab.cap=tab.tbl_growth}

growth.ft

```

```{r tbl-mort, tab.cap = tab.tbl_mort}

mort.ft

```

<!---BLOCK_LANDSCAPE_STOP--->

```{r tbl-summ-grow-mort, tab.cap = tab.summ_grow_mort}

g.m.summ.ft

```

```{r mapfig, fig.height = 7,  fig.dpi=300, fig.cap=fig.cwdmap_cap, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "cwd_maps_figure.png"))

```

```{r domainsfig, fig.width = 6,  fig.dpi=300, fig.cap=fig.domains, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "Example_spp_figure.png"))

```

```{r diff_grow_fig, fig.height = 6.5, width = 6.5, fig.dpi=300, fig.cap=fig.diff_grow, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Growth_figs_cwd/Growth_Panel_Plot.png"))

```

```{r diff_mort_fig, fig.height = 6.5, width = 6.5, fig.dpi=300, fig.cap=fig.diff_mort, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Mort_figs_cwd/Mort_Panel_Plot.png"))

```