### 2025-12-12 Manuscript.Rmd
Incorporated edits from B. Frank.  This added a figure, changed the supplimental and main RMarkdown files, and altered many of the scripts involved in generating plots and tables.

---
title: "Tree growth, mortality, and climatic water deficit in west-coast states, USA"
author:
  - name: "Jeremiah D. Groom"
    affiliations:
      - name: "Groom Analytics LLC"
        address: "1975 SE Crystal Lake Dr., Unit 173"
        city: "Corvallis"
        state: "OR"
        postal-code: "97333"
    orcid: "0000-0002-0110-1036"
    email: "jeremy@groomanalytics.com"
  - name: "Bryce Frank"
    affiliations:
      - name: "USDA Forest Service"
        department: "Pacific Northwest Field Station"
        address: "3625 93rd Ave SW"
        city: "Olympia"
        state: "WA"
        postal-code: "98512"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  officedown::rdocx_document:
    reference_docx: "template_use2.docx"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(kableExtra)   # Enhanced table formatting
library(scales)       # Scale functions for visualization
library(flextable)    # Professional tables for Word output
library(officedown)
library(officer)
library(rvg)
#devtools::install_github("adletaw/captioner")
library(captioner) # For figure captions.

```

```{r other_settings}
fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)
```

```{r constants}

source("Global.R")
source(paste0(CODE.LOC, "Functions.R"))

# Define any custom functions
format_p_value <- function(p) {
  case_when(
    p < 0.001 ~ "< 0.001",
    p < 0.01 ~ sprintf("= %.3f", p),
    TRUE ~ sprintf("= %.2f", p)
  )
}

set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11,
  theme_fun = "theme_vanilla"
)

centeredP <- fp_par(text.align = "center")

```

```{r load-data}
n.spp <- length(sel.spp)

# Some general analysis info + plots
analysis.stats <- read_rds(paste0(RESULTS.OTHER, "analysis.stats.RDS"))

# Percent of plots sampled with REMPER = 9, 10, and 11 years (rounded)
pct.remeas <- as.vector(analysis.stats$pct.remeas)

# Species scientific names
spp.names.use <- analysis.stats$gym.angio.dat

# Correlation/fits for Initial CWD vs Delta CWD:
CWD.rels <- read_rds(paste0(RESULTS.OTHER, "CWD_deltaCWD_relationship.rds"))
CWD.lin.cor <- format(round(CWD.rels$lm1$r.squared^0.5, 2), nsmall = 2) # Last digit = 0, want to show.
CWD.dAIC <- round(CWD.rels$aic.cwd$AIC[1] - CWD.rels$aic.cwd$AIC[2], 0)

# Growth data for species overall, states, differences
grow.state <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/State_Growth_", CLIM.VAR.USE, ".csv"))
grow.all <- read_csv(paste0(RESULTS1.LOC, "Growth_figs_", CLIM.VAR.USE, "/All_Growth_", CLIM.VAR.USE, ".csv"))
grow.diff <- read_rds(paste0(save.loc.fcn(1), "Processed_6Domain_Data.RDS"))

# Mortality data for species overall and states
mort.state <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/State_Mortality_", CLIM.VAR.USE, ".csv"))
mort.all <- read_csv(paste0(RESULTS1.LOC, "Mort_figs_", CLIM.VAR.USE, "/All_Mortality_", CLIM.VAR.USE, ".csv"))
mort.diff <- read_rds(paste0(save.loc.fcn(2), "Processed_6Domain_Data.RDS"))

# Function to extract a value out of the contrast figures by species and contrast
fig.val.fcn <- function(dat, spp.x, cont, rnd){
dat.source <- if(dat == "g") grow.diff else mort.diff
k <- if(dat == "g") 1 else 2
IvS2 <- cm2.fcn(k, results.table = dat.source$IvS2)
I_LMH2 <- cm2.fcn(k, results.table = dat.source$I_LMH2)
S_LMH2 <- cm2.fcn(k, results.table = dat.source$S_LMH2)
tab.comb <- bind_rows(IvS2, I_LMH2, S_LMH2)

val.out <- round(tab.comb$Means[tab.comb$Species == spp.x & tab.comb$Domain == cont], rnd)
return(val.out)
}

# Significant findings, Growth
grow.n.IvS <- nrow(grow.diff$IvS2)
grow.n.I <- nrow(grow.diff$I_LMH2)
grow.n.S <- nrow(grow.diff$S_LMH2)

grow.IvS.sig <- length(grow.diff$IvS2$significant[grow.diff$IvS2$significant == TRUE])
grow.IvS.signeg <- length(grow.diff$IvS2$significant[grow.diff$IvS2$significant == TRUE & grow.diff$IvS2$Means < 0])
grow.I.sig <- length(grow.diff$I_LMH2$significant[grow.diff$I_LMH2$significant == TRUE])
grow.I.signeg <- length(grow.diff$I_LMH2$significant[grow.diff$I_LMH2$significant == TRUE & grow.diff$I_LMH2$Means < 0])
grow.S.sig <- length(grow.diff$S_LMH2$significant[grow.diff$S_LMH2$significant == TRUE])
grow.S.signeg <- length(grow.diff$S_LMH2$significant[grow.diff$S_LMH2$significant == TRUE & grow.diff$S_LMH2$Means < 0])

grow.tot.sig <- grow.IvS.sig + grow.I.sig + grow.S.sig
grow.tot <- grow.n.IvS + grow.n.I + grow.n.S

# Significant findings, Mort
mort.n.IvS <- nrow(mort.diff$IvS2)
mort.n.I <- nrow(mort.diff$I_LMH2)
mort.n.S <- nrow(mort.diff$S_LMH2)

mort.IvS.sig <- length(mort.diff$IvS2$significant[mort.diff$IvS2$significant == TRUE])
mort.IvS.signeg <- length(mort.diff$IvS2$significant[mort.diff$IvS2$significant == TRUE & mort.diff$IvS2$Means < 0])
mort.I.sig <- length(mort.diff$I_LMH2$significant[mort.diff$I_LMH2$significant == TRUE])
mort.I.signeg <- length(mort.diff$I_LMH2$significant[mort.diff$I_LMH2$significant == TRUE & mort.diff$I_LMH2$Means < 0])
mort.S.sig <- length(mort.diff$S_LMH2$significant[mort.diff$S_LMH2$significant == TRUE])
mort.S.signeg <- length(mort.diff$S_LMH2$significant[mort.diff$S_LMH2$significant == TRUE & mort.diff$S_LMH2$Means < 0])

mort.tot.sig <- mort.IvS.sig + mort.I.sig + mort.S.sig
mort.tot <- mort.n.IvS + mort.n.I + mort.n.S

# Summary of significant Growth and Mortality findings for table creation.
datasets.grow <- list(grow.diff$IvS2, grow.diff$I_LMH2, grow.diff$S_LMH2)
datasets.mort <- list(mort.diff$IvS2, mort.diff$I_LMH2, mort.diff$S_LMH2)
datatypes <- list("Increasing vs. Stable", "Increasing", "Stable")
n.comp.grow <- list(grow.n.IvS, grow.n.I, grow.n.S)
n.comp.mort <- list(mort.n.IvS, mort.n.I, mort.n.S)

growth.summ <- pmap(list(datasets.grow, datatypes, n.comp.grow), summ.fig.fcn) %>% bind_rows()
mortality.summ <- pmap(list(datasets.mort, datatypes, n.comp.mort), summ.fig.fcn) %>% bind_rows()

```

```{r table-creation}
# State growth
grow.state.cm <- cm2.fcn(1, grow.state)
grow.all.cm <- cm2.fcn(1, grow.all)
s.g.dat <- state.table.fcn(grow.state.cm, grow.all.cm, 1) %>%
  left_join(spp.cat %>% dplyr::select(-spp_cat), by = c(`Species Name` = "Species")) %>%
  rename("Category" = "cat_abbr") %>%
  relocate(Category, .after = `Species Name`)

growth.ft <- flextable(s.g.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
    italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = TRUE)
min.grow.spp <- maxmin.fcn(grow.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = TRUE)

# State mortality
s.m.dat <- state.table.fcn(mort.state, mort.all, 2) %>%
  left_join(spp.cat %>% dplyr::select(-spp_cat), by = c(`Species Name` = "Species")) %>%
  rename("Category" = "cat_abbr") %>%
  relocate(Category, .after = `Species Name`)

mort.ft <- flextable(s.m.dat) %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  italic(j = 1, part = "body") %>%
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",  
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
  set_table_properties(layout = "autofit", width = 0)

max.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "max", cm2 = FALSE, digits = 2)
min.mort.spp <- maxmin.fcn(mort.all, "Means", "LCI.95", "UCI.95", extreme = "min", cm2 = FALSE, digits = 2)

# CWD amounts, tapping values from the initial & delta CWD tables in the suppliment
init.cwd.stats <- analysis.stats$init.cwd.stats %>%
  mutate(range05.95 = q95 - q05) %>%
  arrange(Mean)

delta.cwd.stats <- analysis.stats$diff.cwd.stats %>%
  mutate(range05.95 = q95 - q05) %>%
  arrange(Mean)

cwd.summ.fcn <- function(data.2use, col.interest, place, stat.use) {
end.start <- switch(stat.use, "min" = 1, "max" = n.spp) # Which row end (top or bottom) to start on?
place2 <- if(stat.use == "min") place else n.spp - place + 1

x <- data.2use %>% arrange(get(col.interest))
x2 <- x[place2, ]
x2.range <- paste0("(5th percentile = ", x2$q05, "mm, 95th percentile = ", x2$q95, "mm)")
x2.range.full <- paste0("(5th-95th percentile range = ", x2$range05.95, "mm, 5th percentile = ", x2$q05, "mm, 95th percentile = ", x2$q95, "mm)")
x2.mean.full <- paste0("(mean = ", x2$Mean, "mm, 5th percentile = ", x2$q05, "mm, 95th percentile = ", x2$q95, "mm)")

return(list(spp.out = x2$Species, range.alone = x2$range05.95, mean.alone = x2$Mean,
            spp.range = x2.range, 
            spp.range.full = x2.range.full, spp.mean.full = x2.mean.full
            ))
}

#Initial CWD: Tree species with the min/max means and ranges.
init.mean.sm <- cwd.summ.fcn(init.cwd.stats, col.interest = "Mean", place = 1, stat.use = "min") #smallest mean
init.mean.sm2 <-   cwd.summ.fcn(init.cwd.stats, col.interest = "Mean", place = 2, stat.use = "min") # 2nd smallest mean
init.range.sm <- cwd.summ.fcn(init.cwd.stats, col.interest = "range05.95", place = 1, stat.use = "min") # smallest range
init.range.sm2 <- cwd.summ.fcn(init.cwd.stats, col.interest = "range05.95", place = 2, stat.use = "min")
init.mean.lg <- cwd.summ.fcn(init.cwd.stats, col.interest = "Mean", place = 1, stat.use = "max") # largest mean
init.mean.lg2 <- cwd.summ.fcn(init.cwd.stats, col.interest = "Mean", place = 2, stat.use = "max")
init.range.lg <- cwd.summ.fcn(init.cwd.stats, col.interest = "range05.95", place = 1, stat.use = "max")
init.range.lg2 <- cwd.summ.fcn(init.cwd.stats, col.interest = "range05.95", place = 2, stat.use = "max")

#Initial CWD: Tree species with the min/max means and ranges.
delta.mean.sm <- cwd.summ.fcn(delta.cwd.stats, col.interest = "Mean", place = 1, stat.use = "min") #smallest mean
delta.mean.sm2 <-   cwd.summ.fcn(delta.cwd.stats, col.interest = "Mean", place = 2, stat.use = "min") # 2nd smallest mean
delta.range.sm <- cwd.summ.fcn(delta.cwd.stats, col.interest = "range05.95", place = 1, stat.use = "min") # smallest range
delta.range.sm2 <- cwd.summ.fcn(delta.cwd.stats, col.interest = "range05.95", place = 2, stat.use = "min")
delta.mean.lg <- cwd.summ.fcn(delta.cwd.stats, col.interest = "Mean", place = 1, stat.use = "max") # largest mean
delta.mean.lg2 <- cwd.summ.fcn(delta.cwd.stats, col.interest = "Mean", place = 2, stat.use = "max")
delta.range.lg <- cwd.summ.fcn(delta.cwd.stats, col.interest = "range05.95", place = 1, stat.use = "max")
delta.range.lg2 <- cwd.summ.fcn(delta.cwd.stats, col.interest = "range05.95", place = 2, stat.use = "max")

min.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.min.mean]
min.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.min.mean]
max.mean.init <- analysis.stats$init.cwd.stats$Mean[analysis.stats$init.cwd.max.mean]
max.mean.init.spp <- analysis.stats$init.cwd.stats$Species[analysis.stats$init.cwd.max.mean]

max.q95.diff <- analysis.stats$diff.cwd.stats$q95[analysis.stats$diff.cwd.95max]
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]
max.range.diff <- analysis.stats$diff.cwd.90range.max
max.q95.diff.spp <- analysis.stats$diff.cwd.stats$Species[analysis.stats$diff.cwd.95max]

# Adding a section identifier to each table
growth.summ2 <- growth.summ %>%
  mutate(Type = "Growth", .before = 1)

mortality.summ2 <- mortality.summ %>%
  mutate(Type = "Mortality", .before = 1)

# Combine the tables
grow.mort.summ <- bind_rows(growth.summ2, mortality.summ2)

# Create flextable
g.m.summ.ft <- flextable(grow.mort.summ) %>%
  merge_v(j = "Type") %>%  # Merge cells in Section column
  valign(j = "Type", valign = "top") %>%  # Top align section labels
  flextable::align(j = "Type", align = "left") %>%   # Left align section labels
  line_spacing(i = NULL, j = NULL, space = 0.5, part = "body") %>% # Control space between lines. 
  set_table_properties(layout = "autofit", width = 0) %>%
  theme_zebra(
    odd_header = "#FFFFFF",   # White background for all
    even_header = "#FFFFFF",
    odd_body = "#FFFFFF",     
    even_body = "#FFFFFF"     
  ) %>%
    flextable::align(j = 3:7, align = "center", part = "all")

```

```{r set_captions}

# 
fig_nums <- captioner::captioner(prefix = "")
tab_nums <- captioner::captioner(prefix = "Table")

f.ref <- function(x) {
 caption_text <- fig_nums(x)
  # Remove any leading numbers and extract just "Figure X"
  stringr::str_extract(caption_text, "\\d+")
}

t.ref <- function(x)  {
  stringr::str_extract(tab_nums(x), "[^:]*")
}

tab.tbl_growth <- tab_nums(name = "tbl_growth", 
                        caption = "Decadal growth estimates by state and overall for analysis species (cm\u00B2/decade). Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots. Ecological category abbreviations (Category) are provided in the text.")

tab.tbl_mort <- tab_nums(name = "tbl_mort", 
                        caption = "Decadal mortality estimates by state and overall for analysis species. Mean estimates with confidence intervals and plot number are provided. Estimates are not made for states with fewer than 10 plots. Ecological category abbreviations (Category) are provided in the text.")

fig.cwdmap_cap <- fig_nums(name = "fig_cwdmap", 
                        caption = "Fuzzed FIA plot locations displayed with initial visit CWD values (A) and change in CWD between visits (C) for the U.S. states, north to south, of Washington, Oregon, and California (see inset map B for the location of these states within the USA).")

fig.domains <- fig_nums(name = "fig_domains",
                        caption = "This figure illustrates our domain analysis approach using *Calocedrus decurrens* as an example. Here we depict decadal (A) growth and (B) mortality domain estimates.  The six domains are defined as FIA plots that fall into the [I]ncreasing or [S]table CWD change threshold with [L]ow, [M]edium, or [H]igh initial CWD values (see text for details). E.g., 'IL' is in the Increasing CWD change category, Low initial CWD value. The distribution of plot CWD change and initial values (C) is presented as well, with the blue horizontal line indicating the CWD change category and the orange vertical lines representing the quantile boundaries between Low, Medium, and High initial CWD values." )

fig.cwd_initdelt <- fig_nums(name = "fig_cwdmeans", 
                          caption = paste0("Mean species plot values for the climatic water deficit initial and change values."))

fig.diff_grow <- fig_nums(name = "fig_diffgrow", 
                          caption = paste0("Differences in mean growth rates (cm\u00B2/decade) between domains by species.  Domain abbreviations are described in the legend for Figure ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are conifers; those below are hardwoods."))

fig.diff_mort <- fig_nums(name = "fig_diffmort", 
                          caption = paste0("Differences in mean decadal mortality rates between domains by species.  Domain abbreviations are described in the legend for Figure ",
                                           f.ref("fig_domains"), 
                                           ". See each column legend for the key of domain differences.  Mean domain differences (circles) with 95% CIs (bars) that do not intersect zero are filled. Species above the dashed line are conifers; those below are hardwoods."))

tab.summ_grow_mort <- tab_nums(name = "tbl_summ_grow_mort",
                               caption = paste0("Summary of number of contrasts with 95% CIs that do not overlap with zero ([T]otal, [P]ositive-, and [N]egative-valued estimates) for Figures ", fig_nums("fig_diffgrow", display = "num"), " and ", fig_nums("fig_diffmort", display = "num"), " by analysis (T)ype (Growth or Mortality) and type of (C)omparison.  The column Domains lists the domains compared; All is the sum of values within a Comparison. See text for other domain labels.  'n' is the total number of differences examined within a Comparison."))

```

# Abstract

Western North American forests are experiencing accelerating tree mortality rates in response to intensifying drought and water stress driven by climate change.  Tree mortality and growth reflect cumulative effects of direct physiological stress (xylem cavitation, carbon starvation), and indirect climate-mediated impacts (wildfire, insects, pathogens).  Relating design-based estimates for mortality and growth to a climate variable can be challenging, as directly assessing mortality and growth requires the visitation of plots at two different points in time.  Plots can differ both in their initial climate variable value and the change in that value between visits, necessitating estimation domains that jointly consider initial conditions and temporal change. We developed an analysis framework that partitions plots into domains defined by their initial-visit climatic water deficit (CWD) value and the change in the CWD value between visits.  We applied this approach to `r comma(analysis.stats$plot.num)` National Forest Inventory plots spanning California, Oregon, and Washington, tracking `r comma(analysis.stats$tree.num)` trees from `r n.spp` species over decade-long intervals.  We estimated decadal growth and mortality overall, by state, and across six CWD domains per species, along with associated measures of uncertainty.  Species responses were highly variable, with no universal pattern of increased mortality or reduced growth in domains experiencing elevated water stress. Several species exhibited overall mortality rates exceeding 0.20 (i.e., 20% decadal mortality), with estimates of 0.34 for *Picea engelmannii* in Washington, 0.27 for *Acer macrophyllum* in California, and 0.37 for *Pinus contorta* in Washington. Domain-specific patterns aligned with known regional disturbances (bark beetles, drought, fire) for multiple species. Our study contributes a template for monitoring forest vital rates as ecosystem-level indicators of climate change response, and our species-specific findings identify or validate populations warranting immediate conservation assessment that can inform forest management strategies.

# Introduction

Western North American forests are experiencing increases in tree mortality across multiple species in concert with changing climate conditions.  Species and tree groups have incurred extraordinary losses in specific regions.  Johnston et al. (2025) documented a 29% 10-year mortality rate of 300+ yr old trees in Eastern Oregon.  High-severity fires in California in 2020 and 2021 killed 13%-19% of extant giant sequoia (*Sequoiadendron gianteum*) with >1.07-m bole diameters (Stephenson and Brigham 2021, Shive et al. 2022); recruitment in these burns was depressed (Soderberg et al. 2022).  Species and genera have also seen their mortality rates increase overall. van Mantgem et al. (2009) modeled mortality rates for the genera *Abies* and *Pinus* in the western USA as increasing from less than 0.05% annually in 1960 to almost 0.2% by 2010. Perret et al. (2023) estimated a decadal range-wide mortality rate of 25% for Subalpine fir (*Picea engelmannii*).    Five of eight common tree species in western North America exhibit declines as measured by relative stand density (Stanke et al. 2021).  Shriver et al. (2022) found that four of five juniper (*Juniperus* spp.) and piñon pine (*Pinus* spp.) species were declining in portions of their range, comparing mortality to recruitment rates.

Climate change as manifest through increases or intensification in drought and aridity can affect tree mortality directly and indirectly.  Direct impacts include physiological damage while indirect causes include increasing wildfire, insect, and disease outbreak frequency and severity.  Hammond et al. (2022) found globally that the intensity of drought and heat events are directly associated with widespread forest die-off events.  Wildfire frequency and area burned in the western United States has increased in concert with the longer fire seasons, hotter summer temperatures, and earlier snowmelt (Abatzoglou and Williams 2016, Westerlin 2016, Juang et al. 2022).  There has also been a regional increase in the area of bark beetle (subfamily *Scolytinae* spp.) outbreaks (Meddens et al. 2012, Hicke et al. 2016).  Many of these indirect climate-influenced mortality causes may be attributable to the physical condition of trees following abnormal or prolonged exposure to water stress.

Carbon starvation and hydraulic failure can drive drought-induced mortality (see reviews by Adams et al. 2010, Sala et al. 2010, 2012, and McDowell et al. 2008).  Carbon starvation occurs when isohydric species (e.g., piñon pine, *Pinus edulus*; Breshears et al. 2009), or those which mainly rely on stomatal closure to avoid excessive water loss through evapotranspiration (ET), receive insufficient atmospheric carbon dioxide for photosynthesis.  Anisohydric plants, or those that do not strongly regulate water loss through stomatal closure, risk hydraulic failure if sufficient xylem cavitation occurs due to insufficient water uptake needed to meet ET demands (e.g., as seen in *Populus tremuloides*; Anderegg et al. 2015). Adams et al. (2017) performed a review of drought-induced mortality across 26 species and concluded that hydraulic failure occurred broadly across taxa while non-structural carbohydrate reserves at mortality differed across taxa.  Regardless of mechanism, drought (specifically, water deficit) appears to be an important direct driver of mortality for tree taxa (Hammond et al. 2022).

Given the importance of water stress on tree mortality, selecting an appropriate metric is critical for rangewide assessments.  One regionally useful metric of water stress is Climatic Water Deficit (CWD; Stephenson 1998).  Stephenson (1990) defines CWD as evaporative demand not met by available water, or the difference between plant water demand (actual evapotranspiration, AET), and plant water supply (potential evapotranspiration); these variables are driven by temperature and precipitation, respectively.  The Anderegg  et al. (2015) tree mortality study relied on CWD to capture drought occurrence.  Das et al. (2013) determined that low-elevation California tree mortality was best modeled using CWD.  Restaino et al. (2016) found decreased Douglass fir growth with an increase in CWD.

Data from the U.S. National Forest Inventory, collected by the US Forest Service Forest Inventory and Analysis Program (FIA), is well suited for evaluating regional species mortality and growth. Since 1999 the FIA program has gathered a variety of forestry data at plots nationwide using a spatially balanced sampling design (McRoberts 2005, Reams et al. 2005, Woudenberg et al. 2010).  The sampling design allows for approximately unbiased design-based inference of population characteristics across large geographical regions.  Data from plot revisits and individual tree data permit direct estimation of change features such as mortality and growth estimation.

Climatic data such as mean annual temperature and CWD can be combined with FIA data to obtain climate-related design-based estimates. Monleon and Linz (2015) evaluated the evidence of tree species range shifts by subtracting the design-based mean estimates of FIA plot temperatures for seedlings from those of trees. However, FIA plot revisit data were mostly incomplete at the time of their analysis. In contrast, Groom and Monleon (2023) used a mostly-completed revisit data set to compare plot climate variable values (vapor pressure deficit [VPD], temperature, and precipitation) to estimate shifts in plot occupancy and density.  Both studies used a single climate variable value associated with each plot for estimation; e.g., Groom and Monleon (2023) examined shifts in species density by estimating the difference in VPD means between plots that gained trees of a given species and plots that lost trees.

Groom et al. (2023) documented, however, that the initial climate variable values could (and often did) change between plot visits, with species' plots often exhibiting positive or negative correlations between the initial visit values and between-visits change value of the climate variable.  They identified that their design-based analysis would potentially improve if they could simultaneously include both initial visit and change values for the climate variable.

Our study has two objectives.  The first is to demonstrate, across 23 species, a design-based estimation approach for tree growth and mortality that simultaneously accounts for initial climatic conditions and change between measurements.  This addresses a gap in design-based analyses of forest change.  Second, we provide current growth and mortality estimates for our analysis species across our study area and by state to establish species status and to inform our climate analysis findings.

# Methods

## FIA data

Our analysis relies on FIA data for forestland in California, Oregon, and Washington, USA.  Forestland is defined as an area of at least 4,050 m$^2$ (i.e., one acre) that is at least 10 percent stocked or potentially stocked with tree species, excluding urban and agricultural land uses (Reams et al. 2005).  We constrained our analysis to only include remeasurements for areas considered forestland during both visits.

In the coterminous United States, the FIA program uses a spatially balanced, quasi-systematic sampling design. Plots are randomly located and produce sampling intensities of approximately 2,400 hectares per plot, with some denser areas such as those located on national forest system lands. In the western United States, each plot is assigned to one of ten panels, and approximately one panel is measured each year in perpetuity, yielding ten year remeasurement intervals on each plot, barring exceptions owing to optional temporal intensifications (Thompson 2015). In this study, we use only those plots that are part of the standard sampling intensity, and have been remeasured at an interval of ten years. Each FIA plot consists, in part, of four 7.32 m radius subplots where all trees with stem diameters greater than or equal to 12.7 cm (for most species measured at 1.37 m above the ground) were tallied with additional trees sampled in an annulus around each subplot (macroplot), with a total radius 18 m from the center point and a total area of 1012 m$^2$.  We did not include data from microplots, where seedlings and saplings are measured, as the analysis was not intended to estimate species recruitment but rather to follow the trajectory of established trees. Reams et al. (2005) and Bechtold and Scott (2005) provide plot design details and measurement protocols.

Our analysis depends on examining the fate and growth of trees measured during two sequential visits (referred to as the first and second plot visit), ideally from a complete sample panel remeasure. Tree growth is the observed increase in marked tree bole diameter between visits (cm$^2$) while mortality is captured from the change in tree condition. We obtained our FIA data from the USDA FIA DataMart web portal (https://research.fs.usda.gov/products/dataandtools/fia-datamart; downloaded May 5, 2025).  The database included plot data sampled from 2001 to 2023.

Remeasured trees within plots were retained in the analysis if they were alive at the initial visit and met the stem diameter criterion during that visit (i.e., plot ingrowth was not included).  The growth analysis additionally restricted trees considered to those that were alive at both visits.  The mortality analysis excluded trees killed through silvicultural practices.  Both analyses excluded trees that shrank in diameter, physically moved due to landslides or other events, were erroneously tallied at either visit, were missed due to data collection procedural changes, or existed at either visit in a nonsampled area.

## Climate data

We obtained plot-level CWD data from the TerraClimate dataset (Abatzoglou et al. 2018) for June, July, and August (summer months) between 1980 and 2023. B. Frank determined that the Pearson's correlation between the fuzzed and unfuzzed CWD plot values was > 0.99, so for simplicity we used the fuzzed CWD values and spatial coordinates. We defined 'initial CWD' values as the 20-year mean of summer CWD values for each plot prior to and including the first visit year (initial CWD = mean(summer CWD from visit year$_{-19}$ to year$_0$)).  We then found the mean of summer CWD values between visits and including the second visit year (second visit CWD = mean(summer CWD from year$_1$ to year$_10$)).  We then defined 'delta CWD' as the second visit CWD values minus initial CWD values.  Figure `r fig_nums("fig_cwdmap", display = "num")` presents maps of both the initial CWD and delta CWD values for forested and nonforested FIA plots.

## Species  Classification

We coarsely categorized tree species by several taxonomic and ecological characteristics, such as tree type (angiosperm [hardwood] vs. conifer), drought tolerance, fire adaption, elevation, shade tolerance, and mesic and riparian specialization.  We obtained the category information from https://plants.usda.gov/ (accessed 12/12/2025) and Burns and Honkala (1990a, 1990b).  Categorization was done to aid the discussion of results; some species could fall under more than one category (e.g., we classified *Picea engelmannii* as a high elevation conifer [HEC] while it could have been classified as a shade-tolerant conifer [STC]).  Our categories with abbreviations also include fire-adapted conifers (FAC), drought-tolerant conifers (DTC), drought-tolerant hardwoods (DTH) and riparian and mesic hardwoods (RMH).  These abbreviations are used in Tables `r tab_nums("tbl_growth", display = "num")` and `r tab_nums("tbl_mort", display = "num")`.

## Analysis
We limited the number of species examined to those with more than 400 occupied plots across the states of Oregon, Washington, and California.  We obtained point estimates of species growth and mortality using ratio estimators (Eqn 4.16 in Scott et al. 2005).  We expanded the number of sampled trees in each plot by the area measured (subplot or macroplot area) adjusted by the proportion of the plot area available for measurement and the number of years between visits. The numerator for species-specific mortality is the sum of stratum-weighted means of the number of trees that died between visits per plot.  The denominator for mortality is the sum of stratum-weighted means of all live trees per plot of a given species at the initial visit. Thus, the estimator describes the proportion of trees that transitioned from live to dead between remeasurements.

The numerator of species-specific growth estimates is the sum of stratum-weighted means of the increase in basal area (cm$^2$) between visits per plot while the denominator is the same as was used for the mortality estimate. This produces a point estimate for a species’ mean growth on an individual tree basis.  We standardized mortality and growth values to a decadal time period.

We report species mortality and growth estimates for all plots across and within Oregon, Washington, and California. Separately, we investigate climatic effects on tree mortality and growth by partitioning trees into subpopulations (domains).  For individual species we estimate domain means and measures of uncertainty for both mortality and growth. In this study we use six domains based on CWD values. We created the domains based on initial CWD plot values by partitioning the values into three species-specific quantiles: values below the first quartile (Low), between the first and third quartiles (Medium), and above the third quartile (High; Figure `r fig_nums("fig_domains", display = "num")`C).  We defined the two domains for delta CWD plot values for all species as those above a 5mm CWD change threshold, which we interpret as increasing in CWD (Increasing), and those below the threshold (Stable). The 5mm threshold was selected because species' delta CWD value distributions did not differ greatly and because 5mm is approximately the mean delta CWD value across species (Table S2).  The Stable domains may include some negative delta CWD values (i.e., mean CWD was larger prior to the initial visit than between visits), however these instances were rare and of small magnitude. The domain designations IL, IM, and IH correspond to Increasing Low, Increasing Medium, and Increasing High. The Stable domains are similarly labeled (SL, SM, SH). These six within-domain estimates allowed us to establish mortality and growth rates for trees that experienced similar climatic conditions for a given species. We calculated estimates for all CWD-based domains with 10 or more plots.

In addition to within-domain estimates, we examined contrasts among growth (Figure `r fig_nums("fig_diffgrow", display = "num")`) and mortality (Figure `r fig_nums("fig_diffmort", display = "num")`) across CWD domains. This allows us to establish if trees between two CWD domains are experiencing meaningful differences in growth and mortality. For example, this allows us to ask whether trees in historically dry sites (High initial CWD) respond differently to increased water stress (Increasing, or above threshold delta CWD) compared to trees in mesic sites (Low initial CWD) experiencing similar increases, among other comparisons.

We obtained empirical means and 95% confidence intervals via a nonparametric bootstrap. For each of the 1000 bootstrap iterations we sampled FIA plots with replacement and obtained point estimates for each domain, yielding a maximum of `r n.spp * 6 ` in-domain point estimates for each iteration.  For each domain we calculated a mean, upper, and lower 95% CI with the 2.5% and 97.5% quantiles from the  bootstrap iterations.  This approach allows for asymmetrical CIs around means.  We estimated contrasts between pairs of CWD domains by subtracting within-iteration domain point estimates.

Our  contrast analysis approach resulted in hundreds of comparisons. Given this large number of tests, we caution readers to interpret individual significant results (i.e., contrast 95% CIs that do not including zero) cautiously, as some proportion likely represent Type-I errors and occurred by chance.  We did not apply a multiple comparisons correction as our analysis is inherently exploratory, aimed at identifying patterns within and across species rather than testing a specific *a priori* hypothesis.  We did not want to apply corrections because we did not want to inflate the Type-II error rate and obscure real relationships in favor of reducing false positives.

## Dashboard
We developed a data dashboard (https://tinyurl.com/FIAgrowmort) to enable ourselves and readers to examine individual species domain estimates, the distribution of initial and delta CWD values, and the geographical distribution of plots and plot domains.  All analysis files, manuscript markdowns, dashboard code, and ancillary data sources are available at the GitHub repository https://github.com/jeremygroom/Tree_Growth_Death.

# Results

Our study tracked the fate of `r comma(analysis.stats$tree.num)` trees from `r n.spp` species in `r comma(analysis.stats$plot.num)` plots that were visited twice by FIA crews. Of the `r n.spp` species, `r analysis.stats$n.gym.spp` were gymnosperms.  The two visits were 10 years apart for `r pct.remeas[2]`% of plots (`r pct.remeas[1]`% were 9 years, `r pct.remeas[3]`% were 11 years). Plots were distributed across `r analysis.stats$strat.num` strata spanning forested regions in all three states, from coastal temperate rainforest to semi-arid interior woodlands.

## Climatic Water Deficit

Delta CWD increased at `r analysis.stats$cwd.gt0.pct`% of plots. The increases occurred in all forested regions of the three states (Figure `r fig_nums("fig_cwdmap", display = "num")`C).  Plots where CWD decreased (indicated by negative delta CWD values) are presented alone in Figure S1.  For Washington and Oregon negative CWD values were almost entirely found in the Coast Ranges and the mesic west slope of the Cascades Range crest.  The Blue Mountains and Northern and northeastern Washington lacked negative delta CWD plots entirely. In California negative delta CWD plots with values < 1mm were primarily found in the northern Sierra Nevada and in the southern portions of the state. Overall, initial and delta CWD values were weakly and positively linearly associated (correlation = `r CWD.lin.cor`).The two variables were better modeled with a 3rd order polynomial although the R^2^ value still indicated that the variance was not well explained by the model ($\Delta AIC$ =  `r CWD.dAIC`, R^2^ = `r round(CWD.rels$lm2$r.squared, 2)`, Figure S2, see model fits in Tables S1 and S2), implying that other variables likely affect the relationship between initial and delta CWD. For most species delta CWD neither linearly related to nor monotonically increased with initial CWD values (Figure S3; see dashboard). For instance, the relationship between the initial and delta CWD values for *Calocedrus decurrens* may be described overall with a concave curve, with some of the most negative delta CWD values occurring in the High domain (along with some of largest delta CWD values; Figures `r fig_nums("fig_domains", display = "num")`C).

The distribution of initial CWD values across the three states (Figure `r fig_nums("fig_cwdmap", display = "num")`A) reflected total annual precipitation normals for 1991-2000 (check against https://www.climate.gov/media/13728).  The most mesic areas, the Coast Ranges and the west slope of the Cascade Range, coincided with the areas with the lowest initial CWD values.

Initial CWD means and ranges differed among tree species (Table S3).  *`r init.mean.sm$spp.out`* had the smallest mean value for initial CWD (`r init.mean.sm$mean.alone` mm) while *`r init.range.sm$spp.out`* had both the largest mean value (`r init.mean.lg$mean.alone` mm) and smallest range `r init.range.sm$spp.range`.  *`r init.range.lg$spp.out`* occupied the plots with the largest range of initial CWD values `r init.range.lg$spp.range` and was the most widely distributed of our species (dashboard).

Species did not differ greatly by delta CWD (Table S4).   The riparian species *`r delta.mean.sm$spp.out`* had the smallest mean delta CWD value of `r delta.mean.sm$mean.alone` mm `r delta.mean.sm$spp.range`.  The species with the second smallest mean (`r delta.mean.sm2$mean.alone`mm) and smallest range of delta CWD values `r delta.mean.sm2$spp.range` was the largely coastal *`r delta.mean.sm2$spp.out`*.  The mostly high-elevation *`r delta.mean.lg$spp.out`* had the greatest mean increase in CWD values of `r delta.mean.lg$mean.alone` mm `r delta.mean.lg$spp.range`.  *`r delta.range.lg$spp.out`*, another high-elevation species, had the greatest range in delta CWD values `r delta.range.lg$spp.range`.

## Growth

Overall growth rates differed by species but in some instances coincided with species’ ecological traits.  For instance, the drought-tolerant hardwoods (*Arbutus menziesii*, *Notholithocarpus densiflorus*, *Quercus chrysolepis*, and *Quercus kelloggii*) exhibited the lowest growth rate, with all mean decadal basal area increases < 75 cm^2^ (Table 1).  *Quercus chrysolepis* had the least mean growth at 40.5 cm^2^/decade.  Riparian and mesic hardwoods like *Acer macrophyllum* and *Alnus rubra* on average grew by 145 cm^2^/decade, although *Acer macrophyllum* growth varied by state; its mean growth in California was 52.3 cm^2^/decade, less than half of its growth rate in Oregon and almost four times less than its growth rate in Washington.  High-elevation conifers like *Abies concolor* and *Picea engelmannii* would both have grown at similar rates of approximately 170-190 cm^2^/decade except that the Washington rate for *Picea engelmannii* was only 125.4 cm/decade.  The other high-elevation conifers, *Abies lasiocarpa*,  and *Tsuga mertensiana*, grew more slowly at almost 70 cm^2^/decade.

Shade tolerant conifers (*Abies amabilis*, *Abies grandis*, *Pinus lambertiana*, *Thuja plicata*, and *Tsuga heterophylla*) grew at moderate (109.9 cm^2^/decade for *Abies amablis*) to high rates (*Pinus lambertiana* had the highest overall growth rate of 253.6 cm^2^/decade). The fire-adapted conifers exhibited a range of mean growth rates, with *Pseudotsuga menziesii* having the greatest mean growth (229.9 cm^2^/decade), *Pinus monticola* and *Larix occidentalis* exhibiting moderate growth (124.8 and 106.4 cm^2^/decade, respectively), and *Pinus contorta* with the lowest mean growth rate (69.3 cm^2^/decade) The three state confidence intervals for *Pinus contorta* did not overlap; growth was greatest in California (83.8 cm^2^/decade) and least in Washington (57.0 cm^2^/decade).  *Pinus monticola* exhibited a somewhat reversed pattern, with the greatest growth in Washington (172.4 cm^2^/decade) and less growth in Oregon and California (approximately 116 cm^2^/decade for both).  The means for drought-tolerant conifers (*Calocedrus decurrens*, *Juniperus occidentalis*, *Pinus jeffreyi*, and *Pinus ponderosa*) were above 140 cm^2^/decade except for *Juniperus occidentalis* with a rate of 68.7 cm^2^/decade.

Individual species growth and mortality domain patterns could differ (e.g., Figures `r fig_nums("fig_domains", display = "num")`A and `r fig_nums("fig_domains", display = "num")`B). Growth domain contrasts (Figure `r fig_nums("fig_diffgrow", display = "num")`, `r t.ref("tbl_summ_grow_mort")`) did not appear to result in substantially more significant  negative or positive estimates, indicating that species overall did not exhibit more or less growth in the Increasing, Low, or High domains. The figure column “Increasing vs. Stable” displays the mean point estimate and 95% CI for contrast between corresponding Increasing and Stable domains (e.g., for *Abies lasiocarpa*, subtracting the estimates for columns SL, SM, and SH from IL, IM, IH, respectively; Figure `r fig_nums("fig_domains", display = "num")`A). Negative estimates for the “Increasing vs. Stable” in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")` indicate that the 10-year growth or mortality rate was greater for a species in plots below the change estimates. `r t.ref("tbl_summ_grow_mort")` summarizes the number of domain differences in each of the columns in Figures `r fig_nums("fig_diffgrow", display = "num")` and `r fig_nums("fig_diffmort", display = "num")` where the 95% CI does not include zero as well as the number of contrasts per figure column.

We observed varying growth responses across the CWD domains (Figure `r fig_nums("fig_diffgrow", display = "num")`). Depending on the species, growth may have been susceptible (negatively or positively) to a change in CWD and/or differed by initial CWD values. We consider a lower initial CWD value to indicate a more favorable condition for plant growth, and indeed *Abies amabilis*, *Abies grandis*, *Calocedrus decurrens*, *Pinus ponderosa*, *Pseudotsuga menziesii*, *Tsuga mertensiana*, *Acer macrophyllum*, and *Alnus rubra* exhibited greater growth in the Low domains than the Medium and/or High domains. This was exemplified by *Acer macrophyllum* and *Abies grandis* with a mean difference of `r fig.val.fcn("g", "X312", "IH - IL", 0)` cm^2^/decade, IH-IL and `r fig.val.fcn("g", "X312", "SH - SL", 0)` cm^2^/decade, SH-SL, respectively while contrasts for species like *Tsuga mertensiana* and *Quercus chrysolepis* were less (-0.20 for IM-IL and SM-SL, and `r fig.val.fcn("g", "X805", "SH - SM", 0)` for SH-SM, respectively).  Except for *Alnus rubra*, these results predominantly occurred in both the Stable domain (2 spp.) or the Increasing and Stable domains (5 spp.). However, this finding is not uniform for all species, and we found that *Thuja plicata* had greater growth in the Increasing (`r fig.val.fcn("g", "X242", "IH - IM", 0)` cm^2^/decade for IH-IM and `r fig.val.fcn("g", "X242", "IM - IL", 0)` cm^2^/decade for IM-IL) and Stable (`r fig.val.fcn("g", "X242", "SH - SM", 0)` cm^2^/decade for SH-SM and `r fig.val.fcn("g", "X242", "SM - SL", 0)` cm^2^/decade for SM-SL) Medium domains than either High or Low domains.  *Abies concolor*, *Larix occidentalis*, *Picea engelmannii*, *Pinus contorta*, *Pinus lambertiana*, *Arbutus menziesii*, and *Quercus kelloggii* had greater growth in the more water stressed High domains than the Medium and/or Low domains, with some of the greatest differences for *Pinus lambertiana* (`r fig.val.fcn("g", "X117", "IH - IL", 0)` cm^2^/decade for IH-IL) and *Picea engelmannii* (`r fig.val.fcn("g", "X93", "SH - SL", 0)` cm^2^/decade for SH-SL).  Except for *Larix occidentalis*, these results predominantly occurred in both the Increasing domain (3 spp.) or the Increasing and Stable domains (3 spp.). Other species exhibited no significant findings (4 spp.) or displayed other result patterns (3 spp.).

## Mortality

Overall decadal mean mortality was highest among high elevation conifers.  Three out of the four species had mortality rates of 0.20/decade or higher, including *Picea engelmannii* with the highest mean mortality rate of 0.28/decade.  The same species exhibited a mean mortality rate that was almost five times greater in Washington (0.34) than in Oregon (0.07).  Drought-tolerant conifers had overall mortality rates that were relatively low (*Calocedrus decurrens* had the greatest mortality of the four species at 0.15/decade), although *Calocedrus decurrens* and *Pinus ponderosa* both exhibited higher mortality and lower growth rates in California than Oregon. The mean mortality rates for drought-tolerant hardwoods ranged from 0.14 to 0.20, with *Quercus chrysolepis* and *Quercus kelloggii* having the highest mortality rates (0.20 and 0.18, respectively).  For the fire-adapted conifers, *Pinus monticola* and *Pinus contorta* had the greatest mortality rates (0.19 and 0.20, respectively), with *Pinus contorta* exhibiting a low decadal mortality rate in California (0.07), more than twice that rate in Oregon (0.17), and five times the California rate in Washington (0.37).  For the riparian and mesic hardwoods, *Alnus rubra* had an overall mortality rate of 0.21, compared 0.13 for *Acer macrophyllum*.  However, *Acer macrophyllum* exhibited a mortality rate of 0.27 in California, almost three times as high as rates for Washington and Oregon.  Mortality rates for shade-tolerant conifers varied greatly, from 0.04 (*Thuja plicata*) to 0.20 (*Pinus lambertiana*).  The Washington mortality rate for *Abies grandis* was approximately twice as high as mortality rates for Oregon and California.

A  suite of species had greater mortality in the Low domains than the Medium and/or High domains including *Larix occidentalis*, *Picea engelmannii*, *Pinus contorta*, *Pinus monticola*, and *Thuja plicata*. Some of the larger mean mortality contrasts included *Picea engelmannii* and *Pinus contorta* (`r fig.val.fcn("m", "X93", "IH - IL", 2)` and `r fig.val.fcn("m", "X108", "IH - IL", 2)`/decade, respectively, for IH-IL) while *Thuja plicata* represented a smaller estimate (`r fig.val.fcn("m", "X242", "IH - IL", 2)`/decade for IH-IL).  These results occurred in Increasing (3 spp.) or both Increasing and Stable (2 spp.) domains. Mortality rates for *Abies lasiocarpa*, *Calocedrus decurrens*, *Pinus lambertiana*, *Pinus ponderosa*, *Pseudotsuga menziesii*, *Tsuga heterophylla*, *Tsuga mertensiana*, *Acer macrophyllum*, and *Notolithocarpus densiflorus* were greater in the High domains than the Medium and/or Low domains.  Larger estimates include `r fig.val.fcn("m", "X19", "SH - SL", 2)`/decade for the *Abies lasiocarpa* SH-SL contrast, and `r fig.val.fcn("m", "X117", "IH - IM", 2)`/decade for the *Pinus lambertiana* IH-IM contrast. *Pseudotsuga mensiesii* exhibited smaller contrast estimates of around `r fig.val.fcn("m", "X202", "SH - SM", 2)`/decade for SH-SL and SH-SM.  These species were approximately evenly split between differences being present in the Increasing (4 spp.) or Stable (5 spp.) domains, but none with significant findings in both the Increasing and Stable domains. *Arbutus menziesii* had greater mortality in Medium domains than the High domains (approximately `r fig.val.fcn("m", "X361", "IH - IM", 2)`/decade for IH-IM and SH-SM). Five species had no significant mortality contrasts.

Mortality domain contrasts are  similar in certain regards to the growth contrasts: results for the three types of domain contrasts did not appear to be predominantly positive or negative (Figure `r fig_nums("fig_diffmort", display = "num")`, `r t.ref("tbl_summ_grow_mort")`). There were fewer mortality estimates with significant findings than growth (46 vs 74). There were fewer differences for “Increasing vs. Stable” contrasts (9) than for Increasing (21) or Stable (16), indicating that mortality may not be as strongly associated with delta CWD as by the initial CWD value.

## Geography  
The interactive dashboard revealed geographic patterns that were associated with specific domains while not necessarily captured by state-level estimates. For the Increasing and Stable Low domains *Abies lasiocarpa* had less mortality in northeastern Washington and *Tsuga mertensiana* exhibited greater growth in the same region. *Pinus contorta* exhibited less growth and greater mortality in northern Washington (Increasing Low). In the southern Sierra Nevada *Pinus lambertiana* had greater growth (Increasing Medium and High) and greater mortality (Increasing High), *Calocedrus decurrens* had greater mortality (Increasing High), and *Pinus monticola* had less mortality (Increasing High). *Acer macrophyllum* exhibited less growth in the Sierra Nevada and southern Klamath Mountains (Increasing High) and greater growth in northwestern Washington in the Coast and Cascade ranges (Increasing and Stable Low).

# Discussion

Analysis of 23 tree species revealed diverse, species-specific associations with climatic water deficit, with different species exhibiting contrasting relationships among growth, mortality, and water stress. Responses were potentially driven by differences in drought tolerance strategies, local disturbance regimes, and biotic interactions. We found more significant CWD contrasts for species growth than mortality, possibly indicating that growth differences by CWD domain were more readily detectable than mortality over a decade, although growth and mortality findings were not necessarily negatively related. Delta CWD increased almost everywhere between plot revisits although the magnitude of change was generally not large relative to initial CWD values. We additionally found concerningly high decadal mortality rates for some species within and across states.

Contrast results serve to highlight differences in mortality and growth rates for different subpopulations of trees undergoing different climatic experiences. Regional analyses detected several species with relatively high mortality rates including *Acer macrophyllum*, *Picea engelmannii*, *Abies lasiocarpa*, *Pinus contorta*, and *Alnus rubra*.  We do not believe our results are concerning for *Alnus rubra*.  Our estimates for this short-lived tree species align with reported background mortality rates (e.g., Brown and Courtin 2018).

The high mortality rate for *Acer macrophyllum* in California appears to be both unreported elsewhere and substantially greater than in Oregon or Washington.  *Acer macrophyllum* growth was negatively associated with initial CWD values, and its growth was lower and mortality almost three times higher in California than in Washington and Oregon.  We cannot find primary literature that corroborates an increase in mortality for this species in California. McNellis et al. (2021) estimated a recent *Acer macrophyllum* overall mortality rate of 0.45%/yr (4.5%/decade) while our estimates for Washington and Oregon were twice this level, and California's estimate was six times as large.  However, Betzen et al. (2021) found widespread evidence in Washington of *Acer macrophyllum* decline (crown dieback, thinning, discoloration, reduced leaf size) and reduced growth and increased mortality in locations with higher temperatures and water limitations.  Although these results align with our contrast findings for growth, as displayed in our dashboard, they are not reflective (yet) of our state mortality findings for Washington.  We encourage a more intensive assessment of *Acer macrophyllum*’s status in California.

Conspecifics *Abies lasiocarpa* and *Picea engelmannii* are frequently codominant and exhibit high mortality rates. Andrus et al. (2018) found that *Picea engelmannii* exhibited higher survival and longevity compared to *Abies lasiocarpa* in undisturbed plots in Colorado, with firs exhibiting a mortality rates of up to 1% annually while *Picea engelmannii* had mortality rates below 0.5% annually.  Mortality estimates for *Abies lasiocarpa* from Perret et al. (2023) approximated our state and overall estimates at 25% +/-2.7; they found the highest level of mortality in the eastern Oregon Cascades as well as in Idaho and Colorado. They attributed 83% of the mortality to fire or biogenic disturbance.  We found that the pressures on the two species appear to differ by CWD values and location.  We found that *Picea engelmannii* had lower mortality in the Increasing High domain, corresponding to stands in the Blue Mountains of Oregon and Washington, and higher mortality in the Increasing Low and Increasing High domains, largely coinciding with northern and northeastern Washington, where spruce beetle infestations were elevated (Washington Department of Natural Resources [WADNR] 2023).  Its mortality rate in Washington was almost 5 times higher than in Oregon.  *Abies lasiocarpa* did not have substantial mortality contrast findings (some evidence of BH - BL difference, but the MC permutation p-value was greater than 0.001) and its mortality was high but similar across all three states.

Our findings for *Pinus contorta* in Washington likely reflected insect pressures on the species.  Our Washington mortality estimate was twice that of Oregon’s and five times greater than Calfornia’s, while the growth confidence intervals for Washington were lower than California’s and did not overlap. The domains with the highest mortality and lowest growth align well with Washington regions that have experienced water stress (Raymond et al. 2023).  Pine bark beetles (*Dendroctonus ponderosae*, *Dendroctonus brevicomis*, and *Ips* spp.) have affected *Pinus contorta* for the past decade in this region (WADNR 2023).  This area also corresponds, at the level of ecoregion subsections, with the portions of *Pinus contorta*’s range in the lowest forest stability index category of Stanke et al. (2021) while California’s Sierra Nevada and the Oregon and Washington Cascade Range correspond to positive stability index values.  Stanke et al. also found, from FIA plot data, that insects (followed by fire) were related to the percent change in relative tree density.

The absence of consistent contrast patterns across species is itself informative, indicating that tree responses to water stress conditions are potentially mediated by local adaptation to conditions, species-specific physiological traits, and local environmental conditions rather than following a general response curve.  The number of significant Increasing vs. Stable threshold contrasts for growth and mortality were approximately evenly distributed between positive and negative contrast estimates, with a positive difference indicating that some species exhibited higher growth or greater mortality in an Increasing domain (e.g., *Pinus lambertiana* for both growth and mortality) versus a Stable domain (*Abies grandis* and *Pinus contorta*, respectively, for growth and mortality).  This lack of a pattern across species may indicate that the distinction of Increasing vs. Stable is not generalizable across species, or that significant findings are due to confounding factors such as geographical locations with different ecological pressures.

*Abies grandis* exhibited higher mortality in Washington than Oregon or California. We found no evidence of a difference in domain mortality geographically, although the greatest growth domain was Stable Low, which corresponded to the more mesic plots west of the Cascade Crest of Washington and Oregon.  The mortality result may be due to an intersection in a fire suppression history promoting a densification of shade-tolerant but drought-intolerant firs in the arid climates of eastern Oregon and Washington, drought, and fir engraver beetle mortality (*Scolytus ventralis*; Washington DNR 2023, Merschel et al. 2021).  Johnson et al. (2025) report mortality rates of *Abies grandis* in protected areas in the Malheur National Forest of Oregon that exceed ours, primarily attributing the cause to fire.

Our dashboard reveals some instances of domains representing geographical mortality differences.  State estimates for *Pinus monticola* mortality do not appear to differ. Mortality for the Increasing High domain was low (see dashboard; MC permutation test results only found p < 0.01 when comparing Increasing High to Increasing Low) and predominantly represented by plots in the southern Sierra Nevada mountains of California.  *Pinus lambertiana* and *Calocedrus decurrens* exhibited high mortality in approximately the same region, a result supported by Fettig et al. (2019).  This region experienced an extreme drought that killed, as estimated from aerial detection surveys, 170 million trees between 2010 and 2021 (Office of Environmental Health Hazard Assessment 2022).  Dudney et al. (2020) also estimated lower mortality rates for *Pinus monticola* than for *Pinus lambertiana* in the Sequoia and Kings Canyon National Parks, which are within this region.  They attributed some of the difference in mortality to elevation, with mountain pine beetle (*Dendroctonus ponderosae*), fire, and white pine blister rust (*Cronartium ribicola*) more prevalent at lower elevations, corresponding to the range of *Pinus lambertiana*.

Our results for *Thuja plicata* did not corroborate with those from other researchers who are currently concerned about mortality in this species. Andrus et al. (2023) sampled *Thuja plicata* at nine locations in Washington, northern Oregon, and Idaho that were without canopy dieback, exhibited canopy dieback, or had died. They found that the mortality was associated with exceptionally warm and dry climate conditions. A collaboration of state and federal agencies is currently reaching out to the public to identify dieback sites for this species (website: Goodrich et al. 2025).  Our analysis found that *Thuja plicata* exhibited the lowest decadal mortality rate and third highest growth rate of any species we examined, and relatively strong sampling precision. It also had its greatest growth in Medium domains. Our analysis may not be temporally sensitive enough to capture the same recent findings as Andrus et al., given that our FIA plot revisit data represented plot visits between 2011 to 2023. If species conditions change quickly (on the order of 3-5 years) the FIA plot data may not reflect the change until a full or nearly full panel of plots is complete following the change in conditions. Another non-exclusive possibility is that our results, based on an extensive, unbiased network of plots, offer a robust and unbiased mortality estimate for this species and others. *Thuja plicata* may indeed be senescing in specific regions, but other species may be overall faring worse.

We believe the value of our analysis lies in its application to a relatively large number of tree species and its approach for investigating life history conditions such as growth and mortality that require the documentation of tree condition change.  The availability of full-panel FIA revisit data in the states we investigated is a relatively recent development.  Prior to the availability of the revisit data, Monleon and Lintz (2015) assessed tree species’ range shifts using the single-visit data set, comparing mean plot temperature values for sapling and mature trees.  Groom and Monleon (2023) accessed to the more complete FIA revisit data set and found evidence of tree species density shifts along initial values of climatic variables but documented evidence of interactions between those variables’ initial and change values.   The literature of model- and design-based estimation of FIA revisit data is growing.  Examples of model-based analyses of mortality and growth using FIA revisit data include Shriver et al. (2022), Ogana et al. (2025), and Brandeis and Brandeis (2021).  Examples of growth and mortality design-based estimation that rely on FIA revisit data, similar to our state and overall analyses, include Vickers and Goff (2025), Goeking and Windmuller-Campione (2021), and Perret et al. (2023). Our examination approach for plot initial and delta CWD values shares some dimensional similarities with Dobrowski et al. (2013)  who examined the velocity of CWD (change in distance over time) along with AET and minimum temperatures. Our approach allows us to capitalize on the amenability of the data to design-based estimation while accounting for the need, when examining revisit conditions, of capturing the initial and change value conditions.

CWD is a defensible variable to use for examining regional water stress as it integrates climatic and soil water availability through considering precipitation and air temperatures (Stephenson 1990, 1998).  CWD correlates well with tree growth (Dyer 2004, Restaino et al. 2016) and mortality (Anderegg et al. 2015, Das et al. 2013, Hammond et al. 2022).  However, there are some challenges in relying on CWD or any other climate variable in a design-based regional analysis such as this one. The challenges stem from the remeasurement timeframe, using a fixed delta CWD domain threshold, the broad, fixed scope of our analysis, and the lack of incorporation of local-scale conditions.

We did not detect sizable changes in CWD between visits, potentially signaling that 10 years is too soon to observe CWD-related species responses (perhaps more true for mortality than growth). Another possibility is that the TerraClimate dataset is relatively spatially coarse (~4 km, Abatzoglou et al. 2018), potentially dampening the range of possible results from a finer resolution model. The CWD changes we observed may have represented greater proportional change for some species than others, with species such as *Abies amablis*, *Alnus rubra*, and *Juniperus occidentalis* exhibiting narrow and low water stress initial CWD ranges. CWD effects can manifest in direct and indirect ways, potentially differing with species by geographic location and the biotic and abiotic impacts to those locations.  It also cannot easily convey differences in impacts to trees within the same stands.

The scale of our analysis and the lack of incorporation of other variables likely limited our analysis because the evidence of water stress can manifest in multiple overlapping ways.  Water stress alone can kill trees (Anderegg et al. 2015, Maherali et al.  2004, and van Nieuwstadt and Sheil 2005), but it can also increase the probability of insect (Hart et al. 2014, Preisler et al. 2012 , examples and counterexamples in Huberty and Denno 2004) and wildfire-caused mortality (Abatzoglou and Williams 2016). With reference to Manion’s (1991) “death spiral” (see also Bennett et al. 2023 for a CWD-related version), teasing apart the cause of mortality from predisposing, inciting, and contributing factors is difficult at best, as these factors can interact (Gaylord et al. 2015).  Our regional analysis necessarily aggregates across local-scale variation in interrelated predisposing conditions for individual tree mortality such as canopy openness (Lemmo et al. 2024), structural complexity (Ma et al. 2023), tree height (Stovall et al. 2019), and within-stand competition (Ford et al. 2017). For management applications, particularly those for individual stands, our regional and domain estimates should be viewed as general results that may be modified by these local factors.

Although our current analysis offers a coarse examination of species status across a large area, future analyses could refine our approach for assessing tree species growth and mortality by domains.  Our climate domains could be divided further into sub-domains: trees could be assigned to tree size, stand competition, plot aspect, or plot elevation categories within the climate domain categories.  This approach has its limits: only estimates for certain widespread tree species would be reasonable to report, and domains could not be divided too finely into sub-domains due to sample size constraints, limiting the number of domain variables that could be considered simultaneously.  However, this approach would permit design-based estimation at a resolution better attuned to capturing factors affecting growth and mortality.

Our analysis captures significant changes in mortality and growth for the tree species we examined, with some trees exhibiting increases in mortality and decreases in growth either in portions of their ranges or across their entire range.  For some species in semi-arid environments such as *Pinus ponderosa*, *Pseudotsuga menziesii*, *Pinus contorta*, *Picea engelmannii*, and *Abies lasiocarpa*, we may be observing ecosystem shifts with a change in water stress combined with disturbances such as wildfire and insect mortality (Halofsky et al. 2020, Parks et al. 2019, Raymond et al. 2023).  Our hope is that our results for individual species serve to bolster or direct others' inquiries into the status of the species we examined.  For managers, the results may be used to inform action for resiliency.   With the shift in ecosystems, landowners can assist with species migration when planting or replanting (Raymond et al. 2023), and forest thinning can help reduce water stress and the impacts of fire and insects (Bernal et al. 2023).

# Conclusion

Our analysis serves as an example of design-based estimation, where we use estimation domains based on initial CWD and delta CWD values of plots.  For the 23 species we examined, we found no general patterns in domains with higher or lower tree growth or mortality; the results appeared species specific.  We believe our analysis approach can be refined for individual species analysis, with the potential for integrating a third variable that could introduce information besides a condition variable and change variable.  Our analysis and the analyses of others provide valuable and reliable insights into the condition of several tree species that exhibit abnormally high mortality (and low growth).  Our hope is that these findings will inform management and encourage focused monitoring efforts for these species in the rapidly changing landscape of western North America.

# Acknowledgments

# References

::: {#refs}
Abatzoglou, J.T. and Williams, A.P., 2016. Impact of anthropogenic climate change on wildfire across western US forests. Proceedings of the National Academy of sciences, 113(42), pp.11770-11775.

Abatzoglou, J.T., Dobrowski, S.Z., Parks, S.A. and Hegewisch, K.C., 2018. TerraClimate, a high-resolution global dataset of monthly climate and climatic water balance from 1958–2015. Scientific data, 5(1), pp.1-12.

Adams, H.D., Macalady, A.K., Breshears, D.D., Allen, C.D., Stephenson, N.L., Saleska, S.R., Huxman, T.E. and McDowell, N.G., 2010. Climate‐induced tree mortality: Earth system consequences. Eos, Transactions American Geophysical Union, 91(17), pp.153-154.

Adams, H.D., Zeppel, M.J., Anderegg, W.R., Hartmann, H., Landhäusser, S.M., Tissue, D.T., Huxman, T.E., Hudson, P.J., Franz, T.E., Allen, C.D. and Anderegg, L.D., 2017. A multi-species synthesis of physiological mechanisms in drought-induced tree mortality. Nature ecology & evolution, 1(9), pp.1285-1291.

Alexander, R.R., Shearer, R.C. and Shepperd, W.D., 1990. Abies lasiocarpa (Hook.) Nutt. subalpine fir. Silvics of North America, 1, pp.60-70.

Anderegg, W.R., Flint, A., Huang, C.Y., Flint, L., Berry, J.A., Davis, F.W., Sperry, J.S. and Field, C.B., 2015. Tree mortality predicted from drought-induced vascular damage. Nature geoscience, 8(5), pp.367-371.

Andrus, R.A., Harvey, B.J., Chai, R.K. and Veblen, T.T., 2018. Different vital rates of Engelmann spruce and subalpine fir explain discordance in understory and overstory dominance. Canadian Journal of Forest Research, 48(12), pp.1554-1562.

Bechtold, W.A., Scott, C.T. 2005. The Forest Inventory and Analysis plot design, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Bennett, M., Shaw, D.C. and Lowrey, L., 2023. Recent Douglas-fir mortality in the Klamath Mountains ecoregion of Oregon: Evidence for a decline spiral. Journal of Forestry, 121(3), pp.246-261.

Betzen, J.J., Ramsey, A., Omdal, D., Ettl, G.J. and Tobin, P.C., 2021. Bigleaf maple, Acer macrophyllum Pursh, decline in western Washington, USA. Forest Ecology and Management, 501, p.119681.

Brandeis, T. and Brandeis, C., 2021. Assessing tree mortality probability in harvested hardwood stands using long-term forest inventory data. Forest Science, 67(2), pp.231-240.

Breshears, D.D., Myers, O.B., Meyer, C.W., Barnes, F.J., Zou, C.B., Allen, C.D., McDowell, N.G. and Pockman, W.T., 2009. Tree die‐off in response to global change‐type drought: mortality insights from a decade of plant water potential measurements. Frontiers in Ecology and the Environment, 7(4), pp.185-189.

Brown, K.R. and Courtin, P.J., 2018. Can phosphorus additions increase long-term growth and survival of red alder (Alnus rubra Bong.) on periodically dry sites?. Forest Ecology and Management, 430, pp.545-557.

Burns, R.M., and B.H. Honkala, tech coords. 1990a. Silvics of North America. Volume 1. Conifers. Agricultural Handbook 654. U.S. Department of Agriculture, Forest Service, Washington DC. 675p.

Burns, R.M., and B.H. Honkala, tech coords. 1990b. Silvics of North America. Volume 2. Hardwoods. Agricultural Handbook 654. U.S. Department of Agriculture, Forest Service, Washington DC. 877p.

Daly C, Halbleib M, Smith JI, Gibson WP, Doggett MK, Taylor GH, et al. Physiographically sensitive mapping of climatological temperature and precipitation across the conterminous United States. Int J Climatol. 2008; 28(15):2031–64. doi: 10.1002/joc.1688 PMID: WOS:000261810700004.

Das, A.J., Stephenson, N.L., Flint, A., Das, T. and van Mantgem, P.J., 2013. Climatic correlates of tree mortality in water-and energy-limited forests. PLoS One, 8(7), p.e69917.

Dobrowski, S.Z., Abatzoglou, J., Swanson, A.K., Greenberg, J.A., Mynsberge, A.R., Holden, Z.A. and Schwartz, M.K., 2013. The climate velocity of the contiguous United States during the 20th century. Global change biology, 19(1), pp.241-251.

Dudney, J.C., Nesmith, J.C., Cahill, M.C., Cribbs, J.E., Duriscoe, D.M., Das, A.J., Stephenson, N.L. and Battles, J.J., 2020. Compounding effects of white pine blister rust, mountain pine beetle, and fire threaten four white pine species. Ecosphere, 11(10), p.e03263.

Dyer, J.M., 2004. A water budget approach to predicting tree species growth and abundance, utilizing paleoclimatology sources. Climate Research, 28(1), pp.1-10.

Fettig, C.J., Mortenson, L.A., Bulaon, B.M. and Foulk, P.B., 2019. Tree mortality following drought in the central and southern Sierra Nevada, California, US. Forest Ecology and Management, 432, pp.164-178.

Ford, K.R., Breckheimer, I.K., Franklin, J.F., Freund, J.A., Kroiss, S.J., Larson, A.J., Theobald, E.J. and HilleRisLambers, J., 2017. Competition alters tree growth responses to climate at individual and stand scales. Canadian Journal of Forest Research, 47(1), pp.53-62.

Gaylord, M.L., Kolb, T.E. and McDowell, N.G., 2015. Mechanisms of piñon pine mortality after severe drought: a retrospective study of mature trees. Tree physiology, 35(8), pp.806-816.

Goeking, S.A. and Windmuller-Campione, M.A., 2021. Comparative species assessments of five-needle pines throughout the western United States. Forest Ecology and Management, 496, p.119438.

Gomez‐Gallego, M., Galiano, L., Martínez‐Vilalta, J., Stenlid, J., Capador‐Barreto, H.D., Elfstrand, M., Camarero, J.J. & Oliva, J. (2022) Interaction of drought‐ and pathogen‐induced mortality in Norway spruce and Scots pine. Plant, Cell & Environment, 45, 2292–2305. https://doi.org/10.1111/pce.14360

Goodrich, B., Fischer, M., Buhl, C.  2025. Western redcedar dieback.  https://storymaps.arcgis.com/stories/1405dab5f59246aa83849ec43f72b15a. Accessed 11 November 2025.

Groom, J.D. and Monleon, V.J., 2023. West coast US tree species show stronger evidence for shifts in density than range. Forest Ecology and Management, 545, p.121262. 
Hammond, W.M., Williams, A.P., Abatzoglou, J.T., Adams, H.D., Klein, T., López, R., Sáenz-Romero, C., Hartmann, H., Breshears, D.D. and Allen, C.D., 2022. Global field observations of tree die-off reveal hotter-drought fingerprint for Earth’s forests. Nature communications, 13(1), p.1761. https://doi.org/10.1038/s41467-022-29289-2

Hart, S.J., Veblen, T.T., Eisenhart, K.S., Jarvis, D. and Kulakowski, D., 2014. Drought induces spruce beetle (Dendroctonus rufipennis) outbreaks across northwestern Colorado. Ecology, 95(4), pp.930-939.

Hicke, J.A., Meddens, A.J. and Kolden, C.A., 2016. Recent tree mortality in the western United States from bark beetles and forest fires. Forest Science, 62(2), pp.141-153.

Huberty, A.F. and Denno, R.F., 2004. Plant water stress and its consequences for herbivorous insects: a new synthesis. Ecology, 85(5), pp.1383-1398.

Johnston, J.D., Stephens, P., Schmidt, M., Jeronimo, S., Brackett, A., Wuenschel, A., Ulrich, D.E., Paine, R., Duloisy, L., Greenler, S., Larson, A.J., and Churchill, D.J.  2025. Significant Mortality of Old Trees Across a Dry Forest Landscape, Oregon, USA. Forest Ecology and Management, Volume 593, 122907, ISSN 0378-1127, https://doi.org/10.1016/j.foreco.2025.122907.

Juang, C.S., Williams, A.P., Abatzoglou, J.T., Balch, J.K., Hurteau, M.D. and Moritz, M.A., 2022. Rapid growth of large forest fires drives the exponential response of annual forest‐fire area to aridity in the western United States. Geophysical Research Letters, 49(5), p.e2021GL097131.

Kolb, T.E., Fettig, C.J., Ayres, M.P., Bentz, B.J., Hicke, J.A., Mathiasen, R., Stewart, J.E. and Weed, A.S., 2016. Observed and anticipated impacts of drought on forest insects and diseases in the United States. Forest Ecology and Management, 380, pp.321-334.

Lemmo, S.L., Kerhoulas, L.P., Sherriff, R.L. and Beckmann, J.J., 2024. Drought effects on tree mortality and regeneration in northern California. Forest Ecology and Management, 563, p.121969.

Ma, Q., Su, Y., Niu, C., Ma, Q., Hu, T., Luo, X., Tai, X., Qiu, T., Zhang, Y., Bales, R.C. and Liu, L., 2023. Tree mortality during long-term droughts is lower in structurally complex forest stands. Nature communications, 14(1), p.7467.

Maherali, H., Pockman, W.T. and Jackson, R.B., 2004. Adaptive variation in the vulnerability of woody plants to xylem cavitation. Ecology, 85(8), pp.2184-2199.

Manion, P.D. 1991. Tree Disease Concepts, 2nd ed, 402. Englewood Cliffs, NJ: Prentice-Hall.

McDowell, N., Pockman, W.T., Allen, C.D., Breshears, D.D., Cobb, N., Kolb, T., Plaut, J., Sperry, J., West, A., Williams, D.G. and Yepez, E.A., 2008. Mechanisms of plant survival and mortality during drought: why do some plants survive while others succumb to drought?. New phytologist, 178(4), pp.719-739.

McNellis, B.E., Smith, A.M., Hudak, A.T. and Strand, E.K., 2021. Tree mortality in western US forests forecasted using forest inventory and Random Forest classification.  Ecosphere, 12(3), p.e03419.

McRoberts, R.E. 2005. The Enhanced Forest Inventory and Analysis Program, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Meddens, A.J., Hicke, J.A. and Ferguson, C.A., 2012. Spatiotemporal patterns of observed bark beetle‐caused tree mortality in British Columbia and the western United States. Ecological Applications, 22(7), pp.1876-1891.

Merschel, A.G., Beedlow, P.A., Shaw, D.C., Woodruff, D.R., Lee, E.H., Cline, S.P., Comeleo, R.L., Hagmann, R.K. and Reilly, M.J., 2021. An ecological perspective on living with fire in ponderosa pine forests of Oregon and Washington: Resistance, gone but not forgotten. Trees, Forests and People, 4, p.100074.

Monleon, V.J., Lintz, H.E., Delzon, S., 2015. Evidence of tree species’ range shifts in a complex landscape. PLoS One 10 (1), e0118069.

Negrón, J.F., McMillin, J.D., Anhold, J.A. and Coulson, D., 2009. Bark beetle-caused mortality in a drought-affected ponderosa pine landscape in Arizona, USA. Forest Ecology and Management, 257(4), pp.1353-1362.

Office of Environmental Health Hazard Assessment. 2022. Indicators of climate change in California, fourth edition, California Environmental Protection Agency.

Ogana, F.N., Green, P.C. and Radtke, P., 2025. Modeling the growth and yield of natural hardwood stands in the southern United States using the Forest Inventory and Analysis data. Forest Ecology and Management, 586, p.122722.

Parks, S.A., Coop, J.D. and Davis, K.T., 2025. Intensifying fire season aridity portends ongoing expansion of severe wildfire in western US forests. Global Change Biology, 31(8), p.e70429.

Perret, D.L., Bell, D.M., Gray, A.N., Shaw, J.D. and Zald, H.S., 2023. Range-wide population assessments for subalpine fir indicate widespread disturbance-driven decline. Forest Ecology and Management, 542, p.121128.

Preisler, H.K., Hicke, J.A., Ager, A.A. and Hayes, J.L., 2012. Climate and weather influences on spatial temporal patterns of mountain pine beetle populations in Washington and Oregon. Ecology, 93(11), pp.2421-2434.

Reams, G.A, Smith, W.D., Hansen, M.H., Bechtold, W.A., Roesch, F.A., Moisen, G.G. 2005. The forest inventory and analysis sampling frame, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Restaino, C.M., Peterson, D.L. and Littell, J., 2016. Increased water deficit decreases Douglas fir growth throughout western US forests. Proceedings of the National academy of Sciences, 113(34), pp.9557-9562.

Scott, C.T., W. A. Bechtold, G.A. Reams, W.D. Smith, J.A. Westfall, M.H. Hansen, Moisen, G.G. 2005. Sample-based estimators used by the Forest Inventory and Analysis National Information Management System, in: Bechtold, W.A., Patterson, P.L. (Eds.), The Enhanced Forest Inventory and Analysis Program – National Sampling Design and Estimation Procedures. Gen. Tech. Rep. U.S. Dep. Agric. For. Serv. SRS-80.

Shive, K.L., Wuenschel, A., Hardlund, L.J., Morris, S., Meyer, M.D. and Hood, S.M., 2022. Ancient trees and modern wildfires: Declining resilience to wildfire in the highly fire-adapted giant sequoia. Forest Ecology and Management, 511, p.120110.

Shriver, R.K., Yackulic, C.B., Bell, D.M., & Bradford, J.B. (2022). Dry forest decline is driven by both declining recruitment and increasing mortality in response to warm, dry conditions. Global Ecology and Biogeography, 31, 2259–2269. https://doi.org/10.1111/geb.13582

Soderberg, D.N., A.J. Das, N. L. Stephenson, M.D. Meyer, C.A. Brigham, and J. Flickinger. 2024. “Assessing giant sequoia Mortality and regeneration following high-severity wildfire.” Ecosphere 15(3): e4789. https://doi.org/10.1002/ecs2.4789

Stanke, H., Finley, A.O., Domke, G.M., Weed, A.S., MacFarlane, D.W., 2021. Over half of western United States’ most abundant tree species in decline. Nat. Commun. 12, 451. https://doi.org/10.1038/s41467-020-20678-z.

Stephenson, N.L., and C. Brigham. 2021. Preliminary estimates of sequoia mortality in the 2020 Castle Fire. National Park Service, Sequoia and Kings Canyon National Parks.

Stephenson, N.L., 1998. Actual evapotranspiration and deficit: biologically meaningful correlates of vegetation distribution across spatial scales. Journal of biogeography, 25(5), pp.855-870.

Stephenson, N.L., 1990. Climatic control of vegetation distribution: the role of the water balance. The American Naturalist, 135(5), pp.649-670.     
Sala, A., Piper, F. and Hoch, G., 2010. Physiological mechanisms of drought-induced tree mortality are far from being resolved. The New Phytologist, 186(2), pp.274-281.

Sala, A., Woodruff, D.R. and Meinzer, F.C., 2012. Carbon dynamics in trees: feast or famine?. Tree physiology, 32(6), pp.764-775.

Stovall, A.E., Shugart, H. and Yang, X., 2019. Tree height explains mortality risk during an intense drought. Nature communications, 10(1), p.4385.
Thompson, J. 2015. PNW-FIADB Users Manual: A data dictionary and user guide for the PNW-FIADB database. Forest Inventory and Analysis, Pacific Northwest Research Station, Portland, Oregon.

Van Mantgem, P.J., Stephenson, N.L., Byrne, J.C., Daniels, L.D., Franklin, J.F., Fulé, P.Z., Harmon, M.E., Larson, A.J., Smith, J.M., Taylor, A.H. and Veblen, T.T., 2009. Widespread increase of tree mortality rates in the western United States. Science, 323(5913), pp.521-524. DOI: 10.1126/science.1165000

Van Nieuwstadt, M.G. and Sheil, D., 2005. Drought, fire and tree survival in a Borneo rain forest, East Kalimantan, Indonesia. Journal of ecology, 93(1), pp.191-201.
Vickers, L.A. and Goff, T.C., 2025. White oak sapling dynamics across the eastern US: Quantifying a bottleneck to sustainability. Forest Ecology and Management, 598, p.123232.

Washington State Department of Natural Resources (WADNR) Forest Reliance Division. 2024. Forest health highlights in Washington: A summary of insect, disease, and other disturbance conditions affecting Washington’s forests.  Washington State Department of Natural Resources, Forest Resilience Division, Olympia, WA. https://dnr.wa.gov/sites/default/files/2025-03/rp_2023_fhh_report_final_wr.pdf, accessed 9/18/2025

Westerling ALR. 2016 Increasing western US forest wildfire activity: sensitivity to changes in the timing of spring. Phil. Trans. R. Soc. B 371: 20150178. http://dx.doi.org/10.1098/rstb.2015.0178

Woudenberg, S.W., Conkling, B.L., O'Connell, B.M., LaPoint, E.B., Turner, J.A. and Waddell, K.L., 2010. The Forest Inventory and Analysis Database: Database description and users manual version 4.0 for Phase 2. Gen. Tech. Rep. RMRS-GTR-245. Fort Collins, CO: US Department of Agriculture, Forest Service, Rocky Mountain Research Station. 336 p.

:::

<!---BLOCK_LANDSCAPE_START--->

```{r tbl-growth, tab.cap=tab.tbl_growth}

growth.ft

```

```{r tbl-mort, tab.cap = tab.tbl_mort}

mort.ft

```

<!---BLOCK_LANDSCAPE_STOP--->

```{r tbl-summ-grow-mort, tab.cap = tab.summ_grow_mort}

g.m.summ.ft

```

```{r mapfig, fig.height = 7,  fig.dpi=300, fig.cap=fig.cwdmap_cap, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "cwd_maps_figure.png"))

```

```{r domainsfig, fig.width = 6,  fig.dpi=300, fig.cap=fig.domains, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "Example_spp_figure.png"))

```

```{r cwdmeansfig, fig.width = 7, fig.height=6, fig.dpi=300, fig.cap=fig.cwd_initdelt, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS.OTHER, "Spp_InitDeltCWD.png"))

```

```{r diff_grow_fig, fig.height = 6.5, width = 6.5, fig.dpi=300, fig.cap=fig.diff_grow, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Growth_figs_cwd/Growth_Panel_Plot.png"))

```

```{r diff_mort_fig, fig.height = 6.5, width = 6.5, fig.dpi=300, fig.cap=fig.diff_mort, fig.cap.sep=""}

# Output from Manuscript_information.R
knitr::include_graphics(paste0(RESULTS1.LOC, "/Mort_figs_cwd/Mort_Panel_Plot.png"))

```